### 2019-03-15 scripts/00_datawrangling.Rmd
update data wrangling

---
title: "00_datawrangling.Rmd"
output: md_document
---

```{r setup, include=FALSE}
# https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html
library(tidyverse)
```

# Data Wrangling

## countData and geneinfo

```{r countData}
# import count data, set rows at entreziz
countData <- read.table("../results/kallistocounts.txt", 
                        sep = ",", row.names = NULL)

## set row names as entrezid for deseq2
row.names(countData) <- countData$entrezid

# replace dashes with periods (to make deseq happy)
colnames(countData) <- gsub('-', '.', colnames(countData))

# create df with gene info and remove this from countData
geneinfo <- countData %>%
  select(row.names, Name, geneid, entrezid)
head(geneinfo)
countData <- countData %>%
  select (-c(row.names, Name, geneid, entrezid))

# because DESeq require integers
countData <- round(countData)

# print tolal num of genes and samples
dim(countData)

# make a tibble for the sole purpose of previewing this huge df
countDataTibble <- as.tbl(countData)
print(countDataTibble, n_extra = 0)
```

## wrangle colData

```{r colData}
colData <- read.table(file.path( "../metadata/kallistosamples.txt"), header = F, stringsAsFactors = F)

# use strsplit to cut the filename into meaningful columns
colData$bird <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 1)
colData$sex <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 2)
colData$tissue <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 3)

# to create columns with variable names in the filename 
colData$temp <- sapply(strsplit(as.character(colData$V1),'\\_'), "[", 4)
colData$treatmentTemp <- sapply(strsplit(as.character(colData$temp),'\\.'), "[", 1)
colData$NYNO <- sapply(strsplit(as.character(colData$temp),'\\.'), "[", 2)

# relabel incorrectly named variables 
unique(colData$treatmentTemp)
colData$treatment <- ifelse(grepl("extend-hatch", colData$treatmentTemp), "extend", colData$treatmentTemp)

# drop no-longer-needed cols
colData <- colData %>% dplyr::select(-c(temp, treatmentTemp))

# replace dashes with periods and make rowname, (to make deseq happy)
colData$bird <- gsub("-", ".", colData$bird)
colData$treatment <- gsub("-", ".", colData$treatment)
colData$V1 <- gsub("-", ".", colData$V1)
row.names(colData) <- colData$V1

# make variables factors and specify level
cols <- c("sex", "treatment", "tissue")
colData[cols] <- lapply(colData[cols], factor) 
colData$tissue <- factor(colData$tissue, 
                         levels = c("hypothalamus", "pituitary", "gonad"))

# create a combinatorial variable
colData$group <- paste(colData$sex, colData$tissue, colData$treatment, sep=".")

## view colData
str(colData)
head(colData, 5)

# check that rownames and colnames match for DESeq
ncol(countData) == nrow(colData)
```

## subset for manipulation study

```{r manip}
# create a new dfs for just the manipulation study
colData_manipluation <- colData %>%
  dplyr::filter(grepl('m-|extend|prolong', treatment)) %>%
  droplevels()
row.names(colData_manipluation) <- colData_manipluation$V1

# print sample sizes
colData_manipluation %>% select(sex,treatment, tissue)  %>%  summary()

# create list with filenames and select countData columns matching the file names
savecols <- as.character(colData_manipluation$V1) 
savecols <- as.vector(savecols) 
countData_manipluation <- countData %>% dplyr::select(one_of(savecols))

# confirm that row and col lenghts are equal
ncol(countData_manipluation) == nrow(colData_manipluation)
```

## subset for characterization study

```{r characterization}
# filter out data related to the manipulation study
colData_characterization <- colData %>%
  dplyr::filter(!grepl('m-|extend|prolong', treatment)) %>%
  droplevels()
row.names(colData_characterization) <- colData_characterization$V1

# print sample sizes
colData_characterization %>% select(sex,treatment, tissue)  %>%  summary()

savecols <- as.character(colData_characterization$V1) 
savecols <- as.vector(savecols) 
countData_characterization <- countData %>% dplyr::select(one_of(savecols))

# check that row and col lenghts are equal
ncol(countData_characterization) == nrow(colData_characterization)
```

## save files for downstream use

```{r write}
write.csv(colData_manipluation, "../results/00_colData_manipluation.csv", row.names = TRUE) 
write.csv(countData_manipluation, "../results/00_countData_manipluation.csv", row.names = TRUE) 
write.csv(colData_characterization, "../results/00_colData_characterization.csv", row.names = TRUE) 
write.csv(countData_characterization, "../results/00_countData_characterization.csv", row.names = TRUE) 
write.csv(geneinfo, "../results/00_geneinfo.csv", row.names = TRUE)
```