### 2019-03-15 scripts/01_limma_all.Rmd
add limma pipeline fisrt attempt

---
title: "01_limma_voom"
output: md_document
---

DESeq2 is _not_ recommended for experiments with more than 100 samples ([see Mike Love's post](https://mikelove.wordpress.com/2016/09/28/deseq2-or-edger/)), so I decided to try the limma package. I followed [this tutorial](https://www.bioconductor.org/packages/devel/workflows/vignettes/RNAseq123/inst/doc/limmaWorkflow.html).

```{r}
library(limma)
library(Glimma)
library(edgeR)
library(kableExtra)

knitr::opts_chunk$set(fig.path = '../figures/',cache=TRUE)

```

First, I read in the data I processed in 00_datawrangling.Rmd.

```{r read}
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

Then, I followed the steps from <https://github.com/macmanes-lab/RockDove/blob/master/parental_care/parental_analysis.Rmd>.

```{r edgeR}
# create a large DGEList with 3 elements
parentalobject <- DGEList(counts=countData, genes=geneinfo, group=colData$group)

# transform raw counts to countspermillion
cpms <- cpm(parentalobject)

# calculate number of lowly lowly expressed genes and remove them
table(rowSums(parentalobject$counts==0)==10)
keep_genes <- rowSums(cpms >= 1) >= 10
dge <- parentalobject[keep_genes, ]

# magic
parentaldesign <- model.matrix(~ colData$group )
colnames(parentaldesign) <- levels(colData$group)
parentalobject <- calcNormFactors(parentalobject)
parentalobject <- estimateCommonDisp(parentalobject)
parentalobject <- estimateTagwiseDisp(parentalobject)
parentalobject <- estimateDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMCommonDisp(parentalobject, parentaldesign, verbose=TRUE)
parentalobject <- estimateGLMTrendedDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMTagwiseDisp(parentalobject, parentaldesign)
fit <- glmFit( parentalobject, parentaldesign, robust=T)
treatres <- glmTreat(fit, lfc = 1)
```

# specify contrasts

```{r 01-contrasts}
levels(colData$group)

my.contrasts <- makeContrasts(
	         FG_HL = female.gonad.hatch - female.gonad.lay,
	         FH_HL = female.hypothalamus.hatch - female.hypothalamus.lay,
	         FP_HL = male.pituitary.hatch - male.pituitary.lay,
	         MP_HL = male.pituitary.hatch - male.pituitary.lay,
	         MH_HL = male.hypothalamus.hatch - male.hypothalamus.lay,
	         MG_HL = male.gonad.hatch - male.gonad.lay,	         
levels=parentaldesign)

cont <- "FG_HL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female gonad lay to hatch', frame.plot=F)

cont <- "FH_HL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female hypothalamus lay to hatch', frame.plot=F)

cont <- "FP_HL"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
#kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)
plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female pituitary lay to hatch', frame.plot=F)

```