### 2019-04-05 analysis/04_DESeq2_gon.Rmd
hyp and gonad

---
title: "04_DESeq_gon"
output: md_document
---

```{r setup}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/gon/',cache=TRUE)
```

This anlaysis will _exclude_ the control timepoint but _combine_ incubation and nestling timepoints.

```{r read}
# import "colData" which contains sample information and "countData" which contains read counts
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1) 
```

```{r colDataWrangle}
# making new groups
colData$group <- NULL
colData$tempgroup <- ifelse(colData$treatment == "bldg", "bldg",
                  ifelse(colData$treatment == "control", "control",
                   ifelse(colData$treatment == "hatch", "hatch",
                    ifelse(grepl("inc", colData$treatment), "inc",
                     ifelse(colData$treatment == "lay", "lay", "nestl")))))
colData$tempgroup <- as.factor(colData$tempgroup)

colData$group <- paste(colData$sex, colData$tempgroup, sep = "")
colData$group <- as.factor(colData$group)
str(colData$group)

colData$treatment <- factor(colData$treatment, levels = 
                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                                "inc.d17", "hatch", "n5", "n9"))
```

```{r filter}
#select samples of interest from colData
colData <- colData %>%
  dplyr::filter(grepl('gonad', tissue)) %>%
  #dplyr::filter(treatment != "control") %>%
  dplyr::filter(sex == "female") %>%
  droplevels()
row.names(colData) <- colData$V1

# print sample sizes
colData %>% select(group, tissue)  %>%  summary()

#select samples of interest from countData
savecols <- as.character(colData$V1) 
savecols <- as.vector(savecols) 
countData <- countData %>% dplyr::select(one_of(savecols))

# check that row and col lenghts are equal
ncol(countData) == nrow(colData)  
```

```{r deseq2}
dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ treatment )
dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
dds <- DESeq(dds) # Differential expression analysis
vsd <- vst(dds, blind=FALSE) # variance stabilized  
```

```{r res}
levels(colData$treatment)
summary(results(dds, contrast=c("treatment",'control', 'bldg'))) 
```

```{res2}
summary(results(dds, contrast=c("treatment",'control', 'lay')))     
summary(results(dds, contrast=c("treatment",'control', 'inc.d3')))  
summary(results(dds, contrast=c("treatment",'control', 'inc.d9')))   
summary(results(dds, contrast=c("treatment",'control', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'control', 'hatch')))    
summary(results(dds, contrast=c("treatment",'control', 'n5')))      
summary(results(dds, contrast=c("treatment",'control', 'n9')))

summary(results(dds, contrast=c("treatment",'bldg', 'lay')))      
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d3')))   
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d9')))   
summary(results(dds, contrast=c("treatment",'bldg', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'bldg', 'hatch')))   
summary(results(dds, contrast=c("treatment",'bldg', 'n5')))      
summary(results(dds, contrast=c("treatment",'bldg', 'n9')))

#summary(results(dds, contrast=c("treatment",'lay', 'bldg')))     
summary(results(dds, contrast=c("treatment",'lay', 'inc.d3')))    
summary(results(dds, contrast=c("treatment",'lay', 'inc.d9')))   
summary(results(dds, contrast=c("treatment",'lay', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'lay', 'hatch')))    
summary(results(dds, contrast=c("treatment",'lay', 'n5')))       
summary(results(dds, contrast=c("treatment",'lay', 'n9')))

#summary(results(dds, contrast=c("treatment",'inc.d3', 'bldg')))     
#summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))       
summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d9')))   
summary(results(dds, contrast=c("treatment",'inc.d3', 'inc.d17')))   
summary(results(dds, contrast=c("treatment",'inc.d3', 'lay')))      
summary(results(dds, contrast=c("treatment",'inc.d3', 'n5')))        
summary(results(dds, contrast=c("treatment",'inc.d3', 'n9')))

summary(results(dds, contrast=c("treatment",'inc.d9', 'inc.d17')))   
summary(results(dds, contrast=c("treatment",'inc.d9', 'lay')))      
summary(results(dds, contrast=c("treatment",'inc.d9', 'n5')))       
summary(results(dds, contrast=c("treatment",'inc.d9', 'n9')))

summary(results(dds, contrast=c("treatment",'inc.d17', 'lay')))  
summary(results(dds, contrast=c("treatment",'inc.d17', 'n5')))   
summary(results(dds, contrast=c("treatment",'inc.d17', 'n9')))

#summary(results(dds, contrast=c("treatment",'hatch', 'bldg')))      
#summary(results(dds, contrast=c("treatment",'hatch', 'lay')))       
#summary(results(dds, contrast=c("treatment",'hatch', 'inc.d3')))    
#summary(results(dds, contrast=c("treatment",'hatch', 'inc.d9')))    
summary(results(dds, contrast=c("treatment",'hatch', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'hatch', 'n5')))       
summary(results(dds, contrast=c("treatment",'hatch', 'n9')))

#summary(results(dds, contrast=c("treatment",'n5', 'bldg')))      
#summary(results(dds, contrast=c("treatment",'n5', 'lay')))       
#summary(results(dds, contrast=c("treatment",'n5', 'inc.d3')))  
#summary(results(dds, contrast=c("treatment",'n5', 'inc.d9')))   
#summary(results(dds, contrast=c("treatment",'n5', 'inc.d17')))  
summary(results(dds, contrast=c("treatment",'n5', 'hatch')))   
summary(results(dds, contrast=c("treatment",'n5', 'n9')))

#summary(results(dds, contrast=c("treatment",'n9', 'bldg')))     
#summary(results(dds, contrast=c("treatment",'n9', 'lay')))      
#summary(results(dds, contrast=c("treatment",'n9', 'inc.d3')))   
#summary(results(dds, contrast=c("treatment",'n9', 'inc.d9')))   
#summary(results(dds, contrast=c("treatment",'n9', 'inc.d17')))  
#summary(results(dds, contrast=c("treatment",'n9', 'hatch')))    
summary(results(dds, contrast=c("treatment",'n9', 'n5')))     
```

```{r totalDEGs}
levels(colData$treatment)

totalDEGs(c("treatment", 'bldg','lay'))
totalDEGs(c("treatment", 'bldg','inc.d3'))
totalDEGs(c("treatment", 'bldg','inc.d9'))
totalDEGs(c("treatment", 'bldg','inc.d17'))
totalDEGs(c("treatment", 'bldg','hatch'))
totalDEGs(c("treatment", 'bldg','n5'))
totalDEGs(c("treatment", 'bldg','n9'))

#totalDEGs(c("treatment", 'lay','bldg'))
totalDEGs(c("treatment", 'lay','inc.d3'))
totalDEGs(c("treatment", 'lay','inc.d9'))
totalDEGs(c("treatment", 'lay','inc.d17'))
totalDEGs(c("treatment", 'lay','hatch'))
totalDEGs(c("treatment", 'lay','n5'))
totalDEGs(c("treatment", 'lay','n9'))

#totalDEGs(c("treatment", 'inc.d3','bldg'))
#totalDEGs(c("treatment", 'inc.d3','lay'))
totalDEGs(c("treatment", 'inc.d3','inc.d9'))
totalDEGs(c("treatment", 'inc.d3','inc.d17'))
totalDEGs(c("treatment", 'inc.d3','hatch'))
totalDEGs(c("treatment", 'inc.d3','n5'))
totalDEGs(c("treatment", 'inc.d3','n9'))

#totalDEGs(c("treatment", 'inc.d9','bldg'))
#totalDEGs(c("treatment", 'inc.d9','lay'))
#totalDEGs(c("treatment", 'inc.d9','inc.d3'))
totalDEGs(c("treatment", 'inc.d9','inc.d17'))
totalDEGs(c("treatment", 'inc.d9','hatch'))
totalDEGs(c("treatment", 'inc.d9','n5'))
totalDEGs(c("treatment", 'inc.d9','n9'))

#totalDEGs(c("treatment", 'inc.d17','bldg'))
#totalDEGs(c("treatment", 'inc.d17','lay'))
#totalDEGs(c("treatment", 'inc.d17','inc.d3'))
#totalDEGs(c("treatment", 'inc.d17','inc.d9'))
totalDEGs(c("treatment", 'inc.d17','hatch'))
totalDEGs(c("treatment", 'inc.d17','n5'))
totalDEGs(c("treatment", 'inc.d17','n9'))

#totalDEGs(c("treatment", 'hatch','bldg'))
#totalDEGs(c("treatment", 'hatch','lay'))
#totalDEGs(c("treatment", 'hatch','inc.d3'))
#totalDEGs(c("treatment", 'hatch','inc.d9'))
totalDEGs(c("treatment", 'hatch','inc.d17'))
totalDEGs(c("treatment", 'hatch','n5'))
totalDEGs(c("treatment", 'hatch','n9'))

#totalDEGs(c("treatment", 'n5','bldg'))
#totalDEGs(c("treatment", 'n5','lay'))
#totalDEGs(c("treatment", 'n5','inc.d3'))
#totalDEGs(c("treatment", 'n5','inc.d9'))
#totalDEGs(c("treatment", 'n5','inc.d17'))
totalDEGs(c("treatment", 'n5','hatch'))
totalDEGs(c("treatment", 'n5','n9'))

#totalDEGs(c("treatment", 'n9','bldg'))
#totalDEGs(c("treatment", 'n9','lay'))
#totalDEGs(c("treatment", 'n9','inc.d3'))
#totalDEGs(c("treatment", 'n9','inc.d9'))
#totalDEGs(c("treatment", 'n9','inc.d17'))
#totalDEGs(c("treatment", 'n9','hatch'))
totalDEGs(c("treatment", 'n9','n5'))
```

```{r PCA, message=FALSE, warning=FALSE}
levels(colData$treatment)

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

ggplot(pcadata, aes(PC1, PC2,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25)

ggplot(pcadata, aes(PC2, PC3,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC2: ", percentVar[2],"% variance")) +
  ylab(paste0("PC3: ", percentVar[3],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25)

ggplot(pcadata, aes(PC3, PC4,color = treatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25)

plotPCA(vsd, intgroup=c("treatment"))
```

PCA statistics

```{r pcastats}
summary(aov(PC1 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC2 ~ treatment, data=pcadata)) 
TukeyHSD(aov(PC2 ~ treatment, data=pcadata), which = "treatment")

summary(aov(PC3 ~ treatment, data=pcadata))

summary(aov(PC4 ~ treatment, data=pcadata))

summary(aov(PC5 ~ treatment, data=pcadata))

summary(aov(PC6 ~ treatment, data=pcadata)) 
```

```{r heatmap}
# see http://bioconductor.org/packages/devel/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#heatmap-of-the-count-matrix
sampleDists <- dist(t(assay(vsd)))

sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- NULL
colnames(sampleDistMatrix) <- colData$treatment
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors,
         fontsize = 6)
```