### 2019-04-09 analysis/06_DESeq2_manipulate_groups.Rmd
diff groups

---
title: "06_DESeq_manipulation_groups"
output: md_document
---

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/manipulation_groups/',cache=TRUE)

```

# Manipulation data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
m.colData <- read.csv("../results/00_colData_manipluation.csv", header = T, row.names = 1)
m.countData <- read.csv("../results/00_countData_manipluation.csv", header = T, row.names = 1)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)

# set levels
m.colData$treatment <- factor(m.colData$treatment, levels = 
                              c("m.inc.d3", "m.inc.d8", "m.inc.d9",
                                "m.inc.d17",  "prolong", "extend", "m.n2"))

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))

m.colData$newtreatment <- ifelse(grepl("m.inc.d3", m.colData$treatment), "remove.eggs",
                          ifelse(grepl("m.inc.d9", m.colData$treatment), "remove.eggs",
                          ifelse(grepl("m.inc.d17", m.colData$treatment), "remove.eggs",
                          ifelse(grepl("m.inc.d8", m.colData$treatment), "add.hatchlings",
                          ifelse(grepl("extend", m.colData$treatment), "add.hatchlings",
                          ifelse(grepl("prolong", m.colData$treatment), "keep.eggs",
                          ifelse(grepl("m.n2", m.colData$treatment), "remove.chicks",
                                   NA)))))))

m.colData$newtreatment <- factor(m.colData$newtreatment, levels = 
                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))

```

# Write for loop to do this one for every tissue and for every treatment

```{r althings}
for (eachgroup in levels(m.colData$sextissue)){

  print(eachgroup)

  colData <- m.colData %>%
      dplyr::filter(sextissue == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- m.countData %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ newtreatment )
  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
  dds <- DESeq(dds) # Differential expression analysis
  vsd <- vst(dds, blind=FALSE) # variance stabilized

#create list of groups
a <- levels(colData$newtreatment)
b <- levels(colData$newtreatment)

numDEGs <- function(group1, group2){
  res <- results(dds, contrast = c("newtreatment", group1, group2), independentFiltering = T)
  sumpadj <- sum(res$padj < 0.1, na.rm = TRUE)
  return(sumpadj)
}

# comapre all contrasts, save to datafrmes
dat=data.frame()
for (i in a){
  for (j in b){
    if (i != j) {
      k <- paste(i,j, sep = "") #assigns usique rownames
      dat[k,1]<-i               
      dat[k,2]<-j
      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
    }
  }
}

head(dat)

# widen data to create table of degs
rownames(dat) <- NULL #remove row names
data_wide <- spread(dat, V2, V3)
print(data_wide)

dat$V1 <- factor(dat$V1, levels = 
                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))
dat$V2 <- factor(dat$V2, levels = 
                              c("remove.eggs", "keep.eggs", "add.hatchlings", "remove.chicks"))

allcontrasts <- dat %>%
  ggplot( aes(V1, V2)) +
    geom_tile(aes(fill = V3)) +
    scale_fill_viridis(na.value="#FFFFFF00", 
                     limits = c(0, 6000),
                     breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000)) + 
    xlab(" ") + ylab("Timepoint") +
    labs(fill = "# of DEGs",
         subtitle = eachgroup)
plot(allcontrasts)

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("newtreatment"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

pca12 <- ggplot(pcadata, aes(PC1, PC2,color = newtreatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca12)

print(summary(aov(PC1 ~ newtreatment, data=pcadata)))
print(TukeyHSD(aov(PC1 ~ newtreatment, data=pcadata), which = "newtreatment"))

print(summary(aov(PC2 ~ newtreatment, data=pcadata))) 
print(TukeyHSD(aov(PC2 ~ newtreatment, data=pcadata), which = "newtreatment"))

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = newtreatment)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca34)
}
```