### 2019-04-19 analysis/99_translate.Rmd
first pass at translation

---
title: "translation-es"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/espanol/')
```

## Tranlsating the sample information for downstreatm analysis and data viz

```{r cars}
library(tidyverse)
library(cowplot)
samples <- read.csv("../metadata/samples.csv" , row.names = 1) 
str(samples)

muestras <- samples
names(muestras) <- c("nombrearchivo", "pajaro", "sexo", "tejido", "tratamiento", "grupo", "estudio")
str(muestras)

levels(muestras$sexo) 
levels(muestras$tejido)
levels(muestras$tratamiento)
levels(muestras$estudio)

levels(muestras$sexo) <- c("mujer", "hombre")
levels(muestras$tejido) <- c("gonada", "hipotalamo", "pituitaria")
levels(muestras$tratamiento) <- c("construir", "control", "extender", "escotilla", "incubación 17", "incubacion 03", "incubacion 09", "poner", "eliminar 17", "eliminar 03"," precoz 09"," eliminar 09"," eliminar 21"," extender n5"," extender n9"," prolongar")
levels(muestras$estudio) <- c("caracterizacion", "manipulacion")
```

```{r datasummary}
theme_rmh <- function(){ 
    theme_bw(base_size=10) +
        theme(
            panel.grid.minor.x  = element_blank(),
           panel.grid.minor.y  = element_blank(),
            strip.background = element_rect(colour="white", fill="white"),
              legend.position = "bottom",           
           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
           legend.key.size = unit(0.5, "cm"))
}

a <- muestras %>%
  #filter(estudio == "caracterizacion")  %>%
ggplot(aes(x = tratamiento, fill = sexo)) +  
  geom_bar() + 
  facet_wrap(~ tejido) +
  theme_rmh() +
  labs(x = "Hitos del cuidado parental (natural y manipulado).",
       y = "número de muestras",
       subtitle = "Muestras de la HPG de padres \nde paloma de roca masculinos y femeninos") + 
  coord_flip()

b <- samples %>%
ggplot(aes(x = treatment, fill = sex)) +  
  geom_bar() + 
  facet_wrap(~ tissue) +
  theme_rmh() +
  labs(x = "Milestones of parental care (natural and manipualted",
       y = "Number of samples",
       subtitle = "Samples from the HPG axis of \nmale and female rock dove parents") + 
  coord_flip()

plot_grid(b,a, labels = c("english", "espanñol"), label_size=7)
```