### 2019-04-23 analysis/01_pca_all.Rmd
pca

---
title: "01_pca"
output: md_document
---

```{r setup}
library(ggfortify) # for PCA analysis
library(cluster)

knitr::opts_chunk$set(echo = TRUE, cache = T, fig.path = '../figures/pca/')
```

```{r rawdata}
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

```{r pca}
# https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html

# pca
pca <- prcomp(t(countData))
autoplot(pca, data = colData, colour = 'tissue') 
autoplot(pca, data = colData, colour = 'sex') 
autoplot(pca, data = colData, colour = 'treatment') 
autoplot(pca, data = colData, colour = 'tissue', shape = "sex") 
autoplot(pca, data = colData, colour = 'tissue', shape = "sex", alpha = "treatment")
autoplot(pca, data = colData, colour = 'treatment', shape = "sex", alpha = "tissue")

# kmeans
autoplot(clara(t(countData),3), frame = TRUE)

# probability ellipse
autoplot(pam(t(countData), 3), frame = TRUE, frame.type = 'norm')
```