### 2019-04-24 analysis/01_pca_all.Rmd
more work on spanish translations

---
title: "01_pca"
output: md_document
---

```{r setup}
library(tidyverse)
library(ggfortify) # for PCA analysis
library(cluster)
library(cowplot)
library(stringr)

theme_rmh <- function(){ 
    theme_bw(base_size=10) +
        theme(
            panel.grid.minor.x  = element_blank(),
           panel.grid.minor.y  = element_blank(),
            strip.background = element_rect(colour="white", fill="white"),
              legend.position = "right",           
           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
           legend.key.size = unit(0.5, "cm"))
}

knitr::opts_chunk$set(echo = TRUE, cache = T, fig.path = '../figures/pca/')
```

```{r rawdata}
counts <- read.csv("../results/00_countData_characterization.csv", row.names = 1, header = T)

samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)
geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)

head(samples)
```

```{r pca}
# https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html

# pca
pca <- prcomp(t(counts))

autoplot(pca, data = samples, colour = 'tissue') 
autoplot(pca, data = samples, colour = 'sex') 
autoplot(pca, data = samples, colour = 'treatment') 
autoplot(pca, data = samples, colour = 'tissue', shape = "sex") 
autoplot(pca, data = samples, colour = 'tissue', shape = "sex", alpha = "treatment")
autoplot(pca, data = samples, colour = 'treatment', shape = "sex", alpha = "tissue")

# kmeans
autoplot(clara(t(counts),3), frame = TRUE)

# probability ellipse
autoplot(pam(t(counts), 3), frame = TRUE, frame.type = 'norm')

head(pca)
```

```{r pca-es}
p <- autoplot(pca, data = samples, colour = 'tissue', shape = "sex")

en <- p + labs(subtitle = "Principle Component Analysis") +
  scale_colour_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
  xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
  ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
  theme_rmh() + guides(colour = guide_legend(order = 2), 
              shape = guide_legend(order = 1))

es <- p + labs(subtitle = "Análisis de Componentes Principales ") +
  scale_colour_discrete(name = "tejido", labels = c("hipotálamo", "pituitaria", "gónada")) +
  scale_shape_discrete(name = "sexo", labels = c("femenino", "masculino")) +
  xlab(stringr::str_replace(p$labels$x, "PC", "Componente Principal ")) +
  ylab(stringr::str_replace(p$labels$y, "PC", "Componente Principal ")) +
  theme_rmh() + guides(colour = guide_legend(order = 2), 
              shape = guide_legend(order = 1))

enes <- plot_grid(en,es)

enes

ggsave("../figures/espanol/pca-en-es.png", width=6, height=3, dpi=300)

pdf(file="../figures/espanol/pca-en-es.pdf", width=6, height=3)
plot(enes)    
dev.off()
```

```{r subset}
for (eachgroup in levels(samples$treatment)){

  print(eachgroup)

  colData <- samples %>%
      dplyr::filter(treatment == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- counts %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  pca <- prcomp(t(countData))

  p <- autoplot(pca, data = colData, colour = 'tissue', shape = "sex")

en <- p + labs(subtitle = eachgroup) +
  theme_rmh()

plot(en) 
}
```