### 2019-04-25 analysis/01_pca_all.Rmd

---
title: "01_pca"
output: md_document
---

```{r setup}
library(tidyverse)
library(ggfortify) # for PCA analysis
library(cluster)
library(cowplot)
library(stringr)
library(gridExtra)
library(grid)

theme_rmh <- function(){ 
    theme_bw(base_size=14) +
        theme(
            panel.grid.minor.x  = element_blank(),
           panel.grid.minor.y  = element_blank(),
            strip.background = element_rect(colour="white", fill="white"),
              legend.position = "right",           
           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
           legend.key.size = unit(0.5, "cm"))
}

treatmentcolors <-  c("control" = "#bdbdbd", 
                              "bldg" =  "#d95f02", 
                              "lay"  = "#a1d99b", 
                              "inc.d3" = "#74c476", 
                              "inc.d9"  = "#41ab5d", 
                              "inc.d17" = "#238b45", 
                              "hatch"  = "#807dba", 
                              "n5"  = "#6a51a3", 
                               "n9" = "#4a1486")

knitr::opts_chunk$set(echo = TRUE, cache = T, fig.path = '../figures/pca/')
```

```{r rawdata}
counts <- read.csv("../results/00_countData_characterization.csv", row.names = 1, header = T)

samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)

samples$treatment <- factor(samples$treatment, levels = 
                              c("control", "bldg", "lay", "inc.d3", "inc.d9", 
                                "inc.d17", "hatch", "n5", "n9"))
samples$tissue <- factor(samples$tissue, levels = 
                              c("hypothalamus", "pituitary", "gonad"))

geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)
```

```{r raw}
# https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html

# pca
pca <- prcomp(t(counts))

p <- autoplot(pca, data = samples, alpha = 'tissue', shape = "sex")

rawPCA112 <- p + labs(subtitle = "Raw counts") +
  scale_colour_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
  #xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
  #ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
  theme_rmh() + guides(alpha = guide_legend(order = 1), 
              shape = guide_legend(order = 2)) +
  stat_ellipse(aes(linetype = sex, colour = tissue)) 
rawPCA112

```

## with normalize edgeR data

```{r normalized}

pseudocounts <- read.csv("../results/01_pseudo.counts.csv", row.names = 1)
head(pseudocounts)
pca <- prcomp(t(pseudocounts))

p <- autoplot(pca, data = samples, alpha = 'tissue', shape = "sex")
p

edgRpca12 <- p + labs(subtitle = "Normalized counts from edgeR") +
  scale_alpha_discrete(name = "tissue", labels = c("hypothalamus", "pituitary", "gonad")) +
  scale_shape_discrete(name = "sex", labels = c("female", "male")) +
  #xlab(stringr::str_replace(p$labels$x, "PC", "Principle Component ")) +
  #ylab(stringr::str_replace(p$labels$y, "PC", "Principle Component ")) +
  theme_rmh() + guides(alpha = guide_legend(order = 1), 
              shape = guide_legend(order = 2)) +
  stat_ellipse(aes(linetype = sex, colour = tissue))
edgRpca12
```

```{r sidebyside}

sidebyside <- plot_grid(rawPCA112 + theme(legend.position = "none"),
                        edgRpca12,
                        rel_widths = c(0.4,0.6))
sidebyside

```

## subsets of normalized data show by tissue and sex

```{r tissuesex}
plotlist = list()
for (eachgroup in levels(samples$treatment)){

  print(eachgroup)

  colData <- samples %>%
      dplyr::filter(treatment == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- pseudocounts %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  pca <- prcomp(t(countData))

  p <- autoplot(pca, data = colData, alpha = 'tissue', shape = "sex", color = "treatment")

en <- p + labs(subtitle = eachgroup) +
  theme_rmh() +
    theme_rmh()  +
  stat_ellipse(aes(linetype = sex, alpha = tissue, color = treatment)) +
   theme_bw(base_size=8) +
    theme(legend.position = "bottom") +
  scale_color_manual(values = treatmentcolors) + 
  guides(color = FALSE)

legend <- get_legend(en)

en <- en  + theme(legend.position = "none")

plotlist[[eachgroup]] = en

}

main <- grid.arrange(grobs=plotlist,ncol=3)

withlegend <- plot_grid( main, legend, rel_heights  = c(3, .3), ncol = 1)
withlegend

```

## sex and timepoint within a tissue

```{r timepointsex}
plotlist = list()
for (eachtissue in levels(samples$tissue)){

  print(eachtissue)

  colData <- samples %>%
      dplyr::filter(tissue == eachtissue) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- pseudocounts %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  pca <- prcomp(t(countData))

  p <- autoplot(pca, data = colData, colour = 'treatment', shape = "sex")

en <- p + labs(subtitle = eachtissue) +
  theme_rmh() +  guides(colour = guide_legend(order = 1), 
              shape = guide_legend(order = 2)) +
  stat_ellipse(aes(linetype = sex, colour = treatment)) +
  theme_bw(base_size=8) +
    theme(legend.position = "bottom") +
  scale_color_manual(values =  treatmentcolors)

legend <- get_legend(en)

en <- en  + theme(legend.position = "none")

plotlist[[eachtissue]] = en

#plot(en) 
}

main <- grid.arrange(grobs=plotlist,ncol=3)

withlegend <- plot_grid( main, legend, rel_heights  = c(3, .3), ncol = 1)
withlegend

```