### 2019-05-19 analysis/05_DESeq2_manipulate_p2.Rmd
cleaning up, prep for separarate functions

## 2nd for loop on data clustered into early or late termination

```{r allthings-outcome, eval = T}

for (eachgroup in levels(m.colData$sextissue)){

  print(eachgroup)

  colData <- m.colData %>%
      dplyr::filter(sextissue == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- m.countData %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData,
                              design = ~ outcome )
  dds <- dds[ rowSums(counts(dds)) > 2, ] ## pre-filter genes 
  dds <- DESeq(dds) # Differential expression analysis
  vsd <- vst(dds, blind=FALSE) # variance stabilized 
head(vsd)

numDEGs <- function(group1, group2){
  res <- results(dds, contrast = c("outcome", group1, group2), independentFiltering = T)
  sumpadj <- sum(res$padj < 0.1, na.rm = TRUE)
  return(sumpadj)}

#create list of groups
a <- levels(colData$outcome)
b <- levels(colData$outcome)

# comapre all contrasts, save to datafrmes
dat=data.frame()
for (i in a){
  for (j in b){
    if (i != j) {
      k <- paste(i,j, sep = "") #assigns usique rownames
      dat[k,1]<-i               
      dat[k,2]<-j
      dat[k,3]<- numDEGs(i,j) #caluculates number of DEGs
    }
  }
}

head(dat)

# widen data to create table of degs
rownames(dat) <- NULL #remove row names
data_wide <- spread(dat, V2, V3)
print(data_wide)

dat$V1 <- factor(dat$V1, levels = 
                              c("end inc",  "end hatch",
                                "prolong inc", "delay hatch"))
dat$V2 <- factor(dat$V2, levels = 
                              c("end inc",  "end hatch",
                                "prolong inc", "delay hatch"))

allcontrasts <- dat %>%
  ggplot( aes(V1, V2)) +
    geom_tile(aes(fill = V3)) +
    scale_fill_viridis(na.value="#FFFFFF00", 
                     limits = c(0, 6000),
                     breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000)) + 
    xlab(" ") + ylab("Timepoint") +
    labs(fill = "# of DEGs",
         subtitle = eachgroup)
plot(allcontrasts)

# create the dataframe using my function pcadataframe
pcadata <- pcadataframe(vsd, intgroup=c("outcome"), returnData=TRUE)
percentVar <- round(100 * attr(pcadata, "percentVar"))
percentVar

pca12 <- ggplot(pcadata, aes(PC1, PC2,color = outcome)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC1: ", percentVar[1],"% variance")) +
  ylab(paste0("PC2: ", percentVar[2],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
print(pca12)

print(summary(aov(PC1 ~ outcome, data=pcadata)))
print(TukeyHSD(aov(PC1 ~ outcome, data=pcadata), which = "outcome"))

print(summary(aov(PC2 ~ outcome, data=pcadata))) 
print(TukeyHSD(aov(PC2 ~ outcome, data=pcadata), which = "outcome"))

pca34 <- ggplot(pcadata, aes(PC3, PC4,color = outcome)) + 
  geom_point(size = 2, alpha = 1) +
  stat_ellipse(type = "t") +
  xlab(paste0("PC3: ", percentVar[3],"% variance")) +
  ylab(paste0("PC4: ", percentVar[4],"% variance")) +
  theme_cowplot(font_size = 8, line_size = 0.25) +
  labs(subtitle = eachgroup)
#print(pca34)

}
```