### 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
half way there

---
title: "05_DESeq_manipulation"
output: md_document
---

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/manipulation/', cache = TRUE)
```

## Manipulation data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
m.colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)
m.countData <- read.csv("../results/00_countData_manipluation.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
m.colData$treatment <- factor(m.colData$treatment, levels = 
                              c("m.inc.d3",  "m.inc.d8",
                                "m.inc.d9", "m.inc.d17",
                                "prolong", "extend", "m.n2"))

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))

m.colData$outcome <- ifelse(grepl("d3|d9|d17", m.colData$treatment), "end inc", 
                     ifelse(grepl("d8|n2", m.colData$treatment),"end hatch",
                     ifelse(grepl("prolong", m.colData$treatment),"prolong inc",
                     ifelse(grepl("extend", m.colData$treatment),"delay hatch", NA))))

m.colData$outcome <- factor(m.colData$outcome, levels = 
                              c("end inc",  "end hatch",
                                "prolong inc", "delay hatch"))
summary(m.colData[c(7,3,4,5,8,9)])
```

## Run DESeq on all subsets of the data

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq("female_hypothalamus")
dds.female_pituitary <- subsetDESeq("female_pituitary" )
dds.female_gonad <- subsetDESeq("female_gonad" )
dds.male_hypothalamus <- subsetDESeq("male_hypothalamus" )
dds.male_pituitary <- subsetDESeq("male_pituitary"  )
dds.male_gondad <- subsetDESeq("male_gonad")
```

## Calculate and plot total DEGs

```{r totalDEGs, eval = F}
#create list of groups for deseq contrasts
group1 <- c("m.inc.d3","m.inc.d8", "m.inc.d9", "m.inc.d17", "prolong","extend", "m.n2" ) 
group2 <- group1

a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
c <- plottotalDEGs(dds.female_gonad, "female gonad")
d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
f <- plottotalDEGs(dds.male_gondad, "male gonad")

plot_grid(a,b,c,d,e,f, nrow = 2)
```

## Calculate and plot principal components

```{r pca}

#mydds <- list(ddsmale_hypothalamus) # for testing only
mydds <- list(dds.female_hypothalamus, dds.female_pituitary, dds.female_gonad,
             dds.male_hypothalamus, dds.male_pituitary, dds.male_gondad)

for (eachdds in mydds){

  vsd <- vst(eachdds, blind=FALSE) # variance stabilized

  # create the dataframe using my function pcadataframe
  pcadata <- pcadataframe(vsd, intgroup=c("treatment"), returnData=TRUE)
  percentVar <- round(100 * attr(pcadata, "percentVar"))
  print(percentVar)

  pca1 <- ggplot(pcadata, aes(treatment, PC1,color = treatment)) + 
    geom_boxplot() +
    ylab(paste0("PC1: ", percentVar[1],"% variance")) +
    xlab(NULL) +
    theme_cowplot(font_size = 8, line_size = 0.25) +
    labs(subtitle = "eachgroup") +
    theme(legend.position = "none")

  pca2 <- ggplot(pcadata, aes(treatment, PC2,color = treatment)) + 
    geom_boxplot() +
    ylab(paste0("PC2: ", percentVar[2],"% variance")) +
    xlab(NULL) +
    theme_cowplot(font_size = 8, line_size = 0.25) +
    labs(subtitle = "eachgroup") +
    theme(legend.position = "none")

  mypca <- plot_grid(pca1, pca2)
  plot(mypca)

  print(summary(aov(PC1 ~ treatment, data=pcadata)))
  print(TukeyHSD(aov(PC1 ~ treatment, data=pcadata), which = "treatment"))

  print(summary(aov(PC2 ~ treatment, data=pcadata))) 
  print(summary(aov(PC3 ~ treatment, data=pcadata))) 
  print(summary(aov(PC4 ~ treatment, data=pcadata)))

}
```

## heamap with minimum pvalue

```{r pheatmap, eval = F}

for (eachdds in mydds){

  vsd <- vst(eachdds, blind=FALSE) # variance stabilized

  # make dataframe counts
  DEGs <- assay(vsd)
  DEGs <- as.data.frame(DEGs)

  a <- levels(m.colData$treatment)
  b <- a

  for (i in a){
    for (j in b){
      if (i != j) {
        print(i)
        results <- returnpadj(i,j)
        DEGs <- cbind(DEGs,results)
      }
    }
     b <- b[-1]  # drop 1st element of second string to not recalculate DEGs
  }

  DEGsmatrix <- DEGs
  DEGsmatrix <- as.matrix(DEGs)
  padjmin <- rowMins(DEGsmatrix, na.rm = T) 
  padjmin <- as.data.frame(padjmin)

  sigDEGs <- cbind(DEGs,padjmin)

  sigDEGs <- sigDEGs %>% arrange(padjmin)
  sigDEGs <- head(sigDEGs,500)
  sigDEGs <- as.data.frame(sigDEGs)
  rownames(sigDEGs) <- sigDEGs$rownames
  drop.cols <-colnames(sigDEGs[,grep("padj|pval|pmin|rownames", colnames(sigDEGs))])
  sigDEGs <- sigDEGs %>% dplyr::select(-one_of(drop.cols))
  sigDEGs <- as.matrix(sigDEGs)
  sigDEGs <- sigDEGs - rowMeans(sigDEGs)

  paletteLength <- 30
  myBreaks <- c(seq(min(sigDEGs), 0, length.out=ceiling(paletteLength/2) + 1), 
                seq(max(sigDEGs)/paletteLength, max(sigDEGs), length.out=floor(paletteLength/2)))

  anndf <- m.colData %>% dplyr::select(treatment)
  rownames(anndf) <- m.colData$V1

  sigDEGs <- as.matrix(sigDEGs) 
  pheatmap(sigDEGs, show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf,
         main = "eachgroup")

  pheatmap(sigDEGs, kmeans_k = 5,
         show_rownames = F, show_colnames = F,
         color = viridis(30),
         breaks=myBreaks,
         annotation_col=anndf,
         main = "eachgroup")
}

```

```{r candidates}
for (eachdds in mydds){

  vsd <- vst(eachdds, blind=FALSE) # variance stabilized

  # make dataframe counts
  DEGs <- assay(vsd)
  DEGs <- as.data.frame(DEGs)

names(geneinfo)
names(geneinfo)[4] <- "rownames"
DEGs$rownames <- row.names(DEGs)

# make dataframe with geneids and names and counts
# how to gather: https://tidyr.tidyverse.org/reference/gather.html

candidates <- full_join(geneinfo, DEGs)
drop.cols <-colnames(candidates[,grep("padj|pval|pmin", colnames(candidates))])
candidates <- candidates %>% dplyr::select(-one_of(drop.cols))
candidates <- candidates %>%
  filter(Name %in% c( #"JUN", "JUND", "EGR",  "AVP", "AVPR1A", "AVPR1B", "AVPR2", "OXT",  
                     "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
                     "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
                     "PGR", "PRL", "PRLR", "VIP", "VIPR1")) 
row.names(candidates) <- candidates$Name
candidates <- candidates %>% select(-row.names, -rownames, -Name, -geneid)
candidates <- candidates %>% drop_na()
candidates <- as.data.frame(t(candidates))
candidates$RNAseqID <- rownames(candidates)
candidates <- candidates %>% gather(gene, value, -RNAseqID)  %>% 
  filter(RNAseqID != "gene")
candidates$value <- as.numeric(candidates$value)
candidates$V1  <- candidates$RNAseqID

candidatecounts <- left_join(candidates, m.colData)
candidatecounts$faketime <- as.numeric(candidatecounts$treatment)
candidatecounts$gene <- as.factor(candidatecounts$gene)

p1 <- candidatecounts %>%
  filter(gene %in% c( "AR", "CYP19A1", "ESR1", "ESR2", "FSHR",
                      "GHRL", "GAL", "NPVF", "GNRH1", "LHCGR",
                      "PGR", "PRL", "PRLR", "VIP", "VIPR1")) %>%
  ggplot(aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot() +
  facet_wrap(~gene, scales = "free") +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom") +
  labs(x = NULL) +
  guides(fill= guide_legend(nrow=1))

plot(p1)
}
```