### 2019-05-20 analysis/05_DESeq2_manipulate.Rmd
now with real subtitles

---
title: "05_DESeq_manipulation"
output: md_document
---

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/manipulation/', cache = TRUE)
```

## Manipulation data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
m.colData <- read.csv("../metadata/00_colData_manipluation.csv", header = T, row.names = 1)
m.countData <- read.csv("../results/00_countData_manipluation.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
m.colData$treatment <- factor(m.colData$treatment, levels = 
                              c("m.inc.d3",  "m.inc.d8",
                                "m.inc.d9", "m.inc.d17",
                                "prolong", "extend", "m.n2"))

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))

m.colData$outcome <- ifelse(grepl("d3|d9|d17", m.colData$treatment), "end inc", 
                     ifelse(grepl("d8|n2", m.colData$treatment),"end hatch",
                     ifelse(grepl("prolong", m.colData$treatment),"prolong inc",
                     ifelse(grepl("extend", m.colData$treatment),"delay hatch", NA))))

summary(m.colData[c(7,3,4,5,8)])
```

## Run DESeq on all subsets of the data

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq("female_hypothalamus")
dds.female_pituitary <- subsetDESeq("female_pituitary" )
dds.female_gonad <- subsetDESeq("female_gonad" )
dds.male_hypothalamus <- subsetDESeq("male_hypothalamus" )
dds.male_pituitary <- subsetDESeq("male_pituitary"  )
dds.male_gondad <- subsetDESeq("male_gonad")
```

## Calculate and plot total DEGs

```{r totalDEGs, eval = T}
#create list of groups for deseq contrasts
group1 <- c("m.inc.d3","m.inc.d8", "m.inc.d9", "m.inc.d17", "prolong","extend", "m.n2" ) 
group2 <- group1

a <- plottotalDEGs(dds.female_hypothalamus, "female hypothalamus")
b <- plottotalDEGs(dds.female_pituitary, "female pituitary")
c <- plottotalDEGs(dds.female_gonad, "female gonad")
d <- plottotalDEGs(dds.male_hypothalamus, "male hypothalamus")
e <- plottotalDEGs(dds.male_pituitary, "male pituitary")
f <- plottotalDEGs(dds.male_gondad, "male gonad")

plot_grid(a,b,c,d,e,f, nrow = 2) 
```

```{r totalDEGs2}
plot_grid(a + theme_minimal(base_size = 8) + theme(legend.position = "none", axis.text.x = element_text(angle = 90)),
          b + theme_minimal(base_size = 8) + theme(legend.position = "none", axis.text.x = element_text(angle = 90)),
          c + theme_minimal(base_size = 8) + theme(axis.text.x = element_text(angle = 90)),
          d + theme_minimal(base_size = 8) + theme(legend.position = "none",axis.text.x = element_text(angle = 90)),
          e + theme_minimal(base_size = 8) + theme(legend.position = "none",axis.text.x = element_text(angle = 90)),
          f + theme_minimal(base_size = 8) + theme(axis.text.x = element_text(angle = 90)),
          nrow = 2, rel_widths = c(0.3, 0.3, 0.4)) 
```

## Calculate and plot principal components

```{r pca}
plotPCAs(dds.female_hypothalamus, "female hypothalamus")
plotPCAs(dds.female_pituitary, "female pituitary")
plotPCAs(dds.female_gonad, "female gonad")
plotPCAs(dds.male_hypothalamus, "male hypothalamus")
plotPCAs(dds.male_pituitary, "male pituitary")
plotPCAs(dds.male_gondad, "male gonad")
```

## heamap with minimum pvalue

```{r pheatmap, eval = F}
makepheatmap(dds.female_hypothalamus, "female hypothalamus")
makepheatmap(dds.female_pituitary, "female pituitary")
makepheatmap(dds.female_gonad, "female gonad")
makepheatmap(dds.male_hypothalamus, "male hypothalamus")
makepheatmap(dds.male_pituitary, "male pituitary")
makepheatmap(dds.male_gondad, "male gonad")
```

```{r candidates}
plotcandidates(dds.female_hypothalamus, "female hypothalamus")
plotcandidates(dds.female_pituitary, "female pituitary")
plotcandidates(dds.female_gonad, "female gonad")
plotcandidates(dds.male_hypothalamus, "male hypothalamus")
plotcandidates(dds.male_pituitary, "male pituitary")
plotcandidates(dds.male_gondad, "male gonad")
```