### 2019-08-13 analysis/06_musicalgenes.Rmd
move musical genes to new script

---
title: "06_musicalgenes"
output: md_document
---

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(cowplot)
library(ggimage)

# load custom functions  
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/characterization/', cache = TRUE)
```

## Characterization data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
levels(c.colData$treatment)
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
summary(c.colData[c(7,3,4,5,8)])
```

## Musical genes

```{r musicalgenes, cache = T, eval=T}
countData <- c.countData
colData <- c.colData

candidataCounts <- countData
candidataCounts$Name <- row.names(candidataCounts)
candidataCounts <- candidataCounts %>%
  filter(Name %in% c("NP_990797.2", "NP_001138861.1", "NP_001107180.1"))
row.names(candidataCounts) <- candidataCounts$Name
candidataCounts$Name <- NULL
candidataCountsT <- t(candidataCounts)
candidataCountsT <- as.data.frame(candidataCountsT)

candidataCountsT$entrez <- row.names(candidataCountsT)
colData$entrez <- row.names(colData)

candidataCountsTcol <- full_join(colData, candidataCountsT)

str(candidataCountsTcol)

candidataCountsTcol2 <- candidataCountsTcol %>%
  group_by(sextissue, treatment) %>%
  summarize(GAL = mean(NP_001138861.1, na.rm = TRUE),
            PRL = mean(NP_990797.2, na.rm = TRUE),
            CALCA = mean(NP_001107180.1, na.rm = TRUE)) %>%
  gather(GAL, PRL, CALCA, key = "gene", value = "mean")

candidataCountsTcol2 <- as.data.frame(candidataCountsTcol2)

musicgenes <- function(mygroup, mysusbtitle){

myplot <- candidataCountsTcol2 %>%
  filter(sextissue %in% c(mygroup)) %>%
ggplot(aes(x = treatment, y = mean)) +
  geom_point(colour = "white") + 
  geom_line(group = 1, color='grey') +
  theme_bw() +
  facet_wrap(~gene, scales = "free", ncol = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_image(aes(image="../figures/characterization/note-black-04.png")) +
  labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL)

  return(myplot)

}

a <- musicgenes("female_hypothalamus", "female hypothalamus")
b <- musicgenes("male_hypothalamus", "male hypothalamus")

c <- plot_grid(a,b)
c

pdf("../figures/characterization/musicalgenes-1.pdf", width = 8, height = 6)
plot(c)
dev.off()

musicgeneslines <- function(mygroup, mysusbtitle, mygene, myimage, myylim, myhline){
  myplot <- candidataCountsTcol2 %>%
  filter(sextissue %in% c(mygroup),
         gene == mygene) %>%
    ggplot(aes(x = treatment, y = mean)) +
    geom_point(colour = "white") + 
    theme_minimal() +
    #facet_wrap(~gene,  ncol = 1) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text = element_text(face = "italic"),
          panel.grid = element_blank(),
          axis.text.y = element_blank()) +
    geom_image(aes(image=myimage)) +
   labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL) +
    ylim(myylim) + 
    geom_hline(yintercept = myhline, colour="grey")
  return(myplot)
}

musicgenesnolines <- function(mygroup, mysusbtitle, mygene, myimage, myylim, myhline){
  myplot <- candidataCountsTcol2 %>%
  filter(sextissue %in% c(mygroup),
         gene == mygene) %>%
    ggplot(aes(x = treatment, y = mean)) +
    geom_point(colour = "white") + 
    theme_minimal(base_size = 10) +
    #facet_wrap(~gene,  ncol = 1) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text = element_text(face = "italic"),
          panel.grid = element_blank(),
          axis.text.y = element_blank()) +
    geom_image(aes(image=myimage)) +
   labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL) +
    ylim(myylim)

  return(myplot)
}

a <- musicgenesnolines("female_hypothalamus", "female hypothalamus", "GAL", "../figures/characterization/note-orange-04.png", c(15,75))
a

b <- musicgenesnolines("male_hypothalamus", "male hypothalamus", "GAL", "../figures/characterization/note-orange-04.png", c(15,75))
b

ab <- plot_grid(a,b) + cowplot::draw_text("GAL", x = 0.925, y = 0.9, size = 8, colour = "#d95f02")
ab

c <- musicgenesnolines("female_hypothalamus", "female hypothalamus", "CALCA", "../figures/characterization/note-teal-04.png", c(0,100))
c

d <- musicgenesnolines("male_hypothalamus", "male hypothalamus", "CALCA", "../figures/characterization/note-teal-04.png", c(0,100))
d

cd <- plot_grid(c,d)  + cowplot::draw_text("CALCA", x = 0.925, y = 0.85, size = 8, colour = "#1b9e77")
cd

e <- musicgenesnolines("female_hypothalamus", "female hypothalamus", "PRL", "../figures/characterization/note-purple-04.png", c(800,3000) )
e

f <- musicgenesnolines("male_hypothalamus", "male hypothalamus", "PRL", "../figures/characterization/note-purple-04.png", c(800,3000))
f

ef <- plot_grid(e,f) + cowplot::draw_text("PRL", x = 0.925, y = 0.8, size = 8, colour = "#7570b3")
ef

pdf("../figures/characterization/musicalgenes-ab.pdf", width = 4, height = 2)
plot(ab)
dev.off()

pdf("../figures/characterization/musicalgenes-cd.pdf", width = 4, height = 2)
plot(cd)
dev.off()

pdf("../figures/characterization/musicalgenes-ef.pdf", width = 4, height = 2)
plot(ef)
dev.off()

```