### 2019-08-26 analysis/08_wgcna.Rmd
now with hypothesis groups

---
title: "wgcna"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = '../figures/wgcna/')

source("08_WGCNA_1.R")

```

# WGCNA analysis, first pass

The anlayses and image was made by following along with and modifing the [first in a series of WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R). The first tutorial focuses on data input, cleaning and pre-processing; outlier removal; clustering of samples by their Euclidean distance; a comparison of sample cluster to sample meta data. I put pretty much the whole script into a single function that I source with [`08_WGCNA_1.R`](https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R).

For the following figures, the top panel is a dendrogram showing Euclidean distance of sample. Because the labeling of samples is hard to read, the colored rows can be used to visualize traits. Here, both sex and tissues are colored in a white to red scaled with the following pattern:

* sex: females (white), males (red), 
* tissue: gonad (white), hypothalamus (pink), pituitary (red).

## All characterization data

```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE}

subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
                           "male.gonad", "male.hypothalamus", "male.pituitary"),
            mytitle = "Characterization dendrogram with trait heatmap")

```

**Fig 1. Characterization dendrogram with trait heatmap.**. 
 Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.

## Females

```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE}
subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary"),
            mytitle = "Female sample dendrogram with trait heatmap")

```

**Fig 2. Female characterization dendrogram with trait heatmap.**. 
Most pituitary samples form a single cluster. More hypothalamus samples cluster with gonads than with pituitary.

## Males

```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE}
subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"),
            mytitle = "Male sample dendrogram with trait heatmap")

```

**Fig 3. Male characterization dendrogram with trait heatmap.**. 
 As with females, most pituitary samples form a single cluster, while more hypothalamus samples cluster with gonads than with pituitary.

## gonads

```{r wgcna-gon, message=FALSE, warning=FALSE, cache=TRUE}
subsetWGCNA(c("female.gonad", "male.gonad"),
            mytitle = "Gonad dendrogram with trait heatmap")
```

**Fig 3. Gonad characterization dendrogram with trait heatmap.**. 
Male and female samples cluster together, but with large, sex-specific regions.... maybe there is a strong treatment effect here.

## hypothalamus

```{r wgcna-hyp, message=FALSE, warning=FALSE, cache=TRUE}
subsetWGCNA(c("male.hypothalamus", "female.hypothalamus"),
            mytitle = "Hypothalamus dendrogram with trait heatmap")
```

Male and female samples cluster together, but with very small, sex-specific regions.... maybe there is a small treatment effect here.

## pituitary (not working)

```{r wgcna-pit, message=FALSE, warning=FALSE, cache=TRUE}

#subsetWGCNA(c("female.pitutary", "male.pitutary"),
#            mytitle = "Pituitary dendrogram with trait heatmap")
```

## female hypothalamus

```{r wgcna-f-hyp, message=FALSE, warning=FALSE, cache=TRUE}
subsetWGCNA(c("female.hypothalamus"),
            mytitle = "Female hypothalamus sample dendrogram")
```

## female gonad

```{r wgcna-f-gon, message=FALSE, warning=FALSE, cache=TRUE}
subsetWGCNA(c("female.gonad"),
            mytitle = "Female gonad sample dendrogram")
```