### 2019-08-27 analysis/09_egdeR.Rmd
now with go and kegg

---
title: "edgeR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(edgeR)
library(org.Gg.eg.db)
library(dplyr)
```

# edgeR: differential expression analysis of digital gene expression data

https://www.bioconductor.org/packages/release/bioc/vignettes/edgeR/inst/doc/edgeRUsersGuide.pdf

## quick start

```{r}
x <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
group <- colData$treatment

# to make go work, I need to swap out row names
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)
x$entrezid <- row.names(x)
x <- left_join(x, geneinfo)
x <- subset(x, !duplicated(subset(x, select=c(row.names))))
row.names(x) <- x$row.names
x <- x[ -c(577:580) ]

y <- DGEList(counts=x,group=group)
y <- calcNormFactors(y)
design <- model.matrix(~group)
y <- estimateDisp(y,design)

# To perform quasi-likelihood F-tests:
fit <- glmQLFit(y,design)
qlf <- glmQLFTest(fit,coef=2)
as.data.frame(topTags(qlf))
```

## Gene ontology (GO) and pathway analysis

```{r}
go <- goana(qlf, species="Gg", geneid = rownames(x))
topGO(go, sort="up")

keg <- kegga(qlf, species="Gg")
topKEGG(keg, sort="up")
```