### 2019-09-03 analysis/08_wgcna.Rmd
only looking at pituitary

---
title: "wgcna"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = '../figures/wgcna/')

source("08_WGCNA_1.R")

```

# WGCNA analysis, first pass

The anlayses and image was made by following along with and modifing the [first in a series of WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R). The first tutorial focuses on data input, cleaning and pre-processing; outlier removal; clustering of samples by their Euclidean distance; a comparison of sample cluster to sample meta data. I put pretty much the whole script into a single function that I source with [`08_WGCNA_1.R`](https://github.com/macmanes-lab/DoveParentsRNAseq/blob/master/analysis/08_WGCNA_1.R).

For the following figures, the top panel is a dendrogram showing Euclidean distance of sample. Because the labeling of samples is hard to read, the colored rows can be used to visualize traits. Here, both sex and tissues are colored in a white to red scaled with the following pattern:

* sex: females (white), males (red), 
* tissue: gonad (white), hypothalamus (pink), pituitary (red).

## All characterization data

```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE}

subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
                           "male.gonad", "male.hypothalamus", "male.pituitary"),
            mytitle = "Characterization dendrogram with trait heatmap")

```

**Fig 1. Characterization dendrogram with trait heatmap.**. 
 Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.

## pituitary (not working)

```{r wgcna-pit, message=FALSE, warning=FALSE, cache=TRUE}

subsetWGCNA(c("female.pitutary", "male.pitutary"),
            mytitle = "Pituitary dendrogram with trait heatmap")
```