### 2019-09-07 docs/characterization_manuscript.Rmd
new figs

---
title: Peaks and valleys of prolatin-related gene expression during pigeon parental care stages
author:
  - name: Rayna M Harris
    email: rmharris@ucdavis
    affiliation: UC Davis
    corresponding: "rmharris@ucdavis"
  - name: Suzanne Austin
    affiliation: UC Davis
  - name: Andrew Lang
    affiliation: UNH
  - name: Matthew MacManes
    affiliation: UNH
  - name: Rebecca M Calisi
    affiliation: UC Davis           
address:
  - code: UC Davis
    address: UC Davis, Davis, CA
  - code: UNH
    address: UNH, smalltown, NH
abstract: |
  The goal of this research was to provide the most comprehensive gene expression profile of gene expression activity in the HPG axis of male and female rock doves to provide a deeper undertand of how the reproductive axis response to typical behavioral transitions associated with parental care. Here we used RNA sequencing to measure gene activity at 8 stages of the parental care cyle from nest building, to egg incubation, and through nestling care. Non-parental groups were added as controls, but gene expressions differences were assesed between and across all time points. The three tissue have unique signatures, but within each tissue we find minial sex differences. The pitutary samples displacy the msot plasticity in gene expression, with many changes in gene expression mirroring the typical rise and fall of circulating prolatin that peaks when chicks hatch. Our analysis provides new insight into how suites of genes respond in concert to the demands of offspring care. This data can be used to develop and test hypotheses about the mechanism regulating parental care behavior. 
author_summary: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
bibliography: mybibfile.bib
output: 
  rticles::plos_article:
    df_print: kable
csl: plos.csl
---

```{r, include=F}
## got error LaTeX Error: File `changepage.sty' not found 
## googled LaTeX Error: File `changepage.sty' not found Rmarkdown
## found this https://stackoverflow.com/questions/34252903/knit-pdf-file-from-rstudio
## ran the following

# install.packages(c('tinytex', 'rmarkdown'))
# tinytex::install_tinytex()

library(cowplot)
library(kableExtra)
```

<!--
_Text based on plos sample manuscript, see [http://journals.plos.org/ploscompbiol/s/latex](http://journals.plos.org/ploscompbiol/s/latex)_
-->

# Introduction

Understanding the mechanisms underlying parental care are critical to circumventing issues with parent-newborn bonding as well, where ultimate explanations are obvious, but specific mechanisms remain elusive. The rock dove (_Columba livia_) is an ideal system to characterize changes in genetic expression during parental care transitions because: 1) ample genomic resources are available (Gillespie et al. 2013), including a complete annotated genome assembly @Shapiro1063 and methodology concerning reproductive physiology and behavior (Dong et al. 2012); and 2) rock doves are prolific, year-round breeders that thrive in captivity, making observation, manipulation, and sampling highly feasible year-round.

Rock doves are socially monogamous and offer bi-parental care, making inter- and intra-sexual comparisons possible. Birds offer two important behavioral transition points into parental care: the incubation of eggs and the caring for chicks. This produces two unique opportunities to study how the brain transitions into two different suites of parental care behaviors. Additionally, rock doves exhibit a parental care strategy analogous to mammals in that they, too, ‘lactate’ to feed their young (Gillespie et al. 2011, 2012). This lactation, unlike simple regurgitation of food, consists of the production and sloughing off of skin cells inside the crop sac of females and males, creating a protein-rich milk-like substance on which they rear their chicks. Many functional similarities between rock dove and mammalian lactation exist concerning the mediation of this event by the hormone prolactin (Dumont 1965). Additionally, like mammalian milk, rock dove milk delivers essential immunoglobulins and nutritional benefits to young, aiding in their immune function and development of microbiota @Gillespie2012. Thus, because rock doves incubate eggs and exhibit mammalian-like mediation and function of lactation for young, they have the potential to serve as a powerful theoretical bridge to understand the neurobiology of both avian and mammalian transitions into parental care. Our working hypothesis is that distinct changes in transcription occur in the brain at the anticipation of, during, and in response to two different types of parental care: incubation behavior and hatchling care.

# Materials and Methods

##  Rock dove housing

Sexually mature male and female rock doves were  housed together in rooftop aviaries on the Barnard Campus. Natural male-female pairings permitted, and small observation windows behind each nest box (64 boxes total) facilitate observations nest building, egg laying, chick hatching, and nesting care every morning from 0900-1100 (correct time?). On specific collection days (Fig 1A, needs to be updated), truck blook, the hypothalamus, the pituitary, and the gonads were collected following common protocols used to sample tissue from birds (Fig 1B). Tissue collection, RNA isolation, cDNA library preparation, sequencing, quality control, and read processing were done as described in Calisi et al. 2018, Lang et al 2019, and Austin et al 2019.

```{r, fig.width= 4, echo = F}
a <- ggdraw() + draw_image("../figures/images/samplingtimepoints.png") 
b <- ggdraw() + draw_image("../figures/images/PigeonHPGaxis.png")

plot_grid(a,b, labels = "AUTO", label_size = 8, ncol = 1)
```

We used EdgeR and DESeq2 to calculate statistical differnces in gene expression between each transition. All subsequent analyses were conducting using variance stabilized data from the DESeq2 model.

# Results

Thosands of genes were found to be differentially expressed between the non-parental control group and the building group (Table 1, Fig. 1). Within the time course of parental care, fewer than 1000 genes change there expression over the course of a few days, with the most pronounced difference in toltal number of changes occuring in the female pituitary between incubation days 9 and 17 (Table 1, Fig. 1).

\begin{table}[]
\begin{tabular}{llllllll}
\hline
\textbf{Method} & \textbf{Contrast} & \textbf{Hyp F} & \textbf{Hyp M} & \textbf{Pit F} & \textbf{Pit M} & \textbf{Gon F} & \textbf{Gon M} \\  \hline
EdgeR & Cont - Bldg & 14674 & 117 & 14593 & 144 & 14859 & 4 \\
EdgeR & Bldg - Lay & 14742 & 0 & 14593 & 0 & 14836 & 0 \\
EdgeR & Lay - Inc3 & 0 & 0 & 0 & 0 & 19 & 0 \\
EdgeR & Inc3 - Inc9 & 0 & 1 & 0 & 3 & 2 & 0 \\
EdgeR & Inc9 - Inc 17 & 0 & 2 & 57 & 37 & 9 & 0 \\
EdgeR & Inc17 - Hatch & 0 & 0 & 3 & 1 & 3 & 0 \\
EdgeR & Hatch - Nest5 & 1 & 0 & 3 & 0 & 0 & 0 \\
EdgeR & Nest5 - Nest9 & 0 & 0 & 0 & 0 & 24 & 0 \\  \hline
DESeq2 & Cont - Bldg & 3452 & 4499 & 3950 & 4680 & 4597 & 2135 \\
DESeq2 & Bldg - Lay & 0 & 0 & 101 & 0 & 104 & 1 \\
DESeq2 & Lay - Inc3 & 0 & 0 & 18 & 1 & 440 & 1 \\
DESeq2 & Inc3 - Inc9 & 1 & 0 & 0 & 1 & 8 & 0 \\
DESeq2 & Inc9 - Inc 17 & 1 & 0 & 640 & 186 & 1 & 0 \\
DESeq2 & Inc17 - Hatch & 2 & 0 & 3 & 18 & 1 & 0 \\
DESeq2 & Hatch - Nest5 & 166 & 0 & 340 & 3 & 1 & 0 \\
DESeq2 & Nest5 - Nest9 & 0 & 0 & 1 & 23 & 19 & 0 \\ \hline
\end{tabular}
\caption{Comparison of total DEGs determined by edgeR and DESeq2}
\label{tab:my-table}
\end{table}

```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
ggdraw() + draw_image("../figures/characterization/linegraphs-1.png") 
```

**Fig 1. The magnitude of gene expression changes between each parental transition**.
Using DESeq 2, we identified 4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad. 500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5.

## Fig 2: Ability to use gene expression to distingush groups

```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
a <- ggdraw() + draw_image("../figures/sexes/LDA-4.png") 
b <- ggdraw() + draw_image("../figures/sexes/DoveParentsRNAseq_model2.png")

plot_grid(a,b, labels = "AUTO", label_size = 8, ncol = 1)
```

## Fig 3: Prolactin expression in PIT (but not HYP or GON) follows same pattern as in blood

```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
ggdraw() + draw_image("../figures/characterization/prolactin-1.png") 
```

## Fig 4: Hundreds of genes show similar pattern in expression as circulating prolactin

```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
ggdraw() + draw_image("../figures/sexes/wgcna-7.png")
```

# Acknowledgments

This project is a synergistic collaboration between the PI, Rebecca Calisi-Rodríguez (expertise in avian behavior, parental care and neurobiology), co-PI, Matthew MacManes (expertise in next-generation sequencing, transcriptome assembly, and gene expression analyses), and Collaborator, Rae Silver (expertise in neurobiology, dove behavior, and decades of successful breeding and maintenance of dove colonies at Barnard College).

<!--

# Funding Statment

```{r, include=F}
## Disclosure of Funding Sources
## All PLOS authors are required to declare what support they received to carry out their research. Declaring funding sources acknowledges funders’ contributions, fulfills funding requirements, and promotes greater transparency in the research process.
## Each author must individually declare all sources of funding received for the research submitted to the journal. This information includes the name of granting agencies, grant numbers, and a description of each funder’s role. If the funder has played no role in the research, this must be stated as well. Authors are not required to provide the complete list of every single grant that supports them if the grant is not related to the research published.
## A Funding Statement is included in the metadata of each published article. The Funding Statement includes the funding information declared by the authors. Inaccurate information about funding discovered after publication may require a correction. 
## from https://journals.plos.org/plosone/s/disclosure-of-funding-sources

## See https://www.nsf.gov/awardsearch/showAward?AWpD_ID=1455957&HistoricalAwards=false
```

1. Collaborative Research: RUI: The Neural Basis of Becoming a Parent: From Genotype to Phenotype. NSF-IOS:1455957.

-->

```{r include=F}
## About the authors

#Roles, Affiliations, Orcids, competing interests
```

# References {#references .unnumbered}