### 2019-10-11 docs/ms1-charactization.Rmd
new colors

---
title: "The one characterizing the parental care stages"
output: word_document
author:
- Principal Investigators - Rebecca Calisi, Matt MacManes, John Wingfield
- Postdocs - Suzanne Austin, Rayna Harris, Jesse Krause, 
- Grad students - Victoria Farrar, April Booth, Andrew Lang
- Undergrads - Tanner Feustel, Candice Lee, Annie Bond?
- Others - Fred?, Rechelle Viernes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

- What genes differ in expression between timepoints?
- What stages are most different, stressful, responsive?
- How do genes in the HPG affect genes in other regions?
- How do male and females differ?
- How do tissues differ?
- What is the relationship between genes and hormones across timepoints?

## Methods

### Fig 1: graphical abstract of methods

![](../figures/images/timeline_timeline.png)

## Results

## Fig 2: hormones

![](../figures/hormones/characterization-1.png)

## Figure 3: general rnaseq pattern, tissue > sex > stage

_Needs to be replaced with a better PCA plot_

![](../figures/limma/plotMDS-colored-1.png)

## Figure 4: model showing ability to predict non-parental, egg care, and nestingling care

![](../figures/characterization/DoveParentsRNAseq_model2.png)

## Figure 5: PRL in prolatin mirrors PRL in bloo

![](../figures/characterization/prolactin-1.png)

## Figure 6: many genes mirror the PRL expression pattern

![](../figures/sexes/wgcna2-1.png)

## Discussion

## Conclusion

## Acknowledgements

## Data availabilty