### 2019-10-15 docs/grant.Rmd
add docs

---
title: "Grant"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Collaborative Research: RUI: The Neural Basis of Becoming a Parent: From Genotype to Phenotype

[Award Abstract #1455957](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1455957&HistoricalAwards=false)

Important changes in behavior occur with the transition to parenthood. For animals that exhibit parental care, raising offspring successfully involves transitioning from aggressive and sexual behaviors to more caring and nurturing ones. However, it is not clear how the brain brings about this transition. In a rapidly changing world, understanding how the environment affects the brain and how the brain affects the behavioral transition into parental care will shed light on how changes in environment can ultimately affect how all organisms reproduce. In an interdisciplinary collaboration between Drs. Calisi-Rodríguez and MacManes they will characterize changes in gene expression and specific protein levels in important tissues essential for reproduction, namely the hypothalamus, pituitary, and gonads, in the rock dove (Columba livia) during the transition into parental care behaviors. Like most birds, rock doves incubate eggs, but are also unique in that they "lactate" to feed their young, showing stereotypical avian and mammalian parental behaviors. This makes doves a potentially powerful theoretical bridge to understand the neurobiology of transitions into parenting. Drs. Calisi-Rodríguez and MacManes will generate a large genomic dataset and identify critical genes involved in important stages of parental care. The results will have broad potential applications including the management of captive breeding for conservation or agricultural purposes, and to inform the study of neural deficiencies in parent-offspring bonding. In addition, the researchers will develop and host an annual summer program, "Summer City Science", to expose high school students from low-income neighborhoods and underrepresented backgrounds who excel in the sciences to a range of scientific fields. This will provide students the opportunity to work closely and personally with faculty, and their research and labs. Students in both the Calisi-Rodríguez and MacManes labs will help to serve as mentors and instructors for Summer City Science to increase the scope of the training of all parties involved. The overall wider objective of this program is to broaden participation in science, creating a more efficient and diverse workforce.

Previous research has implicated specific hormones that play a role in the maintenance of parental care behavior in vertebrates, including oxytocin and vasopressin (avian homologs: mesotocin and vasotocin), vasoactive intestinal peptide, and prolactin. However, very little is known about when and how the brain transitions into parental care behavior in any vertebrate, and this knowledge is fundamentally important to understand the mechanisms that mediate parental care. The Principal Investigator, during her tenure as an NSF Postdoctoral Fellow, discovered that the hypothalamic neurohormone, gonadotropin inhibitory hormone (GnIH), caused significant changes in abundance of GnIH containing cells in both birds and mammals during the transition to parental care. Here, the researchers propose to expand this knowledge using classic offspring removal and replacement manipulations to uncover the role of GnIH during the transition to parental care behavior. In addition to this targeted approach, they propose a powerful untargeted approach to explore what other changes in the brain may be occurring during this transition. They will use high-throughput sequencing technology and immunohistochemistry to uncover all genetic and specific proteomic changes in the brain, along with their relationship to GnIH, which occur during the transition to parental care. This project will build on previous work, and the data gathered will yield the most complete characterization and understanding of changes in neural transcription and translation reported to date in any vertebrate during the transition to parental care. All data and computer code will be made publicly available in established repositories and shared with the scientific community.