### 2019-10-16 docs/ms1-charactization.Rmd
new pca plots

---
title: "The one characterizing the parental care stages"
output: md_document
author:
- Principal Investigators - Rebecca Calisi, Matt MacManes, John Wingfield
- Postdocs - Suzanne Austin, Rayna Harris, Jesse Krause, 
- Grad students - Victoria Farrar, April Booth, Andrew Lang
- Undergrads - Tanner Feustel, Candice Lee, Annie Bond?
- Others - Fred?, Rechelle Viernes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Methods

### Fig 1: Graphical abstract of methods

![](../figures/images/timeline_timeline.png)

## Results

## Fig 2: Of the hormones measured, prolactin fluctuates the most and peaks near hatch in males and females

![](../figures/hormones/characterization-1.png)

## Figure 3: PCA

_Needs to be replaced with a better PCA plot showing all samples as points_

![](../figures/pca/pca-1.png)

## Figure 4: LDA

![](../figures/sexes/LDA-4.png)

![](../figures/characterization/DoveParentsRNAseq_model2.png)

## Figure 5: PRL gene expression in the pituitary in prolactin mirrors PRL in blood

![](../figures/characterization/prolactin-1.png)

## Figure 6: Many genes mirror the PRL expression pattern

![](../figures/sexes/wgcna2-1.png)

## Discussion

- What genes differ in expression between timepoints?
- What stages are most different, stressful, responsive?
- How do genes in the HPG affect genes in other regions?
- How do male and females differ?
- How do tissues differ?
- What is the relationship between genes and hormones across timepoints?

## Conclusion

## Acknowledgements

## Data availability

https://osf.io/g3kc5/