### 2019-10-18 docs/ms2-timingmanipulation.Rmd
new pca new ms

---
title: "The one where we manipulate timing of hatch"
output: word_document
author:
- Principal Investigators - Rebecca Calisi, Matt MacManes, John Wingfield
- Postdocs - Suzanne Austin, Rayna Harris, Jesse Krause, 
- Grad students - Victoria Farrar, April Booth, Andrew Lang
- Undergrads - Tanner Feustel, Candice Lee, Annie Bond?
- Others - Fred?, Rechelle Viernes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Methods

## Results

### Fig 1: Prolactin levels are low when offspring hatch early

![](../figures/hormones/manipulation-1.png)

### Fig 2: Prolactin drops when offspring removed around hatch

![](../figures/hormones/removal-1.png)

### Fig 3. Estradiol drops when incubation prolnged or hatch extended, other hormones vary less

![](../figures/hormones/manipulation-2.png)

## Discussion

## Conclusion

## Acknowledgements

## Data availabilty