### 2019-11-10 analysis/11_pcalda.Rmd
new pca and tsne figures for talk

---
title: "PCA"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(factoextra)
library(cowplot)
library(Rtsne)
library(viridis)

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/', cache = T)

# custom theme for plots
mytheme <- function(){
  theme_minimal(base_size = 12) + 
    theme(panel.grid = element_blank(),
          panel.background = element_rect(fill = "transparent", colour = NA),
          plot.background = element_rect(fill = "transparent", colour = NA),
          legend.title = element_blank())
}
```

## data wrangle

```{r data}
pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
head(pseudocounts[1:3])

colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", 
                                "hatch", "n5", "n9"))
colData$tissue <- factor(colData$tissue, levels = c("hypothalamus", "pituitary", "gonad"))
colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))

head(colData[1:3])

geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)

subsetcounts <- function(colData, countData){

  # save counts that match colData
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols) 
  countData <- countData %>% 
    dplyr::select(X1,savecols) %>% 
    dplyr::rename(entrezid = X1)

  countData <- left_join(geneinfo, countData)
  countData <- countData %>% distinct(Name, .keep_all = TRUE) %>%
    select(-row.names)  %>% select(-geneid) %>% select(-entrezid)

  row.names(countData) <- countData$Name
  countData$Name <- NULL
  countData <- as.data.frame(t(countData))

  return(countData)
}

countData <- subsetcounts(colData, pseudocounts)
head(countData[1:3])

# confirm
row.names(countData) == row.names(colData)

euclidist <- dist(countData) # euclidean distances between the rows
```

## pca

```{r}
# make pca dataframes
mypca <- prcomp(countData)

# percent conribution
eig.val <- get_eigenvalue(mypca)
head(eig.val)

mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
mypcadf$V1 <- row.names(mypcadf)
mypcadf <- left_join(colData, mypcadf)
mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
head(mypcadf)
```

```{r pca}
a <- ggplot(mypcadf, aes(x = PC1, y = PC2, 
                         color = colData$treatment, 
                         shape = colData$tissue,
                         alpha = colData$sex)) +
  geom_point(size = 2)  + theme_minimal(base_size = 6)  +
  mytheme() +
  theme(legend.title = element_blank(), 
        axis.text = element_blank()) +
  scale_alpha_manual(values = c(0.5,1)) +
  labs(x = "PC1: 35.9% of variance", y = "PC2: 32.5% of variance")

b <-  fviz_pca_var(mypca,  labelsize = 3 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 6)) + 
      mytheme() + 
      labs(x = " ", y = " ", subtitle = NULL, title =  NULL)

png('../figures/pca/pca-1.png',width=600,height=600, units="px",bg = "transparent")
print(a)
dev.off()

png('../figures/pca/pca-2.png',width=600,height=600, units="px",bg = "transparent")
print(b)
dev.off()
```

```{r pcaloadings}
fviz_contrib(mypca, choice = "var", axes = 1, top = 5) 
fviz_contrib(mypca, choice = "var", axes = 2, top = 5) 
```

# tSNE

```{r tSNE}
tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=5, theta=0.5, dims=2)
tsne_df = as.data.frame(tsne_model$Y) 
plot(tsne_df$V1, tsne_df$V2)

# prep for adding columns
colData2 <- colData 
colData2$V1 <- NULL
tsne_df_cols <- cbind(colData2, tsne_df)

c  <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = sex, color = tissue)) +
  geom_point(size = 2) +
  mytheme() +
  scale_color_manual(values = c("#d95f02","#1b9e77", "#7570b3")) +
  labs(x = "tSNE 1", y = "tSNE 2")

d <- ggplot(tsne_df_cols, aes(x = V1, y = V2, shape = tissue, color = treatment, alpha = sex)) +
  geom_point(size = 2) +
  mytheme() +
  labs(x = "tSNE 1", y = "tSNE 2") +
  scale_alpha_manual(values = c(0.5,1))

c
d

png('../figures/pca/tSNE-2.png',width=600,height=600, units="px",bg = "transparent")
print(c)
dev.off()

png('../figures/pca/tSNE-3.png',width=600,height=600, units="px",bg = "transparent")
print(d)
dev.off()
```

# combined pca tsne