### 2019-12-02 analysis/03_DESeq2_characterization.Rmd
remove old stuff

---
title: "02_DESeq_characterization"
output: md_document
---

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)
library(ggimage)
library(knitr)
library(kableExtra)

library(png)
library(grid)
library(ggimage)
library(ggpubr)

library(BiocParallel)
register(MulticoreParam(6))

# load custom functions  
source("../R/themes.R") 
source("../R/functions.R")

knitr::opts_chunk$set(fig.path = '../figures/characterization/', cache = TRUE)
```

## Characterization data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17",
                                "hatch", "n5", "n9"))
c.colData$tissue <- factor(c.colData$tissue, levels = 
                              c("hypothalamus",  "pituitary", "gonad"))

levels(c.colData$treatment)
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
summary(c.colData[c(7,3,4,5,8)])

geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)
head(geneinfo)
```

## Run DESeq on all subsets of the data

```{r dds, cache = T, eval = T}
dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
dds.female_pituitary <- subsetDESeq(c.colData,  c.countData, "female_pituitary" )
dds.female_gonad <- subsetDESeq(c.colData,  c.countData, "female_gonad" )
dds.male_hypothalamus <- subsetDESeq(c.colData,  c.countData, "male_hypothalamus" )
dds.male_pituitary <- subsetDESeq(c.colData,  c.countData, "male_pituitary"  )
dds.male_gondad <- subsetDESeq(c.colData,  c.countData, "male_gonad")
```

## plot candidate genes

```{r candidates, eval = T}
plotcandidates(dds.female_hypothalamus, colDataFH,"female hypothalamus")
plotcandidates(dds.female_pituitary, colDataFP, "female pituitary")
plotcandidates(dds.female_gonad, colDataFG,"female gonad")
plotcandidates(dds.male_hypothalamus, colDataMH,"male hypothalamus")
plotcandidates(dds.male_pituitary,colDataMP, "male pituitary")
plotcandidates(dds.male_gondad,colDataMG, "male gonad")
```

## DEGs

```{r createDEGs}

createDEGdf <- function(mydds, whichfactor, up, down, mytissue){
  res <- results(mydds, contrast =c(whichfactor, up, down),
                 independentFiltering = T, alpha = 0.1)
   data <- data.frame(entrezid = row.names(res),
                     padj = res$padj, 
                     logpadj = -log10(res$padj),
                     lfc = res$log2FoldChange,
                     tissue = mytissue)
  data <- na.omit(data)

  data <- data %>%
    dplyr::mutate(direction = ifelse(data$lfc > 0 & data$padj < 0.1, 
                                     yes = up, no = ifelse(data$lfc < 0 & data$padj < 0.1, 
                                                 yes = down, no = "NS")))
  data$direction <- factor(data$direction, levels = c(down, "NS", up))

  data <- left_join(data, geneinfo) %>%
    mutate(gene = Name) %>%
    select(gene, lfc, padj, logpadj, direction, tissue) %>%
    arrange(desc(lfc))

  DEGs <- data %>% filter(direction != "NS")
  print(str(DEGs))

  myfilename = paste("../results/03_DEGs", mytissue, down,up, "csv", sep = ".")

  # write dataframe of only significant genes
  write.csv(DEGs, myfilename, row.names = F)

  # return data frome with all data, included NS genes
  return(data)
}

control.bldg <- createDEGdf(dds.female_pituitary, "treatment", "bldg", "control", "pituitary")
bldg.lay <- createDEGdf(dds.female_pituitary, "treatment", "lay", "bldg", "pituitary")
lay.inc.d3 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d3", "lay", "pituitary") 
inc.d3.inc.d9 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d9", "inc.d3", "pituitary") 
inc.d9.inc.d17 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
inc.d17.hatch <- createDEGdf(dds.female_pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
hatch.n5 <- createDEGdf(dds.female_pituitary, "treatment", "n5", "hatch", "pituitary") 
n5.n9 <- createDEGdf(dds.female_pituitary, "treatment", "n9", "n5", "pituitary") 
```

```{r plotDEGs}
plot.volcano <- function(data, whichfactor, up, down, mycolors){

  volcano <- data %>%
    ggplot(aes(x = lfc, y = logpadj)) + 
    geom_point(aes(color = direction, shape = tissue), size = 2, 
               alpha = 0.75, na.rm = T) + 
    theme_minimal() +
    scale_color_manual(values = mycolors,
                       name = " ",
                       drop = FALSE) +
    ylim(c(0,30)) +  
    xlim(c(-8,8)) +
    labs(y = "-log10(p)", x = " ")  +
    theme(legend.position = "none",
          legend.direction = "horizontal",
          legend.spacing.x = unit(-0.1, 'cm'),
          legend.margin=margin(t=-0.25, r=0, b=0, l=0, unit="cm"),
          panel.grid = element_blank()) +
    scale_shape_manual(values = myshapes)
  return(volcano)
}

a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)

females <- plot_grid(a + 
                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "female pituitary \n -log10(p)"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e + theme(axis.title = element_blank(), axis.text = element_blank()),
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
females 
```

## Go terms

```{r GOterms}
reproduction <- read.table("../results/goterms/reproduction.txt", sep = "\t", row.names = NULL,  fill=TRUE)
reproduction$GO <- "reproduction, GO:0000003"

parentalbehavior <- read.table("../results/goterms/parentalbehavior.txt", sep = "\t", row.names = NULL,  fill=TRUE)
parentalbehavior$GO <- "parental behavior, GO:0060746"

behavior <- read.table("../results/goterms/parentalbehavior.txt", sep = "\t", row.names = NULL,  fill=TRUE)
behavior$GO <- "behavior, GO:0007610"

systemdevelopment <- read.table("../results/goterms/systemdevelopment.txt", sep = "\t", row.names = NULL,  fill=TRUE)
systemdevelopment$GO <- "system development, GO:0048731"

endocrineprocess <- read.table("../results/goterms/endocrineprocess.txt", sep = "\t", row.names = NULL,  fill=TRUE)
endocrineprocess$GO <- "endocrine process, GO:0050886"

responsetostimulus <- read.table("../results/goterms/responsetostimulus.txt", sep = "\t", row.names = NULL,  fill=TRUE)
responsetostimulus$GO <- "response to external stimulus, GO:0009605"

GOterms <- rbind(parentalbehavior, reproduction, behavior, 
                 systemdevelopment, endocrineprocess, responsetostimulus)

GOterms <- GOterms %>%
  dplyr::mutate(gene = toupper(MGI.Gene.Marker.ID)) %>% 
  dplyr::select(gene, GO) %>% 
  dplyr::distinct(gene, GO) %>% 
 group_by(gene)

head(GOterms)

returnGOgenes <- function(whichGOterms, data, mycomparison){

  df <- inner_join(whichGOterms, data) %>% 
    filter(direction != "NS") %>%  
    select(gene, GO) %>% 
    group_by(GO) %>% 
    summarize(genes = str_c(gene, collapse = ", "))  %>% 
    mutate(comparison = mycomparison) %>% 
    select(comparison, GO, genes)

  return(df)
}

dt1 <- returnGOgenes(GOterms,control.bldg, "control v bldg") 
dt2 <- returnGOgenes(GOterms,bldg.lay , "bldg v lay") 
dt3 <- returnGOgenes(GOterms,lay.inc.d3 , "lay v inc d3")  
dt4 <- returnGOgenes(GOterms,inc.d3.inc.d9 , "inc d3 v inc d9")  
dt5 <- returnGOgenes(GOterms,inc.d9.inc.d17 , "inc d9 v inc d17")  
kable(dt5)

dt6 <- returnGOgenes(GOterms,inc.d17.hatch, "inc d17 v hatch")  
dt7 <- returnGOgenes(GOterms,hatch.n5, "hatch v n5")  
kable(dt7)

dt8 <- returnGOgenes(GOterms,n5.n9, "n5 v n9")

alldt <- rbind(dt1,dt2,dt3,dt4,dt5,dt6,dt7,dt8)
alldt

write.csv(alldt, "../results/03_GO_femalepituitary.csv")
```

## males and females

```{r createDEGsmales}
control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control", "pituitary")
bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg", "pituitary")
lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay", "pituitary") 
inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3", "pituitary") 
inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch", "pituitary") 
n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5", "pituitary")

```

```{r plotDEGsmales}
a2 <- plot.volcano(control.bldg2, "treatment", "bldg", "control", colorsvolcano)
b2 <- plot.volcano(bldg.lay2, "treatment", "lay", "bldg", colorsvolcano)
c2 <- plot.volcano(lay.inc.d32, "treatment", "inc.d3", "lay", colorsvolcano) 
d2 <- plot.volcano(inc.d3.inc.d92, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e2 <- plot.volcano(inc.d9.inc.d172, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f2 <- plot.volcano(inc.d17.hatch2, "treatment", "hatch", "inc.d17", colorsvolcano) 
g2 <- plot.volcano(hatch.n52, "treatment", "n5", "hatch", colorsvolcano) 
h2 <- plot.volcano(n5.n92, "treatment", "n9", "n5", colorsvolcano)

males <- plot_grid(a2 + labs(y = "male pituitary \n -log10(p)") + theme(axis.text.x = element_blank()),
          b2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          c2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          d2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          e2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          f2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          g2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          h2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
males

```

### tissue levels, ignoring sex differences, sort of

```{r gonaddfs, eval=T}

dds.gonads <- subsetDESeq2(c.colData,  c.countData, c("female_gonad","male_gonad"))
```

```{r gonaddfs2, eval=T}
control.bldg <- createDEGdf(dds.gonads, "treatment", "bldg", "control",  "gonads")
bldg.lay <- createDEGdf(dds.gonads, "treatment", "lay", "bldg",  "gonads")
lay.inc.d3 <- createDEGdf(dds.gonads, "treatment", "inc.d3", "lay",  "gonads") 
inc.d3.inc.d9 <- createDEGdf(dds.gonads, "treatment", "inc.d9", "inc.d3",  "gonads") 
inc.d9.inc.d17 <- createDEGdf(dds.gonads, "treatment", "inc.d17", "inc.d9",  "gonads")
inc.d17.hatch <- createDEGdf(dds.gonads, "treatment", "hatch", "inc.d17",  "gonads") 
hatch.n5 <- createDEGdf(dds.gonads, "treatment", "n5", "hatch",  "gonads") 
n5.n9 <- createDEGdf(dds.gonads, "treatment", "n9", "n5",  "gonads") 
```

```{r gonplot, eval=T}
a2 <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
b2 <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
c2 <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
d2 <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e2 <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f2 <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
g2 <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
h2 <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)

gonads <- plot_grid(a2 + labs(y = "gonads \n -log10(p)"),
          b2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          c2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          d2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
            labs(x = "log fold change"), 
          e2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          f2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          g2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          h2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
gonads

```

```{r hypdfs, eval=T}
dds.hypothalamus <- subsetDESeq2(c.colData,  c.countData, c("female_hypothalamus","male_hypothalamus"))
```

```{r hypdfs2, eval=T}
control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control" ,  "hypothalamus")
bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg" ,  "hypothalamus")
lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay" ,  "hypothalamus") 
inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3" ,  "hypothalamus") 
inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9" ,  "hypothalamus")
inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17" ,  "hypothalamus") 
hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch" ,  "hypothalamus") 
n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5" ,  "hypothalamus") 
```

```{r hypplot, eval = T}
a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)

hypothalamus <- plot_grid(a + 
                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "hypothalamus \n -log10(p)"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e + theme(axis.title = element_blank(), axis.text = element_blank()),
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
hypothalamus

```

```{r pitdfs, eval=T}
dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary","male_pituitary"))
```

```{r pitdfs2, eval=T}
control.bldg <- createDEGdf(dds.pituitary, "treatment", "bldg", "control", "pituitary")
bldg.lay <- createDEGdf(dds.pituitary, "treatment", "lay", "bldg", "pituitary")
lay.inc.d3 <- createDEGdf(dds.pituitary, "treatment", "inc.d3", "lay", "pituitary") 
inc.d3.inc.d9 <- createDEGdf(dds.pituitary,  "treatment", "inc.d9", "inc.d3", "pituitary") 
inc.d9.inc.d17 <- createDEGdf(dds.pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
inc.d17.hatch <- createDEGdf(dds.pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
hatch.n5 <- createDEGdf(dds.pituitary, "treatment", "n5", "hatch", "pituitary") 
n5.n9 <- createDEGdf(dds.pituitary, "treatment", "n9", "n5", "pituitary") 
```

```{r pitplot, eval = T}
a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)

pituitary <- plot_grid(a + 
                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "pituitary \n -log10(p)"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e + theme(axis.title = element_blank(), axis.text = element_blank()),
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
pituitary
```

```{r allvolcanos, fig.height=6}
background <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
icons <- ggdraw() +  draw_image(background, scale = 1)

plot_grid(icons, hypothalamus, females, males, gonads, nrow = 5, rel_heights = c(1,1,1,1,1.25))

plot_grid(icons, hypothalamus, pituitary, gonads, nrow = 4, rel_heights = c(1,1,1,1.25))
```