### 2019-12-02 analysis/09_manipvolcanos.Rmd
now with manip volcanos

---
title: "Manip Volcanos"
output: md_document
---

```{r}
library(tidyverse)
library(cowplot)

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")  # for images

knitr::opts_chunk$set(fig.path = '../figures/manipulation/',message=F, warning=FALSE)
```

```{r  eval=T}

inc.d3.m.inc.d3 <- read_csv("../results/DESeqAll.pituitary.inc.d3.m.inc.d3.csv")
inc.d9.m.inc.d9 <- read_csv("../results/04_DEGs.pituitary.m.inc.d3.m.inc.d9.csv")
inc.d17.m.inc.d17 <- read_csv("../results/DESeqAll.pituitary.inc.d17.m.inc.d17.csv")
hatch.m.n2 <- read_csv("../results/DESeqAll.pituitary.hatch.m.n2.csv")
inc.d9.m.inc.d8 <- read_csv("../results/DESeqAll.pituitary.inc.d9.m.inc.d8.csv")
hatch.m.inc.d8 <- read_csv("../results/DESeqAll.pituitary.hatch.m.inc.d8.csv")
inc.d17.prolong <- read_csv("../results/DESeqAll.pituitary.inc.d17.prolong.csv")
hatch.prolong <- read_csv("../results/DESeqAll.pituitary.hatch.prolong.csv")
hatch.extend <- read_csv("../results/DESeqAll.pituitary.hatch.extend.csv")
n5.extend <- read_csv("../results/DESeqAll.pituitary.n5.extend.csv")

```

```{r volcanos}
a <- plot.volcano(inc.d3.m.inc.d3 , "treatment",   "m.inc.d3", "inc.d3",   colorsvolcano) + 
  annotation_custom(lay, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(control, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
b <- plot.volcano(inc.d9.m.inc.d9, "treatment",  "m.inc.d9", "inc.d9" , colorsvolcano)+ 
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(control, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
c <- plot.volcano(inc.d17.m.inc.d17 , "treatment",  "m.inc.d17", "inc.d17" , colorsvolcano) + 
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(control, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
d <- plot.volcano(hatch.m.n2 , "treatment",   "m.n2", "hatch", colorsvolcano) + 
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(control, ymin = 20, ymax = 30, xmin = 2, xmax = 8)

e <- plot.volcano(inc.d9.m.inc.d8 , "treatment",   "m.inc.d8", "inc.d9"  , colorsvolcano)  +
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
f <- plot.volcano(inc.d17.prolong , "treatment",   "prolong", "inc.d17",  colorsvolcano)  +
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
g <- plot.volcano(hatch.prolong , "treatment",  "prolong", "hatch", colorsvolcano)  +
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(inc, ymin = 20, ymax = 30, xmin = 2, xmax = 8)

h <- plot.volcano(hatch.m.inc.d8 , "treatment",    "m.inc.d8", "hatch"  , colorsvolcano)  +
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
i <- plot.volcano(hatch.extend , "treatment",  "extend", "hatch" , colorsvolcano)  +
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = 2) + 
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = 2, xmax = 8) 
j <- plot.volcano(n5.extend , "treatment",  "extend", "n5", colorsvolcano)  +
  annotation_custom(hatch, ymin = 20, ymax = 30, xmin = -8, xmax = -2) + 
  annotation_custom(nestling, ymin = 20, ymax = 30, xmin = 2, xmax = 8)

remove <- plot_grid(a +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "pituitary \n -log10(p)", subtitle = "1771 DEGs"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "277 DEGs"),
          c + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "3909 DEGs"),
          d + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "3162 DEGs"), 
          nrow = 1, rel_widths = c(1.2,1,1,1),
          labels = c("a","b","c","d"), label_size = 9)
remove

manipinc <- plot_grid(e +  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "pituitary \n -log10(p)", subtitle = "120 DEGs"),
          f + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "1007 DEGs"),
          g + theme(axis.title = element_blank(), axis.text = element_blank()) + labs(subtitle = "731 DEGs"),
          nrow = 1, rel_widths = c(1.2,1,1),
          labels = c("e","f","g"," "), label_size = 9)

maniphatch <- plot_grid( h +  labs(subtitle = "4113 DEGs"), 
          i + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) + labs(subtitle = "1894 DEGs"), 
          j + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) + labs(subtitle = "256 DEGs"), 
          nrow = 1, rel_widths = c(1.2,1,1),
          labels = c("h","i","j"," "), label_size = 9)

plot_grid(manipinc, maniphatch, nrow = 2)
```