### 2019-12-20 analysis/20_forrechelle.Rmd
working on candidate gnee plots

---
title: "forrechelle"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)

source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/favegenes/")
```

```{r input}

candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL")
timepoints <- c("lay" , "inc.d9", "hatch", "n9")
hormones <- c("PRL", "CORT")

```

```{r rechelle}

datapath <- "../results/"   # path to the data
datafiles <- dir(datapath, pattern = "*allvsd.csv") # get file names
datapathfile <- paste0(datapath, datafiles)

df <- datapathfile %>%
  setNames(nm = .) %>%
  map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
  mutate(tissue = sapply(strsplit(as.character(filename),'../results/06_'), "[", 2)) %>% 
  mutate(tissue = sapply(strsplit(as.character(tissue),'allvsd.csv'), "[", 1))  %>% 
  select(tissue, X1, everything()) 
head(df)

df <- df %>% filter(tissue != "gon")
head(df)

df2 <-  df  %>%
  filter(X1 %in% candidategenes) %>%
  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "sample", values_to = "vsd") %>%
   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
  filter(treatment %in% timepoints) %>%
  select(bird, sex, treatment, tissue, X1, vsd) %>%
  rename(gene = X1) %>%
  drop_na()
head(df2)

df3 <- df2 %>% 
  mutate(tissue = fct_relevel(tissue,  "hyp", "pit", "gon")) %>% 
  mutate(treatment = fct_relevel(treatment, timepoints)) %>% 
  group_by(sex, treatment, tissue, gene)  %>% 
  summarize(m = mean(vsd, na.rm = T), se = sd(vsd,  na.rm = T)/sqrt(length(vsd))) %>%
  mutate(image = "../figures/images/DoveParentsRNAseq_note.png")
head(df3)

for (i in levels(df3$tissue)) {

p <-  df3 %>%
  filter(tissue == i) %>%
  ggplot(aes(x = treatment, y = m)) +
  geom_image(aes(image=image, color = gene), size=0.1)  +

  scale_alpha_manual(values = c(0.5, 1)) +
  labs(subtitle = i, y = "average expression", x = "parental stage")

pimage <- axis_canvas(p, axis = 'x') + 
  draw_image("../figures/images/icons/DoveParentsRNAseq_lay.png", x = 0.5, scale = 1 ) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_incubation.png", x = 1.5, scale = 1) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_hatch.png", x = 2.5, scale = 1) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_chicklings.png", x = 3.5, scale = 1)

p2 <- ggdraw(insert_xaxis_grob(p, pimage, position = "bottom"))
plot(p2)

}

for (i in levels(df3$tissue)) {

  p <-  df3 %>%
    filter(tissue == i) %>%
    ggplot(aes(x = treatment, y = m)) +
    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
    geom_point(size = 1, aes(color = gene)) +
    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = i, y = "average expression", x = "parental stage") +
    facet_wrap(~sex, nrow = 1) +
    theme_B3()
  plot(p)

}

```