### 2019-12-22 analysis/02_PCA.Rmd
new pca images

---
title: "PCA"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggfortify)
library(factoextra)
library(cowplot)
library(Rtsne)
library(viridis)
library(forcats)

source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/pca/', cache = T)

```

# note. this file is too big for storage in github :(

## Data: Pseudocounts from Limma

```{r read}
pseudocounts <- read_csv("../results/01_pseudo.counts.csv")
head(pseudocounts[1:3])
pseudocounts <- as.data.frame(pseudocounts)
row.names(pseudocounts) <- pseudocounts$X1
pseudocounts$X1 <- NULL
# prep count data for all samples
countData <- as.data.frame(t(pseudocounts))
head(countData[1:3])
```

## All data, characterization and manipulations

```{r colData}
# prep col data for all samples
colData <- read.csv("../metadata/00_samples.csv", header = T, row.names = 1)
colData$treatment <- factor(colData$treatment, levels = alllevels)
colData <- colData %>% mutate(tissue = fct_recode(tissue, "gonads" = "gonad"))
colData$tissue <- factor(colData$tissue, levels = tissuelevels)
row.names(colData) <- colData$V1

# check ready for analysis
#row.names(countData) == row.names(colData)

head(colData)
```

```{r tSNE-df}
# prep for tsne

subsetmaketsne <- function(whichtissue, whichtreatment, whichsex){

  colData <- colData %>%
    dplyr::filter(tissue %in% whichtissue,
                  treatment %in% whichtreatment,
                  sex %in% whichsex) 
  row.names(colData) <- colData$V1

  # save counts that match colData
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- as.data.frame(t(countData))
  countData <- countData %>% dplyr::select(one_of(savecols)) 
  countData <- as.data.frame(t(countData))

  euclidist <- dist(countData) # euclidean distances between the rows

  tsne_model <- Rtsne(euclidist, check_duplicates=FALSE, pca=TRUE, perplexity=10, theta=0.5, dims=2)
  tsne_df = as.data.frame(tsne_model$Y)

  # prep for adding columns
  colData2 <- colData 
  colData2$V1 <- NULL
  tsne_df_cols <- cbind(colData2, tsne_df)
  return(tsne_df_cols)
}

chartsne <- subsetmaketsne(tissuelevels, charlevels, sexlevels)
hyptsne <- subsetmaketsne("hypothalamus", charlevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", charlevels, sexlevels)
gontsne <- subsetmaketsne("gonads", charlevels, sexlevels)

```

```{r tSNE}
plotcolorfultsnes <- function(tsnedf, whichfactor, whichcolors){
  p <- ggplot(tsnedf, aes(x = V1, y = V2, color = whichfactor, shape = tissue)) +
    geom_point(size = 1) +
    theme_B3() +
    labs(x = "tSNE 1", y = "tSNE 2") +
    scale_color_manual(values = whichcolors) +
    theme(legend.position = "none",
          axis.text = element_blank()) +
    scale_shape_manual(values = myshapes)
  print(p)
}

a <- plotcolorfultsnes(chartsne, chartsne$tissue, colorstissue)  
b <- plotcolorfultsnes(chartsne, chartsne$sex, sexcolors)   
c <- plotcolorfultsnes(chartsne, chartsne$treatment, colorscharmaip)

abc <- plot_grid(a,b,c, nrow = 1, labels = c("a", "b", "c"), label_size = 12)
abc

```

```{r subsetmakepca}
subsetmakepca <- function(whichtissue, whichtreatment, whichsex){

  colData <- colData %>%
    dplyr::filter(tissue %in% whichtissue,
                  treatment %in% whichtreatment,
                  sex %in% whichsex) 
  row.names(colData) <- colData$V1

  # save counts that match colData
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- as.data.frame(t(countData))
  countData <- countData %>% dplyr::select(one_of(savecols)) 
  countData <- as.data.frame(t(countData))

  mypca <- prcomp(countData)

  mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],
                  ID = row.names(countData))
  mypcadf$V1 <- row.names(mypcadf)
  mypcadf <- left_join(colData, mypcadf)
  mypcadf <- mypcadf %>% select(bird,sex,tissue,treatment,PC1:PC6)
  return(mypcadf)
}

charpca <- subsetmakepca(tissuelevels, charlevels, sexlevels)
hyppca <- subsetmakepca("hypothalamus", charlevels, sexlevels)
pitpca <- subsetmakepca("pituitary", charlevels, sexlevels)
gonpca <- subsetmakepca("gonads", charlevels, sexlevels)

```

```{r pca}
plotcolorfulpcs <- function(mypcadf,  whichfactor, whichcolors){
  p <- mypcadf %>%
    ggplot(aes(x = PC1, y = PC2, shape = tissue, color = whichfactor )) +
    geom_point(size = 1)  +
    theme_B3() +
    theme(legend.title = element_blank(),
         axis.text = element_blank(),
         legend.position = "none") +
    labs(x = "PC1", y = "PC2")  +
    scale_color_manual(values = whichcolors) +
    scale_shape_manual(values = myshapes)
  print(p)
}

d <- plotcolorfulpcs(charpca, charpca$tissue, colorstissue)  
e <- plotcolorfulpcs(charpca, charpca$sex, sexcolors) 
f <- plotcolorfulpcs(charpca,charpca$treatment, colorscharmaip)

def <- plot_grid(d,e, f, nrow =  1,  labels = c("d", "e", "f"), label_size = 12)
def
```

```{r fviz}

makefvizdf <-  function(whichtissue, whichtreatment, whichsex){
  colData <- colData %>%
      dplyr::filter(tissue %in% whichtissue,
                  treatment %in% whichtreatment,
                  sex %in% whichsex) 
  row.names(colData) <- colData$V1

  # save counts that match colData
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- as.data.frame(t(countData))
  countData <- countData %>% dplyr::select(one_of(savecols)) 
  countData <- as.data.frame(t(countData))

  mypca <- prcomp(countData)
  return(mypca)
}

charfviz <- makefvizdf(tissuelevels, charlevels, sexlevels)
hypv <- makefvizdf("hypothalamus", charlevels, sexlevels)
pitfviz <- makefvizdf("pituitary", charlevels, sexlevels)
gonfviz <- makefvizdf("gonads", charlevels, sexlevels)

g <- fviz_screeplot(charfviz, addlabels = TRUE, ylim = c(0, 50),  ncp = 5, barcolor = "white", barfill = "white") + 
  labs(title = NULL, y = "PC Variance") + theme_B3() 
h <- fviz_pca_var(charfviz,  labelsize = 3 , axes.linetype = "blank", 
                   repel = TRUE ,
                  select.var= list(contrib = 5)) + 
      theme_B3() + 
      labs(x = "PC1 (35.8%)", y = "PC2 (32.5%)", title =  NULL) +
      theme( axis.text = element_blank())

legend <- png::readPNG("../figures/images/DoveParentsRNAseq_legendchar.png")
legend <-  grid::rasterGrob(legend, interpolate=TRUE)

gh <- plot_grid(g,h, legend, nrow = 1, rel_widths = c(1, 1, 1), labels = c("h", "i", NULL ), label_size = 12)
gh
plot_grid(def, gh, nrow = 2)
```

```{r PCA-tSNE, fig.height= 7, fig.width=7}
plot_grid( abc, def, gh, nrow = 3)

```

## tissue specific

```{r tissuespecific, fig.height=7,fig.width=7}

## tsne 
h4 <- plotcolorfultsnes(hyptsne, hyptsne$sex, sexcolors)  + labs(y = "hypothalamus") 
h5 <- plotcolorfultsnes(hyptsne, hyptsne$treatment, colorscharmaip)

p4 <- plotcolorfultsnes(pittsne, pittsne$sex, sexcolors)   + labs(y = "pituitary") 
p5 <- plotcolorfultsnes(pittsne, pittsne$treatment, colorscharmaip)

g4 <- plotcolorfultsnes(gontsne, gontsne$sex, sexcolors)  + labs(y = "gonads")  
g5 <- plotcolorfultsnes(gontsne, gontsne$treatment, colorscharmaip)

## pca

h1 <- plotcolorfulpcs(hyppca, hyppca$sex, sexcolors) 
h2 <- plotcolorfulpcs(hyppca, hyppca$treatment, colorscharmaip)

p1 <- plotcolorfulpcs(pitpca, pitpca$sex, sexcolors) 
p2 <- plotcolorfulpcs(pitpca, pitpca$treatment, colorscharmaip)

g1 <- plotcolorfulpcs(gonpca, gonpca$sex, sexcolors) 
g2 <- plotcolorfulpcs(gonpca, gonpca$treatment, colorscharmaip)

plot_grid(h1,h2,p1,p2,g1,g2, nrow = 3)

# screeplot for variance
fviz_screeplot(hypv, addlabels = TRUE,   ncp = 5) + 
  labs(title = NULL, y = "PC Variance") + theme_B3() 
fviz_screeplot(pitfviz, addlabels = TRUE,   ncp = 5) + 
  labs(title = NULL, y = "PC Variance") + theme_B3() 
fviz_screeplot(gonfviz, addlabels = TRUE,    ncp = 5) + 
  labs(title = NULL, y = "PC Variance") + theme_B3()

# fviz contributions

plotfriz <- function(frizdf){

     p <- fviz_pca_var(frizdf,  labelsize = 3.5 , axes.linetype = "blank", 
                   repel = T ,  select.var= list(contrib = 3))  + 
      labs(title = NULL) + 
      theme_B3() + 
      theme(axis.text = element_blank())
   print(p)
}

h3 <- plotfriz(hypv) + labs(x = "PC1 (59.8%)", y = "PC2 (37.7%)") 
p3 <-plotfriz(pitfviz) + labs(x = "PC1 (54.3%)", y = "PC2 (35.7%)") 
g3 <- plotfriz(gonfviz) + labs(x = "PC1 (82.7%)", y = "PC2 (11.6%)")

## all together now

plot_grid(h4,h5,h2,h3,
          p4,p5,p2,p3,
          g4,g5,g2,g3,
          nrow = 3,
          labels = "auto", label_size = 10)
```