### 2019-12-23 analysis/17_pheatmap.Rmd
first draft at group heatmap

---
title: "heatmap"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pheatmap)

source("../R/themes.R")

knitr::opts_chunk$set(fig.path = '../figures/pheatmap/', cache = TRUE,  message=F, comment=FALSE, warning=FALSE)
```

```{r, pheatmap, fig.height=6.5}
prepvsds <- function(pathtofile){

  df <- read.csv(pathtofile, row.names = 1)

  df <- tail(df, 5000)
  df <- head(df, 1000)

  df <- as.data.frame(t(df))
  df$V1 <- row.names(df)

  df$sex <- sapply(strsplit(as.character(df$V1),'\\_'), "[", 2)
  df$tissue <- sapply(strsplit(as.character(df$V1),'\\_'), "[", 3)
  df$treatment <- sapply(strsplit(as.character(df$V1),'\\_'), "[", 4)
  df$treatment <- sapply(strsplit(as.character(df$treatment), ".NYNO"), "[", 1)

  df$treatment <- ifelse(grepl("extend.hatch", df$treatment), "extend",
                   ifelse(grepl("inc.prolong", df$treatment), "prolong",
                   ifelse(grepl("m.hatch", df$treatment), "m.n2",
                                   df$treatment)))

  df$tissue <- ifelse(grepl("gonad", df$tissue), "gonads", df$tissue)

  df$sextissuetreatment <- paste(df$sex, df$tissue, df$treatment, sep = "_")

  newcolnames <- df$sextissuetreatment

  df$V1 <- NULL
  df$sex <- NULL
  df$tissue <- NULL
  df$treatment <- NULL 
  df$sextissuetreatment <- NULL

  df <- as.data.frame(t(df))

  colnames(df) <- newcolnames

  df <- sapply(unique(colnames(df)), function(i)
    rowMeans(df[,colnames(df) == i]))

  return(df)
}

vsd.hyp <- prepvsds("../results/06_hypallvsd.csv")
vsd.pit <- prepvsds("../results/06_pitallvsd.csv")
vsd.gon <- prepvsds("../results/06_gonallvsd.csv")

makelabels <- function(df){

  mylabels <- as.data.frame(colnames(df))
  mylabels$treatment <- sapply(strsplit(as.character(mylabels$`colnames(df)`),'\\_'), "[", 3)
  mylabels$sex <- sapply(strsplit(as.character(mylabels$`colnames(df)`),'\\_'), "[", 1)
  mylabels$tissue <- sapply(strsplit(as.character(mylabels$`colnames(df)`),'\\_'), "[", 2)
  row.names(mylabels) <- mylabels$`colnames(df)`
  mylabels$`colnames(df)` <- NULL
  return(mylabels)
}

hyplabels <- makelabels(vsd.hyp)
pitlabels <- makelabels(vsd.pit)
gonlabels <- makelabels(vsd.gon)

pheatmap(vsd.hyp, show_colnames = F, annotation_col = hyplabels, 
         annotation_colors = myannotationcolors, show_rownames = F)
pheatmap(vsd.pit, show_colnames = F, annotation_col = pitlabels, 
         annotation_colors = myannotationcolors, show_rownames = F)
pheatmap(vsd.gon, show_colnames = F, annotation_col = gonlabels, 
         annotation_colors = myannotationcolors, show_rownames = F)
```