### 2019-12-28 analysis/04_LDA_v3.Rmd
unsuccessful first attempt at incorporating stress data

---
title: "04_LDA_v3"
output: html_document
---

```{r setup, include=FALSE}
suppressMessages(library(tidyverse)) 
library(forcats)
library(tidyselect) # for vars_select

knitr::opts_chunk$set(echo = TRUE)
```

## stress data

```{r}
# genes info
geneinfo <- read_csv("../metadata/00_geneinfo.csv") %>% dplyr::select(Name, entrezid)

# prep stress data
stress_coldata <- read_csv("../../austinCORT/data/03_colDataSTRESS.csv")
stress_countdata <- read.csv("../../austinCORT/data/03_countDataSTRESS.csv", row.names = 1)
head(stress_coldata)
head(stress_countdata)

subsetcoldata <- function(whichtissue){
  df <- stress_coldata %>% 
    mutate(tissue = fct_recode(tissue,
           "hypothalamus" = "hyp",
           "pituitary" = "pit",
           "gonads" = "gon")) %>% 
    filter(tissue == whichtissue) %>% 
    rename("V1" = "sample")  %>% 
    mutate(V1 = str_replace(V1, '.male', "_male"))  %>%
    mutate(V1 = str_replace(V1, '.f_male', "_female"))  %>% 
    mutate(bird = sapply(strsplit(as.character(V1),'\\_'), "[", 1))  %>% 
    mutate(group = sapply(strsplit(as.character(V1),'\\_'), "[", 2))  %>% 
    mutate(V1 = str_replace(V1, '_', "."))  %>%
    mutate(V1 = str_replace(V1, "r.r.x.R2XR.", "r.r.x.R2XR_")) %>%
    mutate(study = "stress") %>% 
    dplyr::select(V1, bird, sex, tissue, treatment, group, study)  %>% 
    droplevels()
  print(head(df))
  return(df)
}

stress_coldata_hyp <- subsetcoldata("hypothalamus")
stress_coldata_pit <- subsetcoldata("pituitary")
stress_coldata_gon <- subsetcoldata("gonads")

selectvsdstress <- function(stresscolData){

  savecols <- stress_coldata_hyp$V1
  savecols <- as.vector(savecols)

  df <- stress_countdata %>% dplyr::select(one_of(savecols)) 
  df$entrezid <- row.names(df)

  df <- left_join(geneinfo, df) %>% 
    dplyr::select(-entrezid) %>% 
    rename(gene = Name) %>%
    pivot_longer(-gene, names_to = "samples", values_to = "counts") %>%
    pivot_wider(
    names_from = samples, 
    values_from = counts,
    values_fn = list(counts = sum))  %>% 
  arrange(gene) %>% 
  filter(gene != "")

  df <- as.data.frame(df)
  row.names(df) <- df$gene
  df$gene <- NULL
  df <- as.data.frame(t(df))

  print(head(df))
  return(df)
}

stress_countdata_hyp <- selectvsdstress(stress_coldata_hyp)
stress_countdata_pit <- selectvsdstress(stress_coldata_pit)
stress_countdata_pit <- selectvsdstress(stress_coldata_gon)
```

## parentalcare data

```{r}
colData <- read.csv("../metadata/00_samples.csv", row.names = 1)

charHyp <- colData %>% filter(study == "charcterization", tissue == "hypothalamus") %>% droplevels()
charPit <- colData %>% filter(study == "charcterization", tissue == "pituitary") %>% droplevels()
charGon <- colData %>% filter(study == "charcterization", tissue == "gonad") %>% 
  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>% droplevels()

manipHyp <- colData %>% filter(study == "manipulation", tissue == "hypothalamus") %>% droplevels()
manipPit <- colData %>% filter(study == "manipulation", tissue == "pituitary") %>% droplevels()
manipGon <- colData %>% filter(study == "manipulation", tissue == "gonad") %>% 
  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>% droplevels()

selectvsd <- function(pathtofile, colData){

  df <- read.csv(pathtofile, row.names = 1)
  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols) 
  df <- df %>% dplyr::select(one_of(savecols))

  # keep only 100 genes fornow
  #df <- head(df, 100)

  df <- as.data.frame(t(df))
  df$V1 <- row.names(df)
  return(df)
}

vsd.hyp.train <- selectvsd("../results/06_hypallvsd.csv",  charHyp)
vsd.pit.train <- selectvsd("../results/06_pitallvsd.csv",  charPit)
vsd.gon.train <- selectvsd("../results/06_gonallvsd.csv",  charGon)

vsd.hyp.test <- selectvsd("../results/06_hypallvsd.csv",  manipHyp)
vsd.pit.test <- selectvsd("../results/06_pitallvsd.csv",  manipPit)
vsd.gon.test <- selectvsd("../results/06_gonallvsd.csv",  manipGon)

hypcols <- names(vsd.hyp.train)
pitcols <- names(vsd.pit.test)
goncols <- names(vsd.gon.test)

```

## combine stress and sex

```{r}
stress_countdata_hyp <- stress_countdata_hyp %>% tidyselect::vars_select(hypcols, strict = F)

head(stress_countdata_hyp)
stress_countdata_pit  
stress_countdata_pit  
```