### 2019-12-31 analysis/08_wgcna.Rmd
update wgcna markdown file

---
title: "wgcna"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = '../figures/wgcna/')
```

# WGCNA analysis, second pass

The anlayses and image was made by following along with and modifing the [first in a series of WGCNA tutorial](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/Consensus-DataInput.R). The first tutorial focuses on data input, cleaning and pre-processing; outlier removal; clustering of samples by their Euclidean distance; a comparison of sample cluster to sample meta data.

The second tutorial organized genes into modules

- "../analysis/08_WGCNA_1.R" 
- "../analysis/08_WGCNA_2.R"