### 2020-01-14 analysis/07_hormones.Rmd
new correlations

---
title: "hormones"
output: md_document
---

```{r setup}
library(tidyverse)
library(cowplot)
library(readxl)
library(modelr)
library(lubridate)
library(ggsignif)
library(apaTables)

source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")

knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE)
```

```{r}
colData  <- read_csv("../metadata/00_birds.csv") %>%
  mutate(RNAseq = "RNAseq",
         bird_id = bird)  %>%
   select(-X1, -bird)
colData
```

```{r wrangle-prolactin}
prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1) %>% 
  filter(Study %in% c("Baseline", "ParentalCare")) %>%
    dplyr::mutate(sex = fct_recode(Sex,
                            "female" = "f",
                            "male" = "m"),
           treatment = fct_recode(Treatment,
                            "hatch" = "Hatch",
                            "inc.d17" = "Inc_d17",
                            "inc.d17" = "inc_d17",
                            "inc.d3" = "Inc_d3",
                             "inc.d3" = "inc_d3",
                            "inc.d9" = "Inc_d9",
                             "inc.d9" = "inc_d9",
                            "m.inc.d9" = "M_Inc9",
                            "m.inc.d9" = "M_inc9",
                            "m.inc.d3" = "M_Inc3",
                            "m.inc.d8" = "M_Inc8",
                            "m.inc.d8" = "M_inc8",
                            "m.inc.d17" = "M_Inc17",
                            "m.n2" = "M_hatch",
                            "control" = "baseline",
                            "n5" = "N5", 
                            "n9" = "N9"),
           study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels1)) %>%
          dplyr::rename("plasma_conc" = "Prolactin ng/mL") %>%
          mutate(bird_id = gsub("[[:punct:]]", "." , ColorBands)) %>% 
          dplyr::mutate(hormone = "prolactin") %>% 
          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc)  %>% 
          drop_na()
head(prolactin)

prolactin2 <- read_csv("../results/prolactin.csv") %>% 
          dplyr::rename("plasma_conc" = "prolactin_ng_mL",
                        "treatment" = "stage") %>% 
           dplyr::mutate(hormone = "prolactin") %>% 
           dplyr::mutate(treatment = fct_recode(treatment,
                            "m.n2" = "m.hatch",
                            "prolong" = "inc.prolong",
                            "extend" = "extend.hatch"),
                          study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels1)) %>%

          dplyr::mutate(bird_id = gsub(".ATLAS", "" , bird_id)) %>% 
          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) 
head(prolactin2)

same <- inner_join(prolactin, prolactin2, by = "bird_id")
ggplot(same, aes(x = plasma_conc.x, y = plasma_conc.y, label = bird_id)) + geom_point() + geom_text()

all <- full_join(prolactin, prolactin2, by = "bird_id") 
onlyin1 <- anti_join(prolactin, prolactin2, by = "bird_id") 
onlyin2 <- anti_join(prolactin2, prolactin, by = "bird_id")

dim(prolactin)
dim(prolactin2)
dim(all)
dim(same)
dim(onlyin1)
dim(onlyin2)

print("only in old prolactin dataset") 
onlyin1$bird_id

print("only in new prolactin dataset") 
onlyin2$bird_id

prolctin1rnaseq <- full_join(prolactin, colData)
prolctin2rnaseq <- full_join(prolactin2, colData)

prolctin1rnaseq
prolctin2rnaseq
```

```{r wrangle-steroids}
PETC <- read_excel("../results/parental_care_hormone_RIA_data_master.xlsx", sheet = 2)  %>% 
                dplyr::mutate(treatment = fct_recode(stage,
                            "inc.d17" = "inc_d17",
                            "inc.d3" = "inc_d3",
                            "inc.d9" = "inc_d9",
                            "m.inc.d9" = "m_incd9",
                            "m.inc.d3" = "m_incd3",
                            "m.inc.d8" = "m_incd8",
                            "m.inc.d17" = "m_incd17",
                            "m.n2" = "m_hatch",
                            "control" = "stress_hpg")) %>%
                dplyr::filter(treatment %in% alllevels2) %>%   
                dplyr::mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
                       study = fct_collapse(treatment,
                                characterization = charlevels,
                                manipulation = maniplevels1)) %>% 
              dplyr::mutate(bird_id = gsub("[[:punct:]]", "." , band_combo)) %>% 
              dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) %>% 
             drop_na() %>%  droplevels()  
head(PETC)

```

```{r correlations}
PETCP <- left_join(prolactin2, PETC, by = "bird_id") %>% drop_na()
PETCP$treatment.x <- factor(PETCP$treatment.x, levels = alllevels2)

plothormonecorrelations <- function(myhormone, myylab){
  PETCP %>% filter(hormone.y == myhormone) %>%
  ggplot(aes(x = plasma_conc.x, y = plasma_conc.y, color = treatment.x )) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) +
  labs(x = "Prolactin (ng/mL)", y = myylab) +
  scale_color_manual(values = colorscharmaip) +
  facet_wrap(~treatment.x, nrow = 3) +
  theme(legend.position = "none")
}

plothormonecorrelations("cort", "Corticosterone (ng/mL")
plothormonecorrelations("estradiol", "Estradiol (ng/mL")
plothormonecorrelations("progesterone", "Progesterone (ng/mL")
plothormonecorrelations("testosterone", "Testosterone (ng/mL")
```

```{r combine-clean-split}
hormones <- rbind(prolactin, PETC)
hormones <- left_join(hormones, birds)

hormones$treatment <- factor(hormones$treatment, levels = alllevels)

hormones$okay <- ifelse(hormones$hormone == "cort" & hormones$plasma_conc > 30, "bad",
                    ifelse(hormones$hormone == "progesterone" & hormones$plasma_conc > 5, "bad", 
                           ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
                        ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay")))))
hormones <- hormones %>% filter(okay == "okay") %>% droplevels()
summary(hormones)
```

```{r characterizationplots}
hormonecharplot <- function(myhormone, myylab){

  hormones %>% 
    filter(study == "characterization",
           hormone %in% c(myhormone))  %>% 
    droplevels() %>% 
  ggplot(aes(x = as.numeric(treatment), y = plasma_conc)) +
        geom_smooth(aes(colour = sex)) +
    geom_boxplot(aes(fill = treatment, alpha = sex)) +
    mytheme() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          legend.position = "none") +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = myylab, x = NULL) +
    guides(fill = guide_legend(order=1),
         color = guide_legend(order=2)) +
    scale_alpha_manual(values = c(0.75,1)) +
    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
                       labels = charlevels)
  }

```

```{r PRLonly}
hormonecharplot("prolactin", "PRL (ng/mL)")
```

```{r manipulation}
# prolactin, removal only
hormones %>% 
    filter( hormone == c("prolactin"))  %>% 
    filter(treatment %in% c("inc.d3", "m.inc.d3", "inc.d9", "m.inc.d9", "inc.d17", "m.inc.d17", "hatch","m.n2"))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
    geom_boxplot() + 
  theme_B3()  +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = "PRL (ng/mL)", x = NULL) +
    ylim(c(-16,100)) +

   annotation_custom(inc, xmax = 1.6, ymin = -14, ymax = 5) +
   annotation_custom(removeegg, xmax = 3.6, ymin = -14, ymax = 5) +
   annotation_custom(inc, xmax = 5.6, ymin = -14, ymax = 5) +
   annotation_custom(removeegg, xmax = 7.6, ymin = -14, ymax = 5) +
  annotation_custom(hatch, xmax = 9.6, ymin = -14, ymax = 5) +
  annotation_custom(removeegg, xmax = 11.6, ymin = -14, ymax = 5) +
  annotation_custom(nestling, xmax = 13.6, ymin = -14, ymax = 5) +
  annotation_custom(removechick, xmax = 15.6, ymin = -14, ymax = 5) +

    annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -16, ymax = -14, alpha = 0.25) +
    annotate("rect", xmin = 2.6, xmax = 4.4, ymin = -16, ymax = -14, alpha = 0.5) +
    annotate("rect", xmin = 4.6, xmax = 6.4, ymin = -16, ymax = -14, alpha = 0.75)  +
    annotate("rect", xmin = 6.6, xmax = 8.4, ymin = -16, ymax = -14, alpha = 1)

hormones %>% 
    filter( hormone == c("prolactin"))  %>% 
    filter( treatment %in% c("inc.d9", "m.inc.d8", "inc.d17", "prolong", "hatch", "extend", "n5"))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
    geom_boxplot() + 
  theme_B3()  +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = "PRL (ng/mL)", x = NULL) +
  ylim(c(-12,100)) +
   annotation_custom(inc, xmax = 1.6, ymin = -10, ymax = 10) +
   annotation_custom(maniphatch, xmax = 3.6, ymin = -10, ymax = 10) +
   annotation_custom(inc, xmax = 5.6, ymin = -10, ymax = 10) +
   annotation_custom(manipinc, xmax = 7.6, ymin = -10, ymax = 10) +
  annotation_custom(hatch, xmax = 9.6, ymin = -10, ymax = 10) +
  annotation_custom(maniphatch, xmax = 11.6, ymin = -10, ymax = 10) +
  annotation_custom(nestling, xmax = 13.6, ymin = -10, ymax = 10) +

  annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -12, ymax = -10, alpha = 0.33) +
    annotate("rect", xmin = 4.6, xmax = 5.4, ymin = -12, ymax = -10, alpha = 0.33) +
    annotate("rect", xmin = 2.6, xmax = 5.4, ymin = -10, ymax = -8, alpha = 0.66) +
    annotate("rect", xmin = 4.6, xmax = 7.4, ymin = -8, ymax = -6, alpha = 1) 
```

```{r sexsteroids}

hormonecharplot("cort", "CORT (ng/mL)")
hormonecharplot("progesterone", "PROG (ng/mL)") 
d1 <- hormonecharplot("estradiol", "E (ng/mL)")
d2 <- hormonecharplot("testosterone", "T (ng/mL)")
plot_grid(d1,d2, nrow = 1)

hormonemanipSteroids <- function(myhormone, myylab, myymax){

  hormones %>% 
    filter( hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
    geom_boxplot() + 
    mytheme() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = myylab, x = NULL) 
}

hormonemanipSteroids("cort", "CORT (ng/mL)", 10)
hormonemanipSteroids("progesterone", "PROG (ng/mL)",2.5)

a <- hormonemanipSteroids("estradiol", "E (ng/mL)", 1)
b <- hormonemanipSteroids("testosterone", "T (ng/mL)", 3.5)
plot_grid(a,b)

```

```{r aov-char}
prl.char <- hormones %>% filter(hormone == "prolactin", treatment %in% charlevels)   %>%  droplevels()
test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
cort.char <- hormones %>% filter(hormone == "cort", treatment %in% charlevels)   %>%  droplevels()

aovSexTretment <- function(mydata, whichormone){
  aov2 <- aov(data = mydata, plasma_conc ~ treatment * sex)
  print(whichormone)
  print(summary(aov2))
  #print(TukeyHSD(aov2, which = "treatment"))
}

aovTretment  <- function(mydata, whichormone){
  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
  print(whichormone)
  print(summary(aov1))
  #print(TukeyHSD(aov1, which = "treatment"))
}

aovSexTretment(prl.char, "PRL") # yes, sex difference (p = 0.00256), yes treatment effect (p < 2e-16), no interaction
aovSexTretment(cort.char, "CORT") # no sex difference (p = 0.779 ), small treatment effect (p = 0.0238),  no interaction
aovSexTretment(prog.char, "PROG") # no sex difference or treatment effect, signifiant interacion (p = 0.0107)

aovTretment(est.char, "E") # p = 0.101
aovTretment(test.char, "T") # p = 0.609
```

# do control bird with high prolactin hormone have high PRL expression in the pituitary? yes.

```{r PRLvPRL}
PRLpit <- read_csv("../results/10_PRLpit.csv") %>% 
  filter(treatment == "control") %>% 
  arrange(desc(PRL))
head(PRLpit,2)
```

```{r}
write.csv(hormones, "../results/hormones.csv", row.names = F)
write.csv(prolactin2, "../results/07_prolactin2.csv", row.names = F)

```

## for rechelle

```{r forrechelle, fig.height= 4, fig.width=6}

rechelle <- hormones %>%
  dplyr::filter(hormone %in% c("prolactin", "cort"),
                treatment %in% c("lay", "inc.d9", "hatch", "n9")) %>%
  dplyr::select(-icons, -music, -iconpath, -okay)

rechelle$hormone <- factor(rechelle$hormone, levels = c("prolactin", "cort"))

prl.rechelle <- rechelle %>% filter(hormone == "prolactin")   %>%  droplevels()
cort.rechelle <- rechelle %>% filter(hormone == "cort")   %>%  droplevels()

TukeyHSD(aov(plasma_conc ~  treatment, data=prl.rechelle))
TukeyHSD(aov(plasma_conc ~  treatment, data=cort.rechelle))

rechelleplot <- function(mydf, myylab){

  p <- ggplot(mydf, aes(x = treatment, y = plasma_conc, fill = sex)) +
  geom_point(aes(color = sex), position = position_dodge(0.9), size = 2)+
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  scale_color_manual(name = "Sex", values = c("gray50", "maroon"))+
  scale_fill_manual(name = "Sex", values = c("gray50", "maroon"))+

  theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "none",
          strip.text = element_blank()) +
    theme_classic()+
  theme(text = element_text(size= 15)) +
    labs(x = "Stage", y = myylab) 
  return(p)

}

#PRL
rechelleplot(prl.rechelle, "Prolactin (ng/mL)")
rechelleplot(prl.rechelle, "Prolactin (ng/mL)") + 
  scale_y_continuous(breaks = c(0,25,50,75,100, 125), limits = c(-20,125)) +
  geom_signif(comparisons=list(c("lay", "hatch")), annotations= "p < 0.001", y_position = 100, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
  geom_signif(comparisons=list(c("lay", "n9")), annotations= "p < 0.001", y_position = 110, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
  geom_signif(comparisons=list(c("hatch", "n9")), annotations= "p < 0.001", y_position = 120, tip_length = 0.05, vjust = 0, color= "black" , textsize = 5 ) +
  geom_signif(comparisons=list(c("inc.d9", "hatch")), annotations= "p < 0.001", y_position = -5, tip_length = -0.05, vjust = 2, color= "black" , textsize = 5 ) +
  geom_signif(comparisons=list(c("inc.d9", "n9")), annotations= "p < 0.001", y_position = -15, tip_length = -0.05, vjust = 2, color= "black" , textsize = 5 )

#CORT
rechelleplot(cort.rechelle, "Corticosterone (ng/mL)")

```

```{r sex}
hormones$hormone <- factor(hormones$hormone, 
                           levels = c("cort", "progesterone", "prolactin", 
                                      "estradiol", "testosterone"))

hormones %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  facet_wrap(~hormone, nrow = 1, scales = "free") +
  theme(legend.position = "none") +
  labs(y = "concentration (ng/mL")

printaovtablesex <- function(myhormone, mydescirption){

  aov_all = data.frame()
  for(i in alllevels3){
    df <- hormones %>% filter(hormone == myhormone,
                            treatment == i) %>% droplevels()
    aovtable <- apa.aov.table(aov(plasma_conc ~ sex, data  = df))
    aovtable <- as.data.frame(aovtable$table_body)
    totaldf <- aovtable[3, 3]
    aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
    aovtable$ANOVA <- mydescirption
    aovtable$stages <- paste(i)
    aovtable$p <- as.numeric(as.character(aovtable$p))
    aov_all <- rbind(aov_all,aovtable)
  }

  aov_all <- aov_all %>%
    filter(Predictor == "sex")  %>%
    select(stages, ANOVA, df, "F", p) %>%
    mutate(sig = ifelse(p < 0.05, "*", " "))
  aov_all
  return(aov_all)
}

printaovtablesex("cort", "CORT ~ sex")
printaovtablesex("progesterone", "Prog ~ sex")
printaovtablesex("prolactin", "PRL ~ sex")

TukeyHSD(aov(data = cort.char, plasma_conc ~ treatment), which = "treatment")

```

```{r cort}

hormones %>%
  filter(hormone == "cort") %>%
  group_by(sex) %>%
  summarise(mean = mean(plasma_conc),
            median = median(plasma_conc))

p1 <- hormones %>%
  filter(hormone == "cort",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Corticosterone" , x = NULL) +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0,8.5))
p1

cort <- hormones %>%
  filter(hormone == "cort",
         study == "characterization")

p2 <-  ggplot(cort, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
  theme(legend.position = "none") +
  geom_signif(comparisons=list(c("control", "inc.d17")), annotations= "p = 0.49", y_position = 7.5, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
  geom_signif(comparisons=list(c("control", "hatch")), annotations= "p = 0.021", y_position = 8, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
  geom_signif(comparisons=list(c("control", "n5")), annotations= "p = 0.004", y_position = 8.5, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
  geom_signif(comparisons=list(c("n5", "n9")), annotations= "p = 0.001", y_position = 8, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
  scale_y_continuous(limits = c(0,8.5))

p2

p12 <- plot_grid(p1,p2, rel_widths = c(0.2,0.8), labels = c("a","b"), label_size = 8)

p3 <- hormones %>%
  filter(hormone == "cort",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs( x = NULL, y = "concentration (ng/mL")
p3

p4 <- hormones %>%
  filter(hormone == "cort",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs(y = NULL, x = NULL)
p4

p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4), labels = c("c","d"), label_size = 8 )

plot_grid(p12, p34, nrow = 2, rel_heights = c(0.6,0.4))
```

```{r PRL}
hormones %>%
  filter(hormone == "prolactin") %>%
  group_by(sex) %>%
  summarise(m = mean(plasma_conc),
            median = median(plasma_conc))

p1 <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Prolactin" , x = NULL) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,100)) 
p1

PRLf <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization",
         sex == "female")

PRLm <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization",
         sex == "male")

p2a <-  ggplot(PRLf, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = "females") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,100)) +
    geom_text(aes(label = "B", x = 1, y = 100, fontface = "plain"), size = 3) +
   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
geom_text(aes(label = "A", x = 3, y = 100), size = 3) +
geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
  geom_text(aes(label = "C", x = 5, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 9, y = 100), size = 3)

p2b <-  ggplot(PRLm, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = "males") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,100)) +
  geom_text(aes(label = "A", x = 1, y = 100, fontface = "plain"), size = 3) +
   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
geom_text(aes(label = "B", x = 3, y = 100), size = 3) +
geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
  geom_text(aes(label = "B", x = 5, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 9, y = 100), size = 3)

p12 <- plot_grid(p1,p2a,p2b, rel_widths = c(0.25,0.5, 0.5), labels = c("a","b", "c"), label_size = 8, nrow = 1, align = "hv")
p12

p3a <- hormones %>%
  filter(hormone == "prolactin",
         sex == "female", 
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( x = NULL, y = "concentration (ng/mL", subtitle = "females") +
  scale_color_manual(values = sexcolors)
p3a

p3b <- hormones %>%
  filter(hormone == "prolactin",
         sex == "male", 
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( x = NULL, y = NULL, subtitle = "males") +
  scale_color_manual(values = sexcolors)
p3b

p4a <- hormones %>%
  filter(hormone == "prolactin",
         sex == "female", 
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = NULL, x = NULL, subtitle = "females") +
  scale_color_manual(values = sexcolors)
p4a

p4b <- hormones %>%
  filter(hormone == "prolactin",
         sex == "male", 
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = NULL, x = NULL, subtitle = "males") +
  scale_color_manual(values = sexcolors)

p4b

p34 <- plot_grid(p3a , p3b,p4a, p4b, rel_widths = c(0.55,0.55, 0.45,0.45), labels = c("d","e", "f", "g"), label_size = 8, nrow = 1)

plot_grid(p12, p34, nrow = 2, rel_heights = c(0.5,0.5))
```

```{r correlations2 }
hormoneswide <- hormones %>%
  select(study, treatment, sex, bird_id, hormone, plasma_conc) %>%
  pivot_wider( names_from = hormone, values_from  = plasma_conc, values_fn = list(plasma_conc = mean))

plotcorrelations <- function(i,j, myxlab, mylab, mysubtitle){
      p <- ggplot(hormoneswide, aes(x = i, y = j),na.rm = TRUE) +
       geom_point(aes(color = treatment), na.rm = TRUE) +
       geom_smooth( method = lm, color = "darkgrey", na.rm = TRUE ) +
       theme_B3() +
      #facet_wrap(~sex, scales = "free") +
        labs(x = myxlab, y = mylab, subtitle = mysubtitle) +
        theme(legend.position = "none")

      return(p)
  }

p1 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$cort, "prolactin", "cort", "r = 0.221, p = 0.001")
p2 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$progesterone, "prolactin", "progesterone", "r = -0.099, p = 0.2")
p3 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$testosterone, "prolactin", "testosterone", "r = -0.258, p = 0.08")
p4 <- plotcorrelations(hormoneswide$prolactin, hormoneswide$estradiol, "prolactin", "estradiol", "r = 0.05, p = 0.6")

p5 <- plotcorrelations(hormoneswide$cort, hormoneswide$progesterone, "cort", "progesterone", "r = 0.104, p = 0.175")
p6 <- plotcorrelations(hormoneswide$cort, hormoneswide$estradiol, "cort", "estradiol", "r = -0.026, p = 0.8")
p7 <- plotcorrelations(hormoneswide$cort, hormoneswide$testosterone, "cort", "testosterone", "r = -0.108, p = 0.4")

p8 <- plotcorrelations(hormoneswide$progesterone, hormoneswide$estradiol, "progesterone", "estradiol", "r = 0.120, p = 0.3")
p9 <- plotcorrelations(hormoneswide$progesterone, hormoneswide$testosterone, "progesterone", "testosterone", "r = 0.133, p = 0.3")

allcorrs <- plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9, nrow = 3, align = "hv", labels = "auto", label_size = 8)
allcorrs

plot_grid(p1,p3 + theme(legend.position = "right"), align = "hv", rel_widths = c(0.4,0.6))

cor.test(hormoneswide$prolactin, hormoneswide$cort)
cor.test(hormoneswide$prolactin, hormoneswide$progesterone)
cor.test(hormoneswide$prolactin, hormoneswide$estradiol)
cor.test(hormoneswide$prolactin, hormoneswide$testosterone)

cor.test(hormoneswide$cort, hormoneswide$estradiol)
cor.test(hormoneswide$cort, hormoneswide$testosterone)
cor.test(hormoneswide$cort, hormoneswide$progesterone)

cor.test(hormoneswide$progesterone, hormoneswide$estradiol)
cor.test(hormoneswide$progesterone, hormoneswide$testosterone)
```