### 2020-01-23 analysis/22_symphony.Rmd
now with resepct project

---
title: "music"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)

source("../R/themes.R")
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/favegenes/")

```

## R Markdown

```{r respect}
song <- c("what", "you", "want", "ba-", "by", "I", "got.")
position <- c(1,2,3,4,5,6,7)
note1 <- c(11,9,7,8,8,8,6)
note2 <- c(7, 9,5,6,8,8,4)
note3 <- c(4, 9,2,4,8,8,1)

respect <- as.data.frame(t(rbind(position, note1,note2,note3)))
respect$song <- song
respect$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
                        size=7, replace = TRUE)
respect <- respect %>%
  pivot_longer(note1:note3, names_to = "music", values_to = "notes")  %>% 
  mutate(position = as.numeric(position),
         notes = as.numeric(notes))

song <- ggplot(respect, aes(x = position, y = notes)) +
  geom_line(aes(color = music)) +
  geom_image(aes(image=image), size = 0.15) +
  scale_x_continuous(breaks = position,
                     labels = song,
                     name = element_blank()) +
  scale_y_continuous(breaks = c(3, 5, 7, 9),
                     labels = c( "F",   "A",   "C",  "E"),
                     limits = c(-1,12)) +
  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
  theme_B3() +
  theme(axis.ticks = element_blank())

stage <- c( "bldg", "lay", "inc.d9", "inc.d17", "hatch", "n5", "n9")
time <- c(1,2,3,4,5,6,7)
gene1 <- c(11,9,7,8,8,8,6)
gene2 <- c(7, 9,5,6,8,8,4)
gene3 <- c(4, 9,2,4,8,8,1)

hypothesisgenes <- as.data.frame(t(rbind(time, gene1,gene2,gene3)))
hypothesisgenes$stage <- stage
hypothesisgenes$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
                        size=7, replace = TRUE)
hypothesisgenes <- hypothesisgenes %>%
  pivot_longer(gene1:gene3, names_to = "data", values_to = "expression")  %>% 
  mutate( time = as.numeric(time))

theory <- ggplot(hypothesisgenes, aes(x = time, y = expression)) +
  geom_line(aes(color = data)) +
  geom_image(aes(image=image), size = 0.15) +
  scale_x_continuous(breaks = time,
                     labels = stage,
                     name = element_blank()) +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10),
                     limits = c(-1,12)) +
  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15)) +
  theme_B3() +
  theme(axis.ticks = element_blank())

plot_grid(song , theory, nrow = 2)

hypothesisgenes2 <- hypothesisgenes %>% select(-image)

respect <- respect %>% select(-image)

write.csv(respect, "../results/22_respect.csv")
write.csv(hypothesisgenes2, "../results/22_respectgenes.csv" )
```