### 2020-01-23 analysis/22_symphony.Rmd
more symphony foo

---
title: "music"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)

source("../R/themes.R")
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/favegenes/")

```

## R Markdown

```{r respect}
song <- c("What", "you", "want,", "ba-", "by", "I", "got.")
stage <- c( "bldg", "lay", "inc.d9", "inc.d17", "hatch", "n5", "n9")
time <- c(1,2,3,4,5,6,7)
position <- c(1,2,3,4,5,6,7)
gene1 <- c(11,9,7,8,8,8,6)
gene2 <- c(7, 9,5,6,8,8,4)
gene3 <- c(4, 9,2,4,8,8,1)

respect <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
respect$song <- song
respect$stage <- stage
respect$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
                        size=7, replace = TRUE)
respect <- respect %>%
  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")  %>% 
  mutate(songstage = paste(song,stage, sep = "\n\n"))

songstageslim <- unique(respect$songstage)

respectsong <- ggplot(respect, aes(x = position, y = genes)) +
  geom_line(aes(color = music)) +
  geom_image(aes(image=image), size = 0.15) +
  scale_x_continuous(breaks = position,
                     labels = songstageslim,
                     name = element_blank()) +
  scale_y_continuous(breaks = c(3, 5, 7, 9),
                     labels = c( "F",   "A",   "C",  "E"),
                     limits = c(-1,12)) +
  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
  theme_B3() +
  theme(axis.ticks = element_blank()) +
  labs(subtitle = "RESPECT by Otis Redding, Jr.", y = NULL)
respectsong
```

```{r feelinggood}
song <- c("Birds", "fly-", "ing", "high,", "you", "know", "how", "I", "feel.")
stage <- charlevels
position <- c(1,2,3,4,5,6,7,8,9)
gene1 <- c(-2,0,1,2,2,1,1,-2,0)
gene2 <- gene1 - 1
gene1 <- gene2
gene3 <- gene2

feelinggood <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
feelinggood$song <- song
feelinggood$stage <- stage
feelinggood$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
                        size=9, replace = TRUE)
feelinggood <- feelinggood %>%
  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")   %>%
  mutate(songstage = paste(song,stage, sep = "\n\n"))
feelinggood

songstageslim <- unique(feelinggood$songstage)

feelinggoodsong <- ggplot(feelinggood, aes(x = position, y = genes)) +
  geom_line(aes(color = music)) +
  geom_image(aes(image=image), size = 0.15) +
  scale_x_continuous(breaks = position,
                    labels = songstageslim,
                     name = element_blank()) +
  scale_y_continuous(breaks = c(3, 5, 7, 9),
                     labels = c( "F",   "A",   "C",  "E"),
                     limits = c(-6,12)) +
  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
  theme_B3() +
  theme(axis.ticks = element_blank()) +
  labs(subtitle = "Feeling Good by L. Bricusse & A. Newley", y = NULL)
feelinggoodsong

```

```{r sinnerman}
song <- c("Oh,", "sin-", "ner", "man,,", "where", "_", "he", "gon'", "run", "_", "to?")
stage <- c( "bldg", "lay", "inc.d3", "m.inc.d3" ,
            "inc.d9", "m.inc.d9"  , 
            "inc.d17", "m.inc.d17" ,
            "hatch", "m.n2" ,
            "n5")
position <- c(1,2,3,4,5,6,7,8,9,10,11)
gene1 <- c(0,0,0,0,2,2,1,0,0,0,0)
gene2 <- gene1 + 2
gene3 <- gene2 + 2

sinnerman <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
sinnerman$song <- song
sinnerman$stage <- stage
sinnerman$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
                        size=11, replace = TRUE)
sinnerman <- sinnerman %>%
  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")   %>%
  mutate(songstage = paste(song,stage, sep = "\n\n"))
sinnerman

songstageslim <- unique(sinnerman$songstage)

sinnermansong <- ggplot(sinnerman, aes(x = position, y = genes)) +
  geom_line(aes(color = music)) +
  geom_image(aes(image=image), size = 0.1) +
  scale_x_continuous(breaks = position,
                    labels = songstageslim,
                     name = element_blank()) +
  scale_y_continuous(breaks = c(3, 5, 7, 9),
                     labels = c( "F",   "A",   "C",  "E"),
                     limits = c(-1,12)) +
  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
  theme_B3() +
  theme(axis.ticks = element_blank()) +
  labs(subtitle = "Sinnerman by Nina Simone", y = NULL)
sinnermansong

```
```{r soulmusic}
p1 <- plot_grid(respectsong + theme(legend.position = "none"), feelinggoodsong + theme(legend.position = "none"), rel_widths = c(0.45,0.55))

p3 <- plot_grid(p1, sinnermansong, nrow = 2)
p3
```

```{r symphonymodules}
modules <- read_csv("../results/08_genes_modules.csv") %>%
  rename(modulecolor = `net$colors`)

candidategenes <- c("PRL", "PRLR", 
                 "VIP", "VIPR1", "VIPR2", 
                 "OXT", "AVP", "AVPR1A", "AVPR1B", 
                 "GNRH1","GNRHR", "NPVF",
                 "NR3C1", "NR3C2",
                 "ESR1", "ESR2", "AR",
                 "DIO2","LEPR", "DIO3", "DIO1","CYP19A1",
                 "HSPA14", "HSPA12A",
                 "PTGES3", "HSD11B2","DRD5", "DRD1", "DRD2","PGE1", "PGF")

modulescandidates <- modules %>% filter(gene %in% candidategenes)  %>%
  group_by(modulecolor)  %>%
  summarize(gene = str_c(gene, collapse = ", "))

modulescandidates

datapath <- "../results/"   # path to the data
datafiles <- dir(datapath, pattern = "*allvsd.csv") # get file names
datapathfile <- paste0(datapath, datafiles)

df <- datapathfile %>%
  setNames(nm = .) %>%
  map_df(~read_csv(.x, col_types = cols(), col_names = T), .id = "filename") %>% 
  mutate(tissue = sapply(strsplit(as.character(filename),'../results/06_'), "[", 2)) %>% 
  mutate(tissue = sapply(strsplit(as.character(tissue),'allvsd.csv'), "[", 1))  %>% 
  select(tissue, X1, everything())

df2 <-  df  %>%
  filter(X1 %in% c(candidategenes)) %>%
  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "sample", values_to = "vsd") %>%
   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
  filter(treatment %in% charlevels) %>%
  select(bird, sex, treatment, tissue, X1, vsd) %>%
  mutate(tissue = fct_recode(tissue, "hypothalamus" = "hyp",
                    "pituitary" = "pit",
                    "gonads" = "gon"
                    )) %>%
  rename(gene = X1) %>%
  drop_na() %>%
  droplevels()
head(df2)

df2$treatment <- factor(df2$treatment, levels = alllevels)
df2$tissue <- factor(df2$tissue, levels = tissuelevels)
df2$gene <- factor(df2$gene)

df3 <- df2 %>% 
  mutate(treatment = fct_relevel(treatment, charlevels)) %>% 
  group_by(treatment, tissue, gene)  %>% 
  summarize(m = mean(vsd, na.rm = T), se = sd(vsd,  na.rm = T)/sqrt(length(vsd))) %>%
  mutate(image = "../figures/images/DoveParentsRNAseq_note.png") 
head(df3)

d4 <- left_join(df3, modules, by = "gene")

for (i in levels(d4$tissue)) {
  p <-  d4 %>%
    filter(tissue == i) %>%
    ggplot(aes(x = treatment, y = m)) +
    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
    geom_point(size = 1, aes(color = gene)) +
    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = i, y = "average expression", x = "parental stage") +
    facet_wrap(~modulecolor, nrow = 2) +
    theme_B3() +
    theme(legend.position = "bottom")
 print(p)
}

d4 %>%
  filter(modulecolor != "turquoise",
         tissue == "pituitary") %>%
  droplevels() %>% 
    ggplot(aes(x = treatment, y = m)) +
    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
    geom_point(size = 1, aes(color = gene)) +
    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "average expression", x = "parental stage") +
    facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
    theme_B3() +
    theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
    guides(color = guide_legend(nrow = 2, byrow = T))

d4 %>%
  filter( modulecolor != "turquoise",
          # modulecolor != "greenyellow",
          tissue == "pituitary") %>%
  droplevels() %>% 
    ggplot(aes(x = treatment, y = m)) +
  geom_image(aes(image=image), size = 0.15) +
  facet_wrap(~modulecolor, scales = "free_y", nrow = 2) +
  labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "gene expression", x = "parental stage") +
    facet_wrap(~gene, scales = "free_y", nrow = ) +
    theme_B3() +
    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))

d4 %>%
  filter( modulecolor != "turquoise",
          # modulecolor != "greenyellow",
          tissue == "pituitary") %>%
  droplevels() %>% 
    ggplot(aes(x = treatment, y = m)) +
  geom_image(aes(image=image), size = 0.1) +
  labs(subtitle = "WGCNA + canddiate genes in the pituitary", y = "gene expression", x = "parental stage") +
    theme_B3() +
    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_log10()

```

## outline

- data-driven
- hypothesis-drive
- sonically-driven
- visually-driven
- statistcally-driven

Data and hypothesis driven research for the visually, sonically, and statistcially inclined biologists