### 2020-02-03 analysis/07_hormones.Rmd
now with dense figures

---
title: "hormones"
output: md_document
---

```{r setup}
library(tidyverse)
library(cowplot)
library(readxl)
library(modelr)
library(lubridate)
library(ggsignif)
library(apaTables)

source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")

knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE, cache = T)
```

```{r}
colData  <- read_csv("../metadata/00_birds.csv") %>%
  mutate(RNAseq = "RNAseq",
         bird_id = bird)  %>%
   select(-X1, -bird)
colData
```

```{r wrangle-prolactin}
prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1) %>% 
  filter(Study %in% c("Baseline", "ParentalCare")) %>%
    dplyr::mutate(sex = fct_recode(Sex,
                            "female" = "f",
                            "male" = "m"),
           treatment = fct_recode(Treatment,
                            "hatch" = "Hatch",
                            "inc.d17" = "Inc_d17",
                            "inc.d17" = "inc_d17",
                            "inc.d3" = "Inc_d3",
                             "inc.d3" = "inc_d3",
                            "inc.d9" = "Inc_d9",
                             "inc.d9" = "inc_d9",
                            "m.inc.d9" = "M_Inc9",
                            "m.inc.d9" = "M_inc9",
                            "m.inc.d3" = "M_Inc3",
                            "m.inc.d8" = "M_Inc8",
                            "m.inc.d8" = "M_inc8",
                            "m.inc.d17" = "M_Inc17",
                            "m.n2" = "M_hatch",
                            "control" = "baseline",
                            "n5" = "N5", 
                            "n9" = "N9"),
           study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels1)) %>%
          dplyr::rename("plasma_conc" = "Prolactin ng/mL") %>%
          mutate(bird_id = gsub("[[:punct:]]", "." , ColorBands)) %>% 
          dplyr::mutate(hormone = "prolactin") %>% 
          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc)  %>% 
          drop_na()
head(prolactin)

```

```{r wrangle-steroids}
PETC <- read_excel("../results/hormones.xlsx", sheet = 1)  %>% 
              dplyr::rename(corticosterone = cort, progesterone = p4,
                            estradiol = e2,  testosterone = t) %>% 
               dplyr::mutate(treatment = fct_recode(treatment...3,
                            "inc.d17" = "incd17",
                            "inc.d3" = "incd3",
                            "inc.d9" = "incd9",
                            "m.inc.d9" = "minc9",
                            "m.inc.d3" = "minc3",
                            "m.inc.d8" = "minc8",
                            "m.inc.d17" = "minc17",
                            "m.n2" = "m hatch",
                            "control" = "baseline")) %>%
                dplyr::filter(treatment %in% alllevels2) %>%   
                dplyr::mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
                       study = fct_collapse(treatment,
                                characterization = charlevels,
                                manipulation = maniplevels1)) %>% 
              dplyr::mutate(bird_id = gsub("[[:punct:]]", "." , id)) %>% 
              dplyr::select(bird_id, treatment, sex, study, 
                            corticosterone, estradiol, testosterone, progesterone) %>% 
              pivot_longer(cols = corticosterone:progesterone,
                           names_to = "hormone", values_to = "plasma_conc",
                           values_drop_na = TRUE) %>% 
              drop_na() %>% droplevels()  
head(PETC)
```

```{r combine-clean-split}
hormones <- rbind(prolactin, PETC)

hormones$okay <- ifelse(hormones$hormone == "corticosterone" & hormones$plasma_conc > 30, "bad",
                    ifelse(hormones$hormone == "progesterone" & hormones$plasma_conc > 5, "bad", 
                           ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
                        ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay")))))
hormones <- hormones %>% filter(okay == "okay") %>% droplevels() %>% select(-okay)

write.csv(hormones, "../results/hormones.csv", row.names = F)

# make a winder one for correlations
hormoneswide <- hormones %>% pivot_wider(names_from = "hormone", values_from = "plasma_conc", 
                                         values_fn = list(plasma_conc = mean))
hormoneswide$treatment <- factor(hormoneswide$treatment, levels = alllevels)
head(hormoneswide)

# for facet wrapping
hormones <- left_join(hormones, birds)

hormones$treatment <- factor(hormones$treatment, levels = alllevels)

hormones$hormone <- factor(hormones$hormone, 
                           levels = c("prolactin", "corticosterone", "progesterone", 
                                      "estradiol", "testosterone"))
head(hormones)
```

```{r characterizationplots}
hormonecharplot <- function(myhormone, myylab){

  hormones %>% 
    filter(study == "characterization",
           hormone %in% c(myhormone))  %>% 
    droplevels() %>% 
  ggplot(aes(x = as.numeric(treatment), y = plasma_conc)) +
        geom_smooth(aes(colour = sex)) +
    geom_boxplot(aes(fill = treatment)) +
    mytheme() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          legend.position = "none") +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = myylab, x = NULL) +
    guides(fill = guide_legend(order=1),
         color = guide_legend(order=2)) +
    scale_x_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
                       labels = charlevels)
  }

hormonecharplot("prolactin", "PRL (ng/mL)")
hormonecharplot("corticosterone", "CORT (ng/mL)")
hormonecharplot("estradiol", "e2 (ng/mL)")
hormonecharplot("testosterone", "t (ng/mL)")
hormonecharplot("progesterone", "p4 (ng/mL)")
```

```{r manipulation}
# prolactin, removal only
hormones %>% 
    filter( hormone == c("prolactin"))  %>% 
    filter(treatment %in% c("inc.d3", "m.inc.d3", "inc.d9", "m.inc.d9", "inc.d17", "m.inc.d17", "hatch","m.n2"))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
    geom_boxplot() + 
  theme_B3()  +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = "PRL (ng/mL)", x = NULL) +
    ylim(c(-16,100)) +

   annotation_custom(inc, xmax = 1.6, ymin = -14, ymax = 5) +
   annotation_custom(removeegg, xmax = 3.6, ymin = -14, ymax = 5) +
   annotation_custom(inc, xmax = 5.6, ymin = -14, ymax = 5) +
   annotation_custom(removeegg, xmax = 7.6, ymin = -14, ymax = 5) +
  annotation_custom(hatch, xmax = 9.6, ymin = -14, ymax = 5) +
  annotation_custom(removeegg, xmax = 11.6, ymin = -14, ymax = 5) +
  annotation_custom(nestling, xmax = 13.6, ymin = -14, ymax = 5) +
  annotation_custom(removechick, xmax = 15.6, ymin = -14, ymax = 5) +

    annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -16, ymax = -14, alpha = 0.25) +
    annotate("rect", xmin = 2.6, xmax = 4.4, ymin = -16, ymax = -14, alpha = 0.5) +
    annotate("rect", xmin = 4.6, xmax = 6.4, ymin = -16, ymax = -14, alpha = 0.75)  +
    annotate("rect", xmin = 6.6, xmax = 8.4, ymin = -16, ymax = -14, alpha = 1)

hormones %>% 
    filter( hormone == c("prolactin"))  %>% 
    filter( treatment %in% c("inc.d9", "m.inc.d8", "inc.d17", "prolong", "hatch", "extend", "n5"))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment,  color = sex)) +
    geom_boxplot() + 
  theme_B3()  +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = "PRL (ng/mL)", x = NULL) +
  ylim(c(-12,100)) +
   annotation_custom(inc, xmax = 1.6, ymin = -10, ymax = 10) +
   annotation_custom(maniphatch, xmax = 3.6, ymin = -10, ymax = 10) +
   annotation_custom(inc, xmax = 5.6, ymin = -10, ymax = 10) +
   annotation_custom(manipinc, xmax = 7.6, ymin = -10, ymax = 10) +
  annotation_custom(hatch, xmax = 9.6, ymin = -10, ymax = 10) +
  annotation_custom(maniphatch, xmax = 11.6, ymin = -10, ymax = 10) +
  annotation_custom(nestling, xmax = 13.6, ymin = -10, ymax = 10) +

  annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -12, ymax = -10, alpha = 0.33) +
    annotate("rect", xmin = 4.6, xmax = 5.4, ymin = -12, ymax = -10, alpha = 0.33) +
    annotate("rect", xmin = 2.6, xmax = 5.4, ymin = -10, ymax = -8, alpha = 0.66) +
    annotate("rect", xmin = 4.6, xmax = 7.4, ymin = -8, ymax = -6, alpha = 1) 
```

```{r sexsteroids}

hormonecharplot("corticosterone", "CORT (ng/mL)")
hormonecharplot("progesterone", "PROG (ng/mL)") 
d1 <- hormonecharplot("estradiol", "E (ng/mL)")
d2 <- hormonecharplot("testosterone", "T (ng/mL)")
plot_grid(d1,d2, nrow = 1)

hormonemanipSteroids <- function(myhormone, myylab, myymax){

  hormones %>% 
    filter( hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
    geom_boxplot() + 
    mytheme() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
         legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = colorscharmaip) +
    scale_color_manual(values = sexcolors) +
    labs(y = myylab, x = NULL) 
}

hormonemanipSteroids("corticosterone", "CORT (ng/mL)", 10)
hormonemanipSteroids("progesterone", "PROG (ng/mL)",2.5)

a <- hormonemanipSteroids("estradiol", "E (ng/mL)", 1)
b <- hormonemanipSteroids("testosterone", "T (ng/mL)", 3.5)
plot_grid(a,b)

```

```{r aov-char}
prl.char <- hormones %>% filter(hormone == "prolactin", treatment %in% charlevels)   %>%  droplevels()
test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
cort.char <- hormones %>% filter(hormone == "corticosterone", treatment %in% charlevels)   %>%  droplevels()

aovSexTretment <- function(mydata, whichormone){
  aov2 <- aov(data = mydata, plasma_conc ~ treatment * sex)
  print(whichormone)
  print(summary(aov2))
  #print(TukeyHSD(aov2, which = "treatment"))
}

aovTretment  <- function(mydata, whichormone){
  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
  print(whichormone)
  print(summary(aov1))
  #print(TukeyHSD(aov1, which = "treatment"))
}

aovSexTretment(prl.char, "PRL") # yes, sex difference (p = 0.00256), yes treatment effect (p < 2e-16), no interaction
aovSexTretment(cort.char, "corticosterone") # no sex difference (p = 0.779 ), small treatment effect (p = 0.0238),  no interaction
aovSexTretment(prog.char, "PROG") # no sex difference or treatment effect, signifiant interacion (p = 0.0107)

aovTretment(est.char, "E") # p = 0.101
aovTretment(test.char, "T") # p = 0.609
```

# do control bird with high prolactin hormone have high PRL expression in the pituitary? yes.

```{r PRLvPRL}
PRLpit <- read_csv("../results/10_PRLpit.csv") %>% 
  filter(treatment == "control") %>% 
  arrange(desc(PRL))
head(PRLpit,2)
```

```{r sex}
hormones %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  facet_wrap(~hormone, nrow = 1, scales = "free") +
  theme(legend.position = "none") +
  labs(y = "concentration (ng/mL")

printaovtablesex <- function(myhormone, mydescirption){

  aov_all = data.frame()
  for(i in alllevels){
    df <- hormones %>% filter(hormone == myhormone,
                            treatment == i) %>% droplevels()
    aovtable <- apa.aov.table(aov(plasma_conc ~ sex, data  = df))
    aovtable <- as.data.frame(aovtable$table_body)
    totaldf <- aovtable[3, 3]
    aovtable$df <- paste(aovtable$df, ", " , totaldf, sep = "")
    aovtable$ANOVA <- mydescirption
    aovtable$stages <- paste(i)
    aovtable$p <- as.numeric(as.character(aovtable$p))
    aov_all <- rbind(aov_all,aovtable)
  }

  aov_all <- aov_all %>%
    filter(Predictor == "sex")  %>%
    select(stages, ANOVA, df, "F", p) %>%
    mutate(sig = ifelse(p < 0.05, "*", " "))
  aov_all
  return(aov_all)
}

printaovtablesex("prolactin", "PRL ~ sex")
```

```{r cort}

hormones %>%
  filter(hormone == "corticosterone") %>%
  group_by(sex) %>%
  summarise(mean = mean(plasma_conc),
            median = median(plasma_conc))

p1 <- hormones %>%
  filter(hormone == "corticosterone",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Corticosterone" , x = NULL) +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = sexcolors)
p1

cort <- hormones %>%
  filter(hormone == "corticosterone",
         study == "characterization")

p2 <-  ggplot(cort, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
  theme(legend.position = "none") +
  geom_signif(comparisons=list(c("control", "inc.d17")), annotations= "p = 0.49", y_position = 7.5, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
  geom_signif(comparisons=list(c("control", "hatch")), annotations= "p = 0.021", y_position = 8, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
  geom_signif(comparisons=list(c("control", "n5")), annotations= "p = 0.004", y_position = 8.5, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
  geom_signif(comparisons=list(c("n5", "n9")), annotations= "p = 0.001", y_position = 8, tip_length = 0, vjust = 0, color= "black" , textsize = 3 ) +
  scale_y_continuous(limits = c(0,8.5))

p2

p12 <- plot_grid(p1,p2, rel_widths = c(0.2,0.8), labels = c("a","b"), label_size = 8)

p3 <- hormones %>%
  filter(hormone == "corticosterone",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs( x = NULL, y = "concentration (ng/mL")
p3

p4 <- hormones %>%
  filter(hormone == "corticosterone",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs(y = NULL, x = NULL)
p4

p34 <- plot_grid(p3,p4, rel_widths = c(0.6,0.4), labels = c("c","d"), label_size = 8 )
p34

allcortplots <- plot_grid(p12, p34, nrow = 2, rel_heights = c(0.6,0.4))
allcortplots
```

```{r PRL}
hormones %>%
  filter(hormone == "prolactin") %>%
  group_by(sex) %>%
  summarise(m = mean(plasma_conc),
            median = median(plasma_conc))

p1 <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Prolactin" , x = NULL) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,100)) 
p1

PRLf <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization",
         sex == "female")

PRLm <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization",
         sex == "male")

p2a <-  ggplot(PRLf, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = "females") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,100)) +
    geom_text(aes(label = "B", x = 1, y = 100, fontface = "plain"), size = 3) +
   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
geom_text(aes(label = "A", x = 3, y = 100), size = 3) +
geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
  geom_text(aes(label = "C", x = 5, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 9, y = 100), size = 3)

p2b <-  ggplot(PRLm, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = "males") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,100)) +
  geom_text(aes(label = "A", x = 1, y = 100, fontface = "plain"), size = 3) +
   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
geom_text(aes(label = "B", x = 3, y = 100), size = 3) +
geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
  geom_text(aes(label = "B", x = 5, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 9, y = 100), size = 3)

p12 <- plot_grid(p1,p2a,p2b, rel_widths = c(0.25,0.5, 0.5), labels = c("a","b", "c"), label_size = 8, nrow = 1, align = "hv")
p12

p3a <- hormones %>%
  filter(hormone == "prolactin",
         sex == "female", 
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( x = NULL, y = "concentration (ng/mL", subtitle = "females") +
  scale_color_manual(values = sexcolors)
p3a

p3b <- hormones %>%
  filter(hormone == "prolactin",
         sex == "male", 
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( x = NULL, y = NULL, subtitle = "males") +
  scale_color_manual(values = sexcolors)
p3b

p4a <- hormones %>%
  filter(hormone == "prolactin",
         sex == "female", 
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = NULL, x = NULL, subtitle = "females") +
  scale_color_manual(values = sexcolors)
p4a

p4b <- hormones %>%
  filter(hormone == "prolactin",
         sex == "male", 
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,100)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = NULL, x = NULL, subtitle = "males") +
  scale_color_manual(values = sexcolors)

p4b

p34 <- plot_grid(p3a , p3b,p4a, p4b, rel_widths = c(0.55,0.55, 0.45,0.45), labels = c("d","e", "f", "g"), label_size = 8, nrow = 1)

allprolactinplots <- plot_grid(p12, p34, nrow = 2, rel_heights = c(0.5,0.5))
allprolactinplots
```

## figure 2

```{r figure2, fig.width=7, fig.height=7}

plot_grid(allprolactinplots, allcortplots, nrow = 2)
```

## prolactin 2

```{r PRL2}
hormones %>%
  filter(hormone == "prolactin") %>%
  group_by(sex) %>%
  summarise(m = mean(plasma_conc),
            median = median(plasma_conc))

p1 <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Prolactin" , x = NULL) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,100)) 
p1

df <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization")
p2 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,105)) +
    geom_text(aes(label = "B", x = 1, y = 100, fontface = "plain"), size = 3) +
   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
geom_text(aes(label = "A", x = 3, y = 100), size = 3) +
geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
  geom_text(aes(label = "C", x = 5, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 9, y = 100), size = 3) +
  geom_text(aes(label = '*', x = 1, y = 105), size = 5) +
   geom_text(aes(label = '*', x = 2, y = 105), size = 5) 
  p2

p3 <- hormones %>%
  filter(hormone == "prolactin",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,110)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( x = NULL, y = "concentration (ng/mL", subtitle = " ") +
  scale_color_manual(values = sexcolors) +
   geom_text(aes(label = '*', x = 2, y = 105), size = 5) 
p3

p4 <- hormones %>%
  filter(hormone == "prolactin",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,110)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = NULL, x = NULL, subtitle = " ") +
  scale_color_manual(values = sexcolors) 
p4

p12 <- plot_grid(p1, p2, rel_widths = c(0.15,0.8), labels = "auto", label_size = 8)
p34 <- plot_grid(p3,p4, nrow = 1, rel_widths = c(0.55,0.45), labels = c("c", "d"), label_size = 8)

plot_grid(p12, p34, nrow = 2)
```

## correlations

```{r correlations2 }

plotcorrelations <- function(i,j, myxlab, mylab, mysubtitle){
      p <- ggplot(hormoneswide, aes(x = i, y = j), na.rm = TRUE) +
       geom_point(aes(color = treatment), na.rm = TRUE) +
       geom_smooth( method = lm, color = "darkgrey", na.rm = TRUE ) +
       theme_B3() +
      #facet_wrap(~sex, scales = "free") +
        labs(x = myxlab, y = mylab, subtitle = mysubtitle) +
        theme(legend.position = "none") +
        scale_color_manual(values = colorscharmaip2)

      return(p)
}

hormoneswide

corrPRL <- hormoneswide %>%
  pivot_longer(cols = corticosterone:progesterone, names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = prolactin, y = conc)) +
  geom_point(aes(color = treatment)) +
  geom_smooth( color = "darkgrey", method = "lm") +
  facet_wrap(~hormones, nrow = 1, scales = "free_y")  +
  theme_B3() +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_color_manual(values = colorscharmaip2) +
  guides(color = guide_legend(nrow = 1)) +
  labs(y = "conc. (ng/mL)", x = "prolactin (ng/mL)")
corrPRL

corrCORT <- hormoneswide %>%
  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = corticosterone, y = conc)) +
  geom_point(aes(color = treatment)) +
  geom_smooth( color = "darkgrey", method = "lm") +
  facet_wrap(~hormones, nrow = 1, scales = "free_y")  +
  theme_B3() +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_color_manual(values = colorscharmaip2) +
  guides(color = guide_legend(nrow = 1)) +
  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")
corrCORT

hormoneswide %>%
  pivot_longer(cols = c(prolactin, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = corticosterone, y = conc)) +
  geom_point(aes(color = treatment)) +
  geom_smooth(color = "darkgrey", method = "lm") +
  facet_grid(hormones ~ treatment, scales = "free_y")  +
  theme_B3() +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_color_manual(values = colorscharmaip2) +
  guides(color = guide_legend(nrow = 1)) +
  labs(y = "conc. (ng/mL)", x = "corticosterone (ng/mL)")

hormoneswide %>%
  pivot_longer(cols = c(corticosterone, estradiol, testosterone, progesterone), names_to = "hormones", values_to = "conc", values_drop_na = TRUE) %>%
  filter(study == "characterization") %>%
  ggplot(aes(x = prolactin, y = conc)) +
  geom_point(aes(color = treatment)) +
  geom_smooth(color = "darkgrey", method = "lm") +
  facet_grid(hormones ~ treatment, scales = "free_y")  +
  theme_B3() +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  scale_color_manual(values = colorscharmaip2) +
  guides(color = guide_legend(nrow = 1)) +
  labs(y = "conc. (ng/mL)", x = "prolactin (ng/mL)")

ggplot(hormoneswide, aes(x = prolactin, y = corticosterone, color = treatment)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_wrap(~treatment) +
  scale_color_manual(values =  colorscharmaip) + theme(legend.position = "none")

ggplot(hormoneswide, aes(x = prolactin, y = testosterone, color = treatment)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_wrap(~treatment) +
  scale_color_manual(values =  colorscharmaip) + theme(legend.position = "none")

```

## progesterone

```{r progesterone}

hormones %>%
  filter(hormone == "progesterone") %>%
  group_by(sex) %>%
  summarise(mean = mean(plasma_conc),
            median = median(plasma_conc))

p1 <- hormones %>%
  filter(hormone == "progesterone",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Progesterone" , x = NULL) +
  theme(legend.position = "none")   +
  scale_y_continuous(limits = c(0,4.5)) +
  scale_fill_manual(values = sexcolors)

df <- hormones %>%
  filter(hormone == "progesterone",
         study == "characterization")

p2 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
  scale_y_continuous(limits = c(0,4.5)) +
  theme(legend.position = "none")

p3 <- hormones %>%
  filter(hormone == "progesterone",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,4.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs( x = NULL, y = "concentration (ng/mL")

p4 <- hormones %>%
  filter(hormone == "progesterone",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,4.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs(y = NULL, x = NULL)

progesterone <- plot_grid(p1, p2, p3 + labs(y = NULL), p4, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
progesterone
```

## estradiol and testosterone

```{r estradiol}

hormones %>%
  filter(hormone == "estradiol") %>%
  summarise(mean = mean(plasma_conc),
            median = median(plasma_conc))

p5 <- hormones %>%
  filter(hormone == "estradiol",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Estradiol" , x = NULL) +
  theme(legend.position = "none")  +
    scale_y_continuous(limits = c(0,1)) +
  scale_fill_manual(values = sexcolors)

df <- hormones %>%
  filter(hormone == "estradiol",
         study == "characterization")

p6 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
    scale_y_continuous(limits = c(0,1)) +
  theme(legend.position = "none")

p7 <- hormones %>%
  filter(hormone == "estradiol",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,1)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs( x = NULL, y = "concentration (ng/mL")

p8 <- hormones %>%
  filter(hormone == "estradiol",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,1)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs(y = NULL, x = NULL)

estradiol <- plot_grid(p5, p6, p7 + labs(y = NULL), p8, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
estradiol
```

```{r testosterone}

hormones %>%
  filter(hormone == "testosterone") %>%
  summarise(mean = mean(plasma_conc),
            median = median(plasma_conc))

p9 <- hormones %>%
  filter(hormone == "testosterone",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Testosterone" , x = NULL) +
  theme(legend.position = "none")  +
    scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = sexcolors)

df <- hormones %>%
  filter(hormone == "testosterone",
         study == "characterization")

p10 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
    scale_y_continuous(limits = c(0,8.5)) +
  theme(legend.position = "none")

p11 <- hormones %>%
  filter(hormone == "testosterone",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs( x = NULL, y = "concentration (ng/mL")

p12 <- hormones %>%
  filter(hormone == "testosterone",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs(y = NULL, x = NULL)
p12

testosterone <- plot_grid(p9, p10, p11 + labs(y = NULL), p12, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
testosterone
```

```{r supplefig, fig.height=8}
sexsteroids <- plot_grid(p1, p2, p3 + labs(y = NULL), p4, p5, p6, p7 + labs(y = NULL), p8, 
          p9, p10, p11 + labs(y = NULL), p12, nrow = 3, rel_widths =   c(0.175,0.4,0.35,0.25), align = "hv",
          label_size = 8, labels = c("a", " ", " " , " ",
                                     "b", " ", " " , " ",
                                     "c", " ", " " , " "))
sexsteroids
```

## pca of hormones

```{r PCA}

library("factoextra")
library("FactoMineR")

# char only
hormoneswidechar <- hormoneswide %>% filter(treatment %in% charlevels)
df <- hormoneswidechar %>% select(prolactin:progesterone)
res.pca <- PCA(df)

a1 <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization only") 
# Contributions of variables to PC1

b1 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
                                                                      title = element_blank())
# Contributions of variables to PC2
c1 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
  labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())

ones <- plot_grid(a1,b1,c1, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")

# char and manip
df <- hormoneswide %>% select(prolactin:progesterone)
res.pca <- PCA(df)

a2 <- fviz_pca_var(res.pca, col.var = "black", title = "Charactherization & manipluation") 
# Contributions of variables to PC1

b2 <- fviz_contrib(res.pca, choice = "var", axes = 1, top = 10) + labs(y = "Dim 1 contibutions (%)",
                                                                      title = element_blank())
# Contributions of variables to PC2
c2 <- fviz_contrib(res.pca, choice = "var", axes = 2, top = 10) +
  labs(y = "Dim 2 contibutions (%)") + theme(title = element_blank())

twos <- plot_grid(a2 + theme(title = element_blank()),b2,c2, nrow = 1, rel_widths = c(1,0.425,0.425), align = "hv")

plot_grid(ones,twos, nrow = 2)

twos
```

```{r figure3, fig.height=8, fig.width=7}

plot_grid(twos,sexsteroids, nrow = 2, rel_heights = c(0.25,0.75), align = "hv")
```

# smooth funciton

```{r summary}
head(hormones)
ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc, color = hormone)) +
  geom_smooth()

box <- ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc)) +
  geom_boxplot(aes(fill = treatment)) +
  #geom_smooth(aes(color = hormone)) +
  facet_wrap(~hormone, scales = "free_y", nrow = 5) +
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
                     labels = alllevels) +
  theme_B3() +
  theme(#strip.text = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)) +
  labs(y = "plasma conc. (ng/mL)", x = "parental stage")

line <- ggplot(hormones, aes(x = as.numeric(treatment), y = plasma_conc)) +
  #geom_boxplot(aes(fill = treatment)) +
  geom_smooth(aes(color = hormone)) +
  facet_wrap(~hormone, scales = "free_y", nrow = 5) +
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16),
                     labels = alllevels) +
  theme_B3() +
  theme(#strip.text = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none",
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)) +
  labs(y = "plasma conc. (ng/mL)", x = "parental stage")

plot_grid(box, line, nrow = 1)
```