### 2020-02-05 analysis/07_hormones.Rmd
now with only datat wrangling

---
title: "hormones"
output: md_document
---

```{r setup}
library(tidyverse)
library(cowplot)
library(readxl)
library(modelr)
library(lubridate)
library(ggsignif)
library(apaTables)

source("../R/themes.R")  # load custom themes and color palletes
source("../R/icons.R")

knitr::opts_chunk$set(fig.path = '../figures/hormones/',message=F, warning=FALSE, cache = T)
```

```{r}
colData  <- read_csv("../metadata/00_birds.csv") %>%
  mutate(RNAseq = "RNAseq",
         bird_id = bird)  %>%
   select(-X1, -bird)
colData
```

```{r wrangle-prolactin}
prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1) %>% 
  filter(Study %in% c("Baseline", "ParentalCare")) %>%
    dplyr::mutate(sex = fct_recode(Sex,
                            "female" = "f",
                            "male" = "m"),
           treatment = fct_recode(Treatment,
                            "hatch" = "Hatch",
                            "inc.d17" = "Inc_d17",
                            "inc.d17" = "inc_d17",
                            "inc.d3" = "Inc_d3",
                             "inc.d3" = "inc_d3",
                            "inc.d9" = "Inc_d9",
                             "inc.d9" = "inc_d9",
                            "m.inc.d9" = "M_Inc9",
                            "m.inc.d9" = "M_inc9",
                            "m.inc.d3" = "M_Inc3",
                            "m.inc.d8" = "M_Inc8",
                            "m.inc.d8" = "M_inc8",
                            "m.inc.d17" = "M_Inc17",
                            "m.n2" = "M_hatch",
                            "control" = "baseline",
                            "n5" = "N5", 
                            "n9" = "N9"),
           study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels1)) %>%
          dplyr::rename("plasma_conc" = "Prolactin ng/mL") %>%
          mutate(bird_id = gsub("[[:punct:]]", "." , ColorBands)) %>% 
          dplyr::mutate(hormone = "prolactin") %>% 
          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc)  %>% 
          drop_na()
head(prolactin)

```

```{r wrangle-steroids}
PETC <- read_excel("../results/hormones.xlsx", sheet = 1)  %>% 
              dplyr::rename(corticosterone = cort, progesterone = p4,
                            estradiol = e2,  testosterone = t) %>% 
               dplyr::mutate(treatment = fct_recode(treatment...3,
                            "inc.d17" = "incd17",
                            "inc.d3" = "incd3",
                            "inc.d9" = "incd9",
                            "m.inc.d9" = "minc9",
                            "m.inc.d3" = "minc3",
                            "m.inc.d8" = "minc8",
                            "m.inc.d17" = "minc17",
                            "m.n2" = "m hatch",
                            "control" = "baseline")) %>%
                dplyr::filter(treatment %in% alllevels2) %>%   
                dplyr::mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
                       study = fct_collapse(treatment,
                                characterization = charlevels,
                                manipulation = maniplevels1)) %>% 
              dplyr::mutate(bird_id = gsub("[[:punct:]]", "." , id)) %>% 
              dplyr::select(bird_id, treatment, sex, study, 
                            corticosterone, estradiol, testosterone, progesterone) %>% 
              pivot_longer(cols = corticosterone:progesterone,
                           names_to = "hormone", values_to = "plasma_conc",
                           values_drop_na = TRUE) %>% 
              drop_na() %>% droplevels()  
head(PETC)
```

```{r combine-clean-split}
hormones <- rbind(prolactin, PETC)

hormones$okay <- ifelse(hormones$hormone == "corticosterone" & hormones$plasma_conc > 30, "bad",
                    ifelse(hormones$hormone == "progesterone" & hormones$plasma_conc > 5, "bad", 
                           ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
                        ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay")))))
hormones <- hormones %>% filter(okay == "okay") %>% droplevels() %>% select(-okay)

write.csv(hormones, "../results/hormones.csv", row.names = F)

# make a winder one for correlations
hormoneswide <- hormones %>% pivot_wider(names_from = "hormone", values_from = "plasma_conc", 
                                         values_fn = list(plasma_conc = mean))
hormoneswide$treatment <- factor(hormoneswide$treatment, levels = alllevels)
head(hormoneswide)

# for facet wrapping
hormones <- left_join(hormones, birds)

hormones$treatment <- factor(hormones$treatment, levels = alllevels)

hormones$hormone <- factor(hormones$hormone, 
                           levels = c("prolactin", "corticosterone", "progesterone", 
                                      "estradiol", "testosterone"))
hormones$logconc <- log10(hormones$plasma_conc)
head(hormones)
```

# do control bird with high prolactin hormone have high PRL expression in the pituitary? yes.

```{r PRLvPRL}
PRLpit <- read_csv("../results/10_PRLpit.csv") %>% 
  filter(treatment == "control") %>% 
  arrange(desc(PRL))
head(PRLpit,2)
```

## summary for owen

```{r}
meanTE <- hormoneswide %>% 
  dplyr::group_by(treatment) %>%
  dplyr::summarise(meanTestosterone = (mean(testosterone, na.rm = TRUE)), 
                   meanEstradiol = mean(estradiol, na.rm = TRUE)) %>%
  dplyr::mutate(meanTestosterone = round(meanTestosterone,2)) %>%
  dplyr::mutate(meanEstradiol = round(meanEstradiol,2)) %>%
  dplyr::filter(treatment %in% charlevels) %>%
  droplevels() %>%
  dplyr::mutate(timepoint = as.numeric(treatment)) %>%
  select(timepoint, treatment, meanTestosterone, meanEstradiol)
meanTE
```

```{r savefiles}
write.csv(meanTE, "../results/07_meanTE.csv", row.names = F)
write.csv(hormones, "../results/07_hormones.csv", row.names = F)
write.csv(hormoneswide, "../results/07_hormoneswide.csv", row.names = F)
```