### 2020-02-10 analysis/suppl-figure-1.Rmd
suppl fig

---
title: "Suppl.Fig 1"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)

library(factoextra)
source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

```{r read}
source("../R/wrangledata.R")
```

## PCA

```{r supplfig1}
subsetmakepca <- function(whichtissue, whichtreatment, whichsex){	
  colData <- colData %>%	
    dplyr::filter(tissue %in% whichtissue,	
                  treatment %in% whichtreatment,	
                  sex %in% whichsex) 	
  row.names(colData) <- colData$V1	
  # save counts that match colData	
  savecols <- as.character(colData$V1) 	
  savecols <- as.vector(savecols)

  countData <- as.data.frame(t(countData))	
  countData <- countData %>% dplyr::select(one_of(savecols)) 	
  countData <- as.data.frame(t(countData))

  mypca <- prcomp(countData)	
  mypcadf <- data.frame(PC1 = mypca$x[, 1], PC2 = mypca$x[, 2], PC3 = mypca$x[, 3], 	
                  PC4 = mypca$x[, 4],PC5 = mypca$x[, 5],PC6 = mypca$x[, 6],	
                  ID = row.names(countData))	
  mypcadf$V1 <- row.names(mypcadf)	
  mypcadf <- left_join(colData, mypcadf)	
  mypcadf <- mypcadf %>% dplyr::select(bird,sex,tissue,treatment,PC1:PC6)	
  return(mypcadf)	
}

plotcolorfulpcs <- function(mypcadf,  whichfactor, whichcolors){	
  p <- mypcadf %>%	
    ggplot(aes(x = PC1, y = PC2, shape = tissue, color = whichfactor )) +	
    geom_point(size = 1)  +	
    theme_B3() +	
    theme(legend.title = element_blank(),	
         axis.text = element_blank(),	
         legend.position = "none") +	
    labs(x = "PC1", y = "PC2")  +	
    scale_color_manual(values = whichcolors) +	
    scale_shape_manual(values = myshapes)	+
    stat_ellipse( )

  return(p)	
}

makefvizdf <-  function(whichtissue, whichtreatment, whichsex){	
  colData <- colData %>%	
      dplyr::filter(tissue %in% whichtissue,	
                  treatment %in% whichtreatment,	
                  sex %in% whichsex) 	
  row.names(colData) <- colData$V1	
  # save counts that match colData	
  savecols <- as.character(colData$V1) 	
  savecols <- as.vector(savecols)

  countData <- as.data.frame(t(countData))	
  countData <- countData %>% dplyr::select(one_of(savecols)) 	
  countData <- as.data.frame(t(countData))	
  mypca <- prcomp(countData)	
  return(mypca)	
}

plotfriz <- function(frizdf){

  p <- fviz_pca_var(frizdf,  labelsize = 3.5 , axes.linetype = "blank", 
                    repel = T , select.var= list(contrib = 3), col.var = "black")  + 
    labs(title = NULL) + 
    theme_B3() + 
    theme(axis.text = element_blank())
  return(p)
}

fpca <- subsetmakepca(tissuelevels, charlevels, "female")	
mpca <- subsetmakepca(tissuelevels, charlevels, "male")	
ffviz <- makefvizdf(tissuelevels, charlevels, "female")	
mfviz<- makefvizdf(tissuelevels, charlevels, "male")

f1 <- plotcolorfulpcs(fpca,fpca$treatment, allcolors) + labs(subtitle = "females")
f2 <- plotfriz(ffviz) + labs(subtitle = " ")

m1 <- plotcolorfulpcs(mpca,mpca$treatment, allcolors) + labs(subtitle = "males")
m2 <- plotfriz(mfviz) + labs(subtitle = " ")

allpcaplots <- plot_grid(f1,m1,f2,m2, nrow = 2, rel_heights = c(0.6,0.4), 
                labels= c("a"," ", "b"), label_size = 12)

forlegend <- plotcolorfulpcs(mpca,mpca$treatment, allcolors) + 
  theme(legend.position = "left",
        legend.direction = "vertical",
        legend.key.size = unit(0.5, 'lines')) +
  guides(color = guide_legend(ncol = 1),
         shape = guide_legend(ncol = 1))
mylegend <- get_legend(forlegend)

supplefig1 <- plot_grid(allpcaplots, mylegend, nrow = 1, rel_widths = c(1,0.2))
supplefig1
```