### 2020-02-11 analysis/figure-5.Rmd
adding manipulation deseq data

---
title: "Fig5"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

## DEGs

```{r fig2, fig.width=7}

DEG_path <- "../results/DEseq2/manip/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG)

comparisonlevels <- c( "inc.d3_m.inc.d3" ,  "inc.d9_m.inc.d9", 
                        "inc.d17_m.inc.d17", "m.n2_hatch" ,
                        "inc.d9_m.inc.d8", "inc.d17_prolong", "hatch_extend"  )

allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
allDEG$direction <- factor(allDEG$direction, levels = manipulationsandcontrols)

allDEG %>%
  group_by(sex, tissue, comparison) %>%
  summarize(n = n())

allDEG

allDEG %>%
ggplot(aes(x = comparison,  fill = direction)) +
  geom_bar(position = "dodge") +
  facet_grid(tissue~sex) +
  theme_B3() +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text.x = element_blank(),
        legend.position = "none",
        strip.placement = "right")  +
  guides(fill = guide_legend(nrow = 1)) +
  labs(x = "Parental stage-specific manipulations", y = NULL) +
  scale_fill_manual(values = allcolors,
                       name = " ", drop = FALSE) +
  scale_color_manual(values = sexcolors) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
            position = position_dodge(width = 1),
            size = 2.5, color = "black")

```