### 2020-02-17 analysis/figure-7.Rmd

---
title: "Fig7"
output: md_document
---

# Figure 7

```{r setup, include=TRUE}
library(tidyverse)
library(cowplot)

source("../R/themes.R")

knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
```

```{r fig6, fig.height=7, fig.width=7}
geneids <- read_csv("../metadata/00_geneinfo.csv")

DEG_manip_path <- "../results/DEseq2/manip/"   # path to the data
DEG_manip_files <- dir(DEG_manip_path, pattern = "*_DEGs.csv") 
DEG_manip_pathfiles <- paste0(DEG_manip_path, DEG_manip_files)
DEG_manip_files

manipDEGs <- DEG_manip_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::mutate(comparison = sapply(strsplit(file_name,'./results/DEseq2/manip/'), "[", 2),
                comparison = sapply(strsplit(comparison,'_DEGs.csv'), "[", 1)) %>%
  dplyr::select(comparison, gene, padj,logpadj, lfc, sextissue, direction)
head(manipDEGs)

DEG_char_path <- "../results/DEseq2/"   # path to the data
DEG_char_files <- dir(DEG_char_path, pattern = "*_DEGs.csv") 
DEG_char_pathfiles <- paste0(DEG_char_path, DEG_char_files)
DEG_char_files

charDEGs <- DEG_char_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::mutate(comparison = sapply(strsplit(file_name,'./results/DEseq2/'), "[", 2),
                comparison = sapply(strsplit(comparison,'_DEGs.csv'), "[", 1)) %>%
  dplyr::select(comparison, gene, padj,logpadj, lfc, sextissue, direction)
head(charDEGs)

allDEGs <- rbind(charDEGs, manipDEGs)
head(allDEGs)

nearhatch <- allDEGs %>% 
  dplyr::mutate(comparison = str_c(str_split_fixed(comparison,'\\_', 4)[,c(1,3)],collapse="_")) %>% 
  dplyr::filter(grepl("inc.d17|hatch|prolong|extend|m.inc.d8", comparison)) %>%
  dplyr::filter(!grepl("inc.d9|m.inc.d3", comparison))
head(nearhatch)

unique(nearhatch$comparison)
```