### 2020-02-18 analysis/figure-6.Rmd
save files

---
title: "fig6"
output: md_document
---

# Figure 6

```{r setup, include=TRUE}
library(tidyverse)
library(cowplot)

source("../R/themes.R")

knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
```

```{r fig6, fig.height=7, fig.width=7}
geneids <- read_csv("../metadata/00_geneinfo.csv")

DEGchar <- read_csv("../results/DEGs-char.csv")
DEGmanip <- read_csv("../results/DEGs-manip.csv")
DEGall <- rbind(DEGchar, DEGmanip)
tail(DEGall)

hatchnocontrol <- DEGall %>% filter(grepl("inc.d17", comparison)) %>% filter(!grepl("control", comparison))
hatchnocontrol %>% group_by(comparison) %>% summarize(n = n())

incd17hatch <- DEGall %>% filter(comparison == "inc.d17_hatch") %>%
  arrange(gene) %>% distinct(comparison, gene)
incd17hatch

incd17prolong <- DEGall %>% filter(comparison == "inc.d17_prolong") %>%
  arrange(gene) %>% distinct(comparison, gene)
incd17prolong

bothincs <- full_join(incd17hatch,incd17prolong, by = "gene" )
bothincs

onlyhatch <- subset(bothincs, is.na(bothincs$comparison.y))
onlyhatch
```