### 2020-03-06 analysis/figure5-maniptsne.Rmd
now with tsne fig 5

---
title: "Fig5"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(Rtsne)

source("../R/themes.R")
source("../R/functions.R")
source("../R/wrangledata.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin

```{r tSNE-df}

# prep for tsne

hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)

addgroupings <- function(df){

  df <- df %>% mutate(hiloPRL = fct_collapse(treatment, 
                                  lo = c("m.inc.d8", "m.inc.d3", "m.inc.d9", "inc.d3", "inc.d9", "lay", "bldg"),
                                  hi = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" ,  "m.n2", "extend", "n5")),
                extint = fct_collapse(treatment, 
                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
                                  chicks = c( "m.inc.d8", "hatch", "extend", "n5"),
                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
  df$extint <- factor(df$extint, levels = levelsextint)
  return(df)
}

hyptsne <- addgroupings(hyptsne)
pittsne <- addgroupings(pittsne)
gontsne <- addgroupings(gontsne)

```

## make figure

```{r fig5, fig.width=7, fig.height=6}

expdesign2 <- png::readPNG("../figures/images/DoveParentsRNAseq_maniptsne.png")
expdesign2 <- ggdraw() +  draw_image(expdesign2, scale = 1)

b <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) + labs(subtitle = "hypothalamus ~ parental stage", x = NULL)  + facet_wrap(~sex, scales = "free")
c <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors ) + labs(subtitle = "pituitary ~ parental stage", x = NULL) + facet_wrap(~sex, scales = "free") 
d <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors ) + labs(subtitle = "gonads ~ parental stage")  + facet_wrap(~sex, scales = "free")

bcd <- plot_grid(b,c,d, nrow = 3, labels = c("b", "c", "d"), label_size = 8, rel_heights = c(1,1,1))

e <- plottsneelipsev2(hyptsne, hyptsne$hiloPRL, allcolors ) + labs(subtitle = "hypothalamus ~ estimated prolactin", x = NULL) + facet_wrap(~sex, scales = "free")  
f <- plottsneelipsev2(pittsne, pittsne$hiloPRL, allcolors ) + labs(subtitle = "pituitary ~ estimated prolactin", x = NULL) + facet_wrap(~sex, scales = "free")   
g <- plottsneelipsev2(gontsne, gontsne$hiloPRL, allcolors ) + labs(subtitle = "gonads ~ estimated prolactin" ) + facet_wrap(~sex, scales = "free")

efg <- plot_grid(e,f,g, nrow = 3, labels = c("h", "i", "j"), label_size = 8, rel_heights = c(1,1,1))

h <- plottsneelipsev2(hyptsne, hyptsne$extint, allcolors ) + labs(subtitle = "hypothalamus ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")  
i <- plottsneelipsev2(pittsne, pittsne$extint, allcolors ) + labs(subtitle = "pituitary ~ external enviornment", x = NULL) + facet_wrap(~sex, scales = "free")   
j <- plottsneelipsev2(gontsne, gontsne$extint, allcolors ) + labs(subtitle = "gonads ~ external enviornment" ) + facet_wrap(~sex, scales = "free")

hij <- plot_grid(h,i,j, nrow = 3, labels = c("e", "f", "g"), label_size = 8, rel_heights = c(1,1,1))

bcdhijefg <- plot_grid(bcd, hij,  efg, nrow = 1)

fig6 <- plot_grid(expdesign2, bcdhijefg, nrow = 2, labels = c("e"), label_size = 8, rel_heights = c(0.5,1))
fig6

```