### 2020-03-28 analysis/figure-6.Rmd
first attempt at candidate gene figure

---
title: "Fig6"
author: "Rayna M Harris"
date: "3/26/2020"
output: md_document
---

## manipulation box plots for candidate genes

```{r}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)
library(kableExtra)

source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")

candidategenesslim <- c("OXT", "AVP", "GNRH1", "GNRHR", 
                    "AR",  "CYP19A1", 
                     "AVPR1A", "AVPR1B", "AVPR2","VIP",
                  "DRD1", "DRD2", 
                  "PRL", "PRLR",  
                    "ESR1","ESR2", "LBH",

                    "FOS", "JUN", "EGR1", "BDNF"
                      )

candidategenes <- candidategenesslim

```

## variance stabilized gene expression  (vsd)

```{r }
geneids <- read_csv("../metadata/00_geneinfo.csv")

vsd_path <- "../results/DEseq2/manip/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = blk.s031.pu.d_female_gonad_prolong:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") 
allvsd <- as.tibble(allvsd)

getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/manip/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    #dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1))# %>%
   # dplyr::recode(treatment, "m.hatch" = "m.n2", .default = levels(treatment))

     dplyr::mutate(treatment = recode(treatment, "m.hatch" = "m.n2"))%>%

    dplyr::select(sex, tissue, treatment, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)  %>%
    drop_na()
  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
  return(candidates)
}

hypvsd <- getcandidatevsd(candidategenesslim, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd(candidategenesslim, "pituitary", sexlevels)
gonvsd <- getcandidatevsd(candidategenesslim, "gonad", sexlevels)
head(hypvsd)
head(pitvsd)
head(gonvsd)

candidatevsd <- rbind(hypvsd, pitvsd)
candidatevsd <- rbind(candidatevsd, gonvsd)

head(candidatevsd)
levels(candidatevsd$treatment)
```

## DEGs

```{r allDEGs}
DEG_path <- "../results/DEseq2/manip/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG)

allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
allDEG$direction <- factor(allDEG$direction, levels = alllevels)

head(allDEG)

table1 <- allDEG %>%
  mutate(updown = ifelse(lfc > 0, "+", "-"))  %>%
  mutate(geneupdown = paste(gene, updown, sep = "")) %>%
  filter(gene %in% candidategenes) %>%
  arrange(geneupdown) %>%
  group_by(sex, tissue, comparison) %>%
  summarize(genes = str_c(geneupdown, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = genes ) %>%
  select(sex, tissue, contains("m."))  %>%
  arrange( tissue, sex)

(table1$inc.d3_m.inc.d3)
(table1$inc.d9_m.inc.d9)
(table1$inc.d17_m.inc.d17)
(table1$hatch_m.n2)
```

## Figs

```{r fig6, fig.height=7, fig.width=7}
makeboxplotsmanip <- function(df, whichgene, mysubtitle, whichsex, whichtimepoint){
  p <- df %>%
    filter(treatment %in% alllevels,
           gene %in% whichgene,
           sex %in% whichsex) %>%
    filter(treatment %in% whichtimepoint) %>%
    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
    geom_boxplot(outlier.shape = NA) + 
    geom_jitter(size = 0.5, width = 0.1) +
    facet_wrap(~gene, scales = "free_y", nrow = 1) +
    theme_B3() +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          plot.caption = element_text(face = "italic")) +
    labs(y = "gene expression" , x = "Parental Stages and Manipulations", subtitle = mysubtitle) +
    theme(strip.text = element_text(face = "italic"))
  return(p)
}

## hyp 
a2 <- makeboxplotsmanip(hypvsd, c("DRD1","EGR1","PRLR"), "Female hypothalamus", "female",
                  c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17",  "hatch", "m.n2"))  +
  theme(axis.title.x = element_blank())
b2 <- makeboxplotsmanip(hypvsd, c("BDNF","ESR2", "FOS"), " ", "female",
                  c("inc.d3", "m.inc.d3", "inc.d17", "m.inc.d17"))  +
  theme(axis.title.x = element_blank())

c <- makeboxplotsmanip(hypvsd, c("AVP","OXT"), " ", "female", 
                       c("inc.d17", "m.inc.d17"))  +
  theme(axis.title = element_blank())
d <- makeboxplotsmanip(hypvsd, c("AVPR1A","CYP19A1","GNRH1"), " ", "female",
                   c("hatch", "m.n2"))  +
  theme(axis.title = element_blank())
e <- makeboxplotsmanip(hypvsd, c("JUN"), " ", "female", 
                 c("hatch", "m.n2"))  +
  theme(axis.title = element_blank()) + labs(subtitle = " ")

f1 <- makeboxplotsmanip(hypvsd, c("DRD1", "EGR1", "PRLR"), "Male hypothalamus", "male",
                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
  theme(axis.title.x = element_blank())

f2 <- makeboxplotsmanip(hypvsd, c( "PRL"), " ", "male",
                  c( "inc.d17", "m.inc.d17", "hatch", "m.n2"))  +
  theme(axis.title = element_blank())

h <- makeboxplotsmanip(hypvsd, c("AR", "AVP", "BDNF", "ESR1", "LBH"), " ", "male",
                  c( "inc.d17", "m.inc.d17"))  + labs(caption = " ")

i <- makeboxplotsmanip(hypvsd, c("GNRH1"), " ", "male",
                  c("hatch", "m.n2"))  +
  theme(axis.title = element_blank()) +
  labs(x = " ",
       caption = "Candidate genes that were not differentially expressed following removal: AVPR1B, AVPR2, ESR1, GNRHR, VIP")

ab2 <- plot_grid(a2,c, rel_widths = c(18,5), labels = "auto", label_size = 8)
cde <- plot_grid(b2,d,e, rel_widths = c(12,6,2), nrow =  1, align = "h",  
                 labels = c("c", "d", " "), label_size = 8)
f12 <- plot_grid(f1,f2, rel_widths = c(3,1), labels = c("e",  " "), label_size = 8)
hi <- plot_grid(h,i, ncol = 2, rel_widths = c(10,2),labels = c("f",  "g"), label_size = 8, align = "h")

plot_grid(ab2, cde, f12,hi, nrow = 4, rel_heights = c(1,1,1,1.25))

```

```{r writefiles}
write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
```