### 2020-04-21 analysis/figure-6.Rmd
working on manip figure

---
title: "Fig6"
author: "Rayna M Harris"
date: "3/26/2020"
output: md_document
---

## manipulation box plots for candidate genes

```{r setup}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)
library(kableExtra)

source("../R/themes.R")

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

## Candidate Genes (from literature and relevant GO terms)

```{r GO}
source("../R/wrangledata.R")
```

## variance stabilized gene expression  (vsd)

```{r vsd}
vsd_path <- "../results/DEseq2/manip/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") %>% 
  pivot_longer(cols = blk.s031.pu.d_female_gonad_prolong:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") 
allvsd <- as.tibble(allvsd)

head(allvsd)

getcandidatevsd <- function(whichgenes, whichtissue, whichsex){
  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1)) %>%
    dplyr::mutate(sextissue = sapply(strsplit(sextissue, '../results/DEseq2/manip/'), "[", 2)) %>%
    dplyr::mutate(sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4)) %>%
    dplyr::mutate(treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1)) %>%
     dplyr::mutate(treatment = recode(treatment, "m.hatch" = "m.n2",
                                                  "extend.hatch" = "m.n2",
                                      "inc.prolong" = "prolong")) %>%

    dplyr::mutate(day = fct_collapse(treatment,
                                     "day 3" = c("inc.d3", "m.inc.d3" ),
                                     "day 9" = c("inc.d9", "m.inc.d9", "m.inc.d8"  ),
                                     "day 17" = c("inc.d17", "m.inc.d17", "prolong"  ),
                                     "hatch" = c("hatch", "m.n2", "extend"  )
                                     )) %>%

    dplyr::mutate(manip = fct_collapse(treatment,
                                       "reference" = charlevels,
                                       "removal" = c("m.inc.d3", "m.inc.d9", "m.inc.d17","m.n2"),
                                       "replacement" = c("m.inc.d8", "prolong", "extend"))) %>%

    dplyr::select(sex, tissue, treatment, day,  manip, gene, samples, counts) %>%
    filter(tissue == whichtissue, sex %in% whichsex)  %>%
    drop_na()
  candidates$treatment <- factor(candidates$treatment, levels = alllevels)
  return(candidates)
}

hypvsd <- getcandidatevsd(candidategenes, "hypothalamus", sexlevels)
pitvsd <- getcandidatevsd(candidategenes, "pituitary", sexlevels)
gonvsd <- getcandidatevsd(candidategenes, "gonad", sexlevels)
candidatevsd <- rbind(hypvsd, pitvsd)
candidatevsd <- rbind(candidatevsd, gonvsd)
head(candidatevsd)
```

## DEGs

```{r allDEGs}
DEG_path <- "../results/DEseq2/manip/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/manip/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG)

allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevels)
allDEG$direction <- factor(allDEG$direction, levels = alllevels)
head(allDEG)

candidateDEGS <- allDEG %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, contains("m.")) 
candidateDEGS

table2 <- left_join(candidateDEGS, GOgenesWide) %>%
  select(gene, m.inc.d17_prolong:m.inc.d9_m.inc.d17, parentalcare, parentalbehavior, NCBI) %>%
  mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
         parentalbehavior = if_else(is.na(parentalbehavior), " ", "X")) %>%
  rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
table2$numDEGs <- rowSums(is.na(table2)) # count NAs to know how many are NS
table2 <- table2 %>% 
  mutate(sig = ifelse(numDEGs == 6, "NS", "DEG")) %>% 
  arrange(sig, gene)  %>%  select(-sig, -numDEGs)
table2[is.na(table2)] <- " " # replace NA with blank space so it's pretty
table2
```

```{r fig6}
candidatevsd %>%
  filter(gene == "PRL") %>%
  ggplot(aes(y = counts, x = day, color = manip)) +
  geom_boxplot() +
  facet_wrap(tissue~sex, scales = "free") +
  theme(legend.position = "bottom") 
```

```{r writefiles}
write.csv(table2, "../results/table2.csv")

# for victoria
prlmanip <- candidatevsd %>%
  filter(gene == "PRL", 
         treatment %in% c("extend", "m.inc.d8"))
write.csv(prlmanip, "../results/prlmanip.csv")
```