### 2020-05-03 analysis/22_symphony.Rmd
transcriptional symphony

---
title: "Transcriptoinal symphony"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)

source("../R/themes.R")

theme_music <- function () { 
  theme_classic(base_size = 12,
                base_family = 'Helvetica') +
    theme(
      panel.grid.major  = element_blank(),  # remove major gridlines
      panel.grid.minor  = element_blank()  # remove minor gridlines
    )
}

knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/favegenes/")

```

## Music example

```{r transcriptionalsymphony}
song <- c("Oh,", "sin-", "ner", "man,", "where", "_", "he", "gon'", "run", "_", "to?")
stage <- c( "bldg", "lay", "inc.d3", "m.inc.d3" ,
            "inc.d9", "m.inc.d9"  , 
            "inc.d17", "m.inc.d17" ,
            "hatch", "m.n2" ,
            "n5")
position <- c(1,2,3,4,5,6,7,8,9,10,11)
gene1 <- c(4,4,4,4,6,6,5,4,4,4,4)
gene2 <- gene1 + 2
gene3 <- gene2 + 2

sinnerman <- as.data.frame(t(rbind(position, gene1,gene2,gene3)))
sinnerman$song <- song
sinnerman$stage <- stage
sinnerman$image <- sample(c("https://cdn.pixabay.com/photo/2016/03/23/20/49/music-note-1275650_960_720.png"), 
                        size=11, replace = TRUE)
sinnerman <- sinnerman %>%
  pivot_longer(gene1:gene3, names_to = "music", values_to = "genes")

sinnermansong <- ggplot(sinnerman, aes(x = position, y = genes)) +
  #geom_line(aes(color = music)) +
  geom_image(aes(image=image), size = 0.15) +
  scale_x_continuous(breaks = position,
                    labels = song,
                     name = element_blank()) +
  scale_y_continuous(breaks = c(3, 5, 7, 9),
                     labels = c( "F",   "A",   "C",  "E"),
                     limits = c(-1,12)) +
  geom_hline(yintercept=c(2.15,4.15,6.15,8.15,10.15))   +
  theme_music() +
  theme(axis.ticks = element_blank()) +
  labs(subtitle = "Sinnerman by Nina Simone", y = NULL)
sinnermansong

```