### 2020-05-12 analysis/00_datawrangling.Rmd
slimming data prep

---
title: "00_datawrangling.Rmd"
output: md_document
---

```{r setup}
# https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html

library(tidyverse)

source("../R/themes.R")
```

# Data Wrangling

## import Kallisto transcript data, make gene info file

```{r countData}
# import count data, set rows at entreziz
kallistodata <- read.table("../results/kallistocounts.txt", 
                        sep = ",", row.names = NULL) %>%
  rename("NCBI" = "entrezid",
         "gene" = "Name")
head(kallistodata)

## save gene information
geneinfo <- kallistodata %>%
  select(gene, geneid, NCBI) 
head(geneinfo)
```

## # check for genes that have multple transcripts expressed

```{r}
isoforms <- kallistodata %>%
  group_by(gene) %>%
  summarize(n = n()) %>%
  filter(n > 1) %>%
  group_by(n) %>% 
  summarize(genes = str_c(gene, collapse = ", ")) %>%
  arrange(desc(n)) %>%
  rename("n(isoforms)" = "n") %>% 
  mutate("n(counts)" = str_count(genes, ",") + 1)
head(isoforms)

# for example, these gene have 2 and 3 isoforms
geneinfo %>% filter(gene == "GRIN1")
geneinfo %>% filter(gene == "CACNA1C")
```

## group data by gene

```{r}
head(kallistodata)
# aggregate transcript counts to gene counts
countData <- kallistodata %>% 
  select(-row.names, -geneid, -NCBI) %>% 
  pivot_longer(-gene, names_to = "samples", values_to = "counts") %>%  
  pivot_wider(
    names_from = samples, 
    values_from = counts,
    values_fn = list(counts = sum))  %>% 
  arrange(gene) %>% 
  filter(gene != "")
countData <- as.data.frame(countData)
row.names(countData) <- countData$gene ## make gene the row name
countData[1] <- NULL ## make gene the row name
countData <- round(countData) #round all value to nearest 1s place
head(countData[13:15])

# print tolal num of genes and samples
dim(countData)
```

## wrangle colData

```{r colData}
colData <- read.table(file.path( "../metadata/kallistosamples.txt"),
                      header = F, stringsAsFactors = F) %>%
  # use strsplit to cut the filename into meaningful columns
  mutate(bird = sapply(strsplit(V1,'\\_'), "[", 1),
         sex = sapply(strsplit(V1,'\\_'), "[", 2),
         tissue = sapply(strsplit(V1,'\\_'), "[", 3),
         temp = sapply(strsplit(V1,'\\_'), "[", 4)) %>%
  mutate(treatmenttemp = sapply(strsplit(temp,'\\.'), "[", 1),
         NYNO = sapply(strsplit(temp,'\\.'), "[", 2)) %>%

  # rename variables
  mutate(treatment = ifelse(grepl("extend-hatch", treatmenttemp), "extend",
                            ifelse(grepl("inc-prolong", treatmenttemp), "prolong",
                                   ifelse(grepl("m.hatch", treatmenttemp), "m.n2",
                                          ifelse(grepl("m.inc.d8", treatmenttemp), "early",
                                   treatmenttemp))))) %>%
   select(-temp, -NYNO, -treatmenttemp ) %>%
  # replace dashes with periods (to make deseq happy)
  mutate(bird = gsub("-", ".", bird),
         treatment = gsub("-", ".", treatment),
         V1 = gsub("-", ".", V1))
# specify levels that will be factors
cols <- c("sex", "treatment", "tissue")
colData[cols] <- lapply(colData[cols], factor) 
colData <- colData %>%
  mutate(tissue = factor(tissue, levels = c("hypothalamus", "pituitary", "gonad"))) %>% 
  mutate(tissue = fct_recode(tissue, "gonads" = "gonad")) %>%
  mutate(group = paste(sex, tissue, treatment, sep = "."),
         study = ifelse(grepl("m.|extend|prolong", treatment), 
                        "manipulation", "charcterization")) %>%
  mutate(study = factor(study, levels = c("charcterization", "manipulation")))
colData <- as.data.frame(colData)
row.names(colData) <- colData$V1
head(colData)
```

## check that rownames and colnames match for DESeq

```{r}
str(colData)
ncol(countData) == nrow(colData)

colData %>% select(sex, tissue, treatment, study)  %>%  summary()
table(colData$sex, colData$treatment, colData$tissue)
```

## save bird data

```{r}
birds <- colData %>%
  select(bird, sex, treatment) %>%
  distinct()
birds
```

## save files for downstream use

```{r write}
write.csv(countData, "../results/00_counts.csv")
write.csv(isoforms, "../results/00_geneswithisoforms.csv", row.names = T)

write.csv(colData, "../metadata/00_colData.csv")
write.csv(birds, "../metadata/00_birds.csv")
write.csv(geneinfo, "../metadata/00_geneinfo.csv", row.names = TRUE)

```