### 2020-05-13 analysis/06_datawrangling.Rmd
now with WGCNA as step 5, wrangle as 6

---
title: "05_wrangle"
output: md_document
---

```{r setup}
library(tidyverse)
library(readr)

source("../R/themes.R")
```

## Wrangle data for hyotheses testing

```{r }
# col data with internal (hiloPRL) and external  hypothesese
colData <- read_csv("../metadata/04_colData.csv")
tail(colData)

# candidate genes from GO and literature
candidategenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% pull(gene)
candidategenes

## genes WGCNA prl module
WGCNAgenes <- read_csv("../results/05_PRLmodule.csv") %>% pull(x)
WGCNAgenes
```

## candidate genes

```{r}

```

```{R}
pithilo <- makecorrdf("female", "pituitary", hilogenes)
piteggchick <- makecorrdf("female", "pituitary", eggchickgenes)
pitWGCNA <- makecorrdf("female", "pituitary", WGCNAgenes[1:20])

write.csv(WGCNAvsd, "../results/03_WGCNAvsd.csv")
write.csv(hypothesisDEGs, "../results/03_hypothesisDEGs.csv")

```