### 2020-05-26 analysis/02_DESeq2.Rmd
more manip comparisons

---
title: "02_DESeq2"
output: md_document
---

```{r setup}
library(tidyverse)
library(DESeq2)
library(BiocParallel)
register(MulticoreParam(6))
library(cowplot)

source("../R/functions.R")  # load custom functions 
source("../R/themes.R")  # load custom themes and color palletes

knitr::opts_chunk$set(fig.path = '../figures/DESeq2/',  message=F, comment=F, warning=F)
```

# DESeq2

- DESeq2 was not designed to run on 100+ samples. But, I really like it, so I do it anyways. Some of these commands take like 15 min to run using 6 cores. 
_ Also, I don't run every chuck every time. When I want new analysese, I add them to the bottom and set `eval = F` for chunks I don't want to rerurn

```{r coldata}
countData <- read.csv("../results/00_counts.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

colData <- read.csv("../metadata/00_colData.csv", header = T, row.names = 1) %>%
  mutate(tissue = factor(tissue, levels = tissuelevels),
         treatment <- factor(treatment, levels = alllevels)) %>%
  mutate(sextissue = as.factor(paste(sex, tissue, sep = "_"))) 
colData <- colData %>% drop_na()
# 987 samples, 8 variables

ncol(countData) == nrow(colData)

```

## tissue and sex differences

```{r sextissue, eval = F}

## first or sex

for(i in levels(colData$tissue)){

  newcolData <- subsetcolData4(colData, i)
  print(head(newcolData))
  # save counts that match colData
  savecols <- as.character(newcolData$V1) 
  savecols <- as.vector(savecols)

  newcountData <- countData %>% dplyr::select(one_of(savecols))

  dds <- DESeqDataSetFromMatrix(countData = newcountData,
                                colData = newcolData,
                                design = ~ sex )

  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
  print(dds)
  print(dim(dds))
  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis

  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))

  myfilename = paste0("../results/DEseq2/sex/", i, "_vsd.csv")
  write.csv(vsd, myfilename)

  print(head(vsd))

  # save differential gene expression results
  sex <- createDEGdfsavesex("male", "female", i)
}

## then for tissue

dds <- DESeqDataSetFromMatrix(countData = countData,
                                colData = colData,
                                design = ~ tissue )
dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
print(dds)
print(dim(dds))
dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis

vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))
write.csv(vsd, "../results/DEseq2/tissue/tissuevsds.csv")
print(head(vsd))

# save differential gene expression results
HP <- createDEGdfsavestissue("pituitary", "hypothalamus")
HG <- createDEGdfsavestissue("gonad", "hypothalamus")
PG <- createDEGdfsavestissue("gonad", "pituitary")

str(HP)
```

## FOR DEGs of all TREATMENT comparisons, char y manip

```{r treatment, eval = T}

createDEGdftreatment <- function(up, down, mytissue){

  res <- results(dds, contrast = c("treatment", up, down), 
                 independentFiltering = T, alpha = 0.1)

  DEGs <- data.frame(gene = row.names(res),
                        padj = res$padj, 
                        logpadj = -log10(res$padj),
                        lfc = res$log2FoldChange,
                        sextissue = mytissue)
  DEGs <- na.omit(DEGs)
  DEGs <- DEGs %>%
    dplyr::mutate(direction = ifelse(DEGs$lfc > 0 & DEGs$padj < 0.1, 
                                     yes = up, no = ifelse(DEGs$lfc < 0 & DEGs$padj < 0.1, 
                                                           yes = down, no = "NS"))) %>% 
    dplyr::arrange(desc(lfc))

  DEGs$direction <- factor(DEGs$direction, levels = c(down, "NS", up))

  # write DEGsframe of only significant genes
  DEGs <- DEGs %>% dplyr::filter(direction != "NS")
  print(str(DEGs))

  partialfilename = paste("_", down, "_", up, sep = "")
  myfilename = paste0("../results/DESeq2/treatment/", mytissue, partialfilename, "_DEGs.csv")

  write.csv(DEGs, myfilename, row.names = F)
  # return DEGs frome with all data, included NS genes
  #print(head(DEGs))
}

for(i in levels(colData$sextissue)){

  newcolData <- subsetcolData2(colData, i)

  # save counts that match colData
  savecols <- as.character(newcolData$V1) 
  savecols <- as.vector(savecols)

  newcountData <- countData %>% dplyr::select(one_of(savecols))

  dds <- DESeqDataSetFromMatrix(countData = newcountData,
                                colData = newcolData,
                                design = ~ treatment )
  dds <- dds[rowSums(counts(dds) > 1) >= 10]  # filter more than sample with less 0 counts
  print(dds)
  print(dim(dds))
  dds <- DESeq(dds, parallel = TRUE) # Differential expression analysis

  vsd <- as.data.frame(assay(vst(dds, blind=FALSE)))

  myfilename = paste0("../results/DEseq2/treatment/", i, "_vsd.csv")
  write.csv(vsd, myfilename)

  #return(dds)
  #return(vsd)
  #print(head(vsd))

  # save differential gene expression results

  inc.d9.early <- createDEGdftreatment("inc.d9", "early",  i) 
  earyl.hatch <- createDEGdftreatment("early", "hatch",  i)

  inc.d17.prolong <- createDEGdftreatment("inc.d17", "prolong",  i) 
  prolong.hatch <- createDEGdftreatment("prolong", "hatch",  i)

  hatch.extend <- createDEGdftreatment("hatch", "extend",  i) 
  extend.n5 <- createDEGdftreatment("extend", "n5",  i)

  ## previously run samples in comparisons.R

}

```