### 2020-08-06 analysis/figure-1.Rmd
better fig 1

---
title: "Fig1"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(Rtsne)
library(kableExtra) 
library(ggtext) # for italic fonts

source("../R/themes.R")
source("../R/functions.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

#  Experimental design, tSNE analysis, and bar chars

## import limma counts and sample info

```{r data}
countData <- read_csv("../results/01_limma.csv") %>%
  column_to_rownames(var = "X1")
countData <- as.data.frame(t(countData))
head(countData[1:3])

colData <- read_csv("../metadata/00_colData.csv") %>%
  mutate(treatment = factor(treatment, levels = alllevels),
         tissue = factor(tissue, levels = tissuelevels)) %>% 
  column_to_rownames(var = "X1") 
head(colData)

# check ready for analysis
# row.names(countData) == row.names(colData)
head(row.names(countData) == row.names(colData))
```

## tsne

```{r tSNE-df}
# prep for tsne useing count data from limma and the custom `subsetmaketsne` function
hyptsne <- subsetmaketsne("hypothalamus", alllevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", alllevels, sexlevels)
gontsne <- subsetmaketsne("gonads", alllevels, sexlevels)
```

## Figure 1

```{r fig1, fig.width=7, fig.height=4.5}

a <- png::readPNG("../figures/images/fig_fig1a.png")
a <- ggdraw() +  draw_image(a, scale = 1)

e1 <- plottsneelipsev3(hyptsne, hyptsne$treatment, allcolors) + 
  labs(subtitle = "hypothalamus") 
e2 <- plottsneelipsev3(pittsne, pittsne$treatment, allcolors ) + 
  labs(subtitle = "pituitary") 
  theme(strip.text = element_blank())
e3 <- plottsneelipsev3(gontsne, gontsne$treatment, allcolors ) + 
  labs(subtitle = "gonads")

e <- plot_grid(e1,e2,e3, ncol = 3, labels = c("D"), label_size = 8)

fig1 <- plot_grid(a, e, nrow = 2, rel_heights = c(2.4,1.6))
fig1
```

## Save files

```{r write}
pdf(file="../figures/fig1-1.pdf", width=7, height=4.5)
plot(fig1)
dev.off()
```