### 2020-08-06 analysis/boneyard.Rmd
not sure i like it but same 3 figs for hyp pit gon

## Treatment specific DEGs

```{r suppltables}
allDEG <- read_csv("../results/03_allDEG.csv") %>%
  mutate(tissue = factor(tissue, levels = tissuelevel),
         direction = factor(direction, levels = alllevels))

# for figure
DEGchar <- allDEG %>% 
  filter(comparison %in% comparisonlevelschar) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelschar))

# for suppl figures
DEGcontrol <- allDEG %>% 
  filter(grepl("control", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol))

DEGbldg <- allDEG %>% 
  filter(grepl("bldg", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
  drop_na()
```

## Sex and Tissue -related DEGs

```{r DEGs, message=F}

## other DEG stuff

pitsex <- read_csv("../results/DEseq2/sex/pituitary_female_male_DEGs.csv")
gonsex <- read_csv("../results/DEseq2/sex/gonad_female_male_DEGs.csv")

hyppit <- read_csv("../results/DEseq2/tissue/hypothalamus_pituitary_DEGs.csv") 
hypgon <- read_csv("../results/DEseq2/tissue/hypothalamus_gonad_DEGs.csv")
pitgon <- read_csv("../results/DEseq2/tissue/pituitary_gonad_DEGs.csv")

hyppit$direction <- factor(hyppit$direction, levels = c("hypothalamus", "pituitary"))
hypgon$direction <- factor(hypgon$direction, levels = c("hypothalamus",  "gonad"))
pitgon$direction <- factor(pitgon$direction, levels = c( "pituitary", "gonad"))

```

```{r barplots}

b2 <- sexbarplots(hyppit, 0, 8100) + labs(subtitle = " ", y = "DEGs w/ + LFC") +
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  scale_x_discrete(labels=c("hypothalamus" = "hyp", "pituitary" = "pit" ))
b3 <- sexbarplots(hypgon, 0, 8100) + labs(subtitle = " " ) +
  theme(axis.text.y = element_blank(), axis.line.y = element_blank(), 
        axis.ticks = element_blank(), axis.title.y = element_blank()) +
    scale_x_discrete(labels=c("hypothalamus" = "hyp",  "gonad" = "gon"))
b4 <- sexbarplots(pitgon, 0, 8100) + labs(subtitle = " " ) +
  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
        axis.ticks = element_blank(), axis.title.y = element_blank()) +
    scale_x_discrete(labels=c(  "pituitary" = "pit", "gonad" = "gon"))

c3 <- sexbarplots(pitsex, 0, 8100) + labs(subtitle = "pituitary" ) + 
  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
        axis.ticks = element_blank(), axis.title.y = element_blank())
c4 <- sexbarplots(gonsex, 0, 8100) + labs(subtitle = "gonads" ) +
  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),  
        axis.ticks = element_blank(), axis.title.y = element_blank())
bc <- plot_grid(b1,b2,b3,b4, c1,c2,c3,c4, nrow = 1, rel_widths = c(1.5,1.1,0.9,0.9,1.5,1.1,0.9,0.9),
                labels = c("A", "", "", "", "B"), label_size = 8)
```

## suppl fig 1 controls versus bldg

```{r supplfig-1, fig.width=5, fig.height=5}

s1c <- makebargraphsuppl(DEGcontrol, "pituitary","DEGs w/ + LFC", 0, 5000)  +  
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        strip.text.x = element_blank())    + 
  labs(subtitle = "pituitary") 
s1e <- makebargraphsuppl(DEGcontrol, "gonad","DEGs w/ + LFC", 0, 5000) +  
  theme(strip.text.x = element_blank()) +
  scale_x_discrete(labels = comparisonlabelscontrol) +
  labs(subtitle = "gonads")

s1d <- makebargraphsuppl(DEGbldg, "pituitary",NULL, 0, 1000)  +  
  theme(axis.text.x = element_blank(), 
        axis.title.x = element_blank(), 
        strip.text.x = element_blank())   
s1f <- makebargraphsuppl(DEGbldg, "gonad",NULL, 0, 1000) +  
  theme(strip.text.x = element_blank()) +
  scale_x_discrete(labels = comparisonlabelsbldg)

s1 <- plot_grid(s1a, s1b, 
          s1c, s1d, 
          s1e, s1f, nrow = 3,
          rel_heights = c(1,0.9,1.3),
          labels = c("A", "B"), label_size = 8, rel_widths = c(1.1,1))

sa <- png::readPNG("../figures/images/fig_supplfig-1.png")
sa <- ggdraw() +  draw_image(sa, scale = 1)

supplfig1 <- plot_grid(s1, sa, ncol = 1, rel_heights = c(2,0.7))
supplfig1

# for volcano plots

pitinc <- DEGchar %>% filter(tissue == "pituitary", comparison == "inc.d9_inc.d17") 
goninc <- DEGchar %>% filter(tissue == "gonad", comparison == "lay_inc.d3")

s2a <- plot.volcano("hypothalamus", sexlevels,  "hatch_n5") + facet_wrap(~sex) + 
  labs(subtitle = "hypothalamus", title = "Volcano plots") 
s2c <- plot.volcano("pituitary", sexlevels,  "inc.d9_inc.d17") + 
  facet_wrap(~sex) + 
  labs(subtitle = "pitutiary") + 
  theme(legend.title = element_blank(), strip.text = element_blank())
s2e <- plot.volcano("gonad", sexlevels,  "lay_inc.d3") + facet_wrap(~sex) + 
  labs( x = "Log-fold change (LFC)", subtitle = "gonads") + 
  theme(legend.title = element_blank(), strip.text = element_blank())

s2b <- makebargraph(hyphatch, "hypothalamus","Total DEGs with +LFC", 0, 1800)  + 
  labs(subtitle = " ", title = "Bar graphs") + scale_x_discrete(labels = "hatch\nvs n5")
s2d <- makebargraph(pitinc, "pituitary","Total DEGs with +LFC", 0, 1800) + 
  labs(subtitle = " ") + scale_x_discrete(labels = "inc.d9 vs\ninc.d17") + 
  theme(strip.text = element_blank())
s2f <- makebargraph(goninc, "gonad","Total DEGs with +LFC", 0, 1800) + 
  labs(subtitle = " ", x = "stages compared") + scale_x_discrete(labels = "lay vs\ninc.d3") + 
  theme(strip.text = element_blank())
supplfig2 <- plot_grid(s2a, s2b, s2c, s2d, s2e, s2f,
          rel_widths = c(2,1), rel_heights = c(1.1,1,1.1), ncol = 2,
          labels = c("A", "B"), label_size = 8)
supplfig2
```

```{r}
suppltable1 <- DEGchar %>%
  arrange( tissue, sex, direction, gene)
head(suppltable1)
write_csv(suppltable1, "../results/suppltable1.csv")

suppltable2 <- rbind(DEGcontrol, DEGbldg)  %>%
  arrange( tissue, sex, direction, gene)
head(suppltable2)
write_csv(suppltable2, "../results/suppltable2.csv")

```