### 2020-08-07 analysis/figure-1.Rmd

---
title: "Fig1"
output: md_document
---

```{r setup, include=T}
library(tidyverse)
library(cowplot)
library(Rtsne)

source("../R/themes.R")
source("../R/functions.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

#  Experimental design, tSNE analysis, and bar chars

## import limma counts and sample info

```{r data}
countData <- read_csv("../results/01_limma.csv") %>%
  column_to_rownames(var = "X1")
countData <- as.data.frame(t(countData))
head(countData[1:3])

colData <- read_csv("../metadata/00_colData.csv") %>%
  mutate(treatment = factor(treatment, levels = alllevels),
         tissue = factor(tissue, levels = tissuelevels)) %>% 
  column_to_rownames(var = "X1") 
head(colData)

# check ready for analysis
# row.names(countData) == row.names(colData)
head(row.names(countData) == row.names(colData))
```

## tsne

```{r tSNE-df}
# prep for tsne useing count data from limma and the custom `subsetmaketsne` function
hyptsne <- subsetmaketsne("hypothalamus", alllevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", alllevels, sexlevels)
gontsne <- subsetmaketsne("gonads", alllevels, sexlevels)
```

## Figure 1

```{r fig1, fig.width=7, fig.height=4.5}

a <- png::readPNG("../figures/images/fig_fig1a.png")
a <- ggdraw() +  draw_image(a, scale = 1)

e1 <- plottsneelipsev3(hyptsne, hyptsne$treatment, allcolors) + 
  labs(subtitle = "hypothalamus") 
e2 <- plottsneelipsev3(pittsne, pittsne$treatment, allcolors ) + 
  labs(subtitle = "pituitary") 
  theme(strip.text = element_blank())
e3 <- plottsneelipsev3(gontsne, gontsne$treatment, allcolors ) + 
  labs(subtitle = "gonads")

e <- plot_grid(e1,e2,e3, ncol = 3, labels = c("D"), label_size = 8)

fig1 <- plot_grid(a, e, nrow = 2, rel_heights = c(2.4,1.6))
fig1
```

## Supplemental PCA data

```{r}
# pca relies on count data from limma
pca_h1 <- subsetmakepca("hypothalamus", alllevels, sexlevels)	
pca_h2 <- makefvizdf("hypothalamus", alllevels, sexlevels)

pca_p1 <- subsetmakepca("pituitary", alllevels, sexlevels)	
pca_p2 <- makefvizdf("pituitary", alllevels, sexlevels)

pca_g1 <- subsetmakepca("gonads", alllevels, sexlevels)	
pca_g2 <- makefvizdf("gonads", alllevels, sexlevels)	
```

## Supplemental Data-driven genes vsds for box plots

```{r candidates}
# genes lists in "./analysis/03_datawrangling.Rmd" and old lists in "./R/genelists.R"

candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
  select(-X1) %>%
  filter(treatment %in% alllevels) %>%
  mutate(treatment = factor(treatment, levels = charlevels)) %>%
  drop_na() 
head(candidatevsd)
```

## Fig 3

```{r supplfig, fig.height=7.25, fig.width=7.25}

a <- plotpc12(pca_h1, pca_h2, pca_h1$treatment, allcolors, "Hypothalamus", NULL) 
c <- plotpc12(pca_p1, pca_p2, pca_p1$treatment, allcolors, "Pituitary", NULL) 
e <- plotpc12(pca_g1, pca_g2, pca_g1$treatment, allcolors, "Gonads", NULL)

b1 <- fviz_screeplot(pca_h2, addlabels = F, barfill = "grey", barcolor = "grey") +
  theme_B3() + labs(title = " ")
d1 <- fviz_screeplot(pca_p2, addlabels = F, barfill = "grey", barcolor = "grey") +
  theme_B3() + labs(title = " ")
f1 <- fviz_screeplot(pca_g2, addlabels = F, barfill = "grey", barcolor = "grey") +
  theme_B3() + labs(title = " ")

b2 <- candidateboxplot("hypothalamus", c("COX1"), sexlevels)
d2 <- candidateboxplot("pituitary", c("PRL"), sexlevels)
f2 <- candidateboxplot("gonad", c("OVALY"), sexlevels)

supplfig <- plot_grid(a,b1,b2, c,d1,d2,e,f1,f2, nrow = 3, rel_widths = c(2,1,2))
```

## Save files

```{r write}
pdf(file="../figures/fig1-1.pdf", width=7, height=4.5)
plot(fig1)
dev.off()

png("../figures/fig1-1.png", width = 7, height = 4.5, 
    units = 'in', res = 300)
plot(fig1) 
dev.off()

pdf(file="../figures/supplfig-1.pdf", width=7.5, height=7.5)
plot(supplfig)
dev.off()

png("../figures/supplfig-1.png", width = 7.5, height = 7.5, 
    units = 'in', res = 300)
plot(supplfig)  
dev.off()

sessionInfo()
```