### 2020-08-07 analysis/figure-3.Rmd
working on new fig 3 with all manip files, fig 4 is blank

---
title: "Replacement only"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(Rtsne)
library(ggsignif)
library(corrr)

source("../R/themes.R")
source("../R/functions.R")
source("../R/genelists.R")

knitr::opts_chunk$set(echo = TRUE, message = F, warning = F,
                      fig.path = '../figures/', fig.width = 7.25, fig.height = 7.25)
```

# Figure 3 and 4. What are the effects of removal  and replacement?

## import limma counts and sample info

```{r data}
countData <- read_csv("../results/01_limma.csv") %>%
  column_to_rownames(var = "X1")
countData <- as.data.frame(t(countData))
head(countData[1:3])

colData <- read_csv("../metadata/00_colData.csv") %>%
  mutate(treatment = factor(treatment, levels = alllevels),
         tissue = factor(tissue, levels = tissuelevels)) %>% 
  column_to_rownames(var = "X1") 
head(colData)
# check ready for analysis
# row.names(countData) == row.names(colData)
head(row.names(countData) == row.names(colData))
```

## tsne

```{r tSNE}

# removal
hyptsne <- subsetmaketsne("hypothalamus", removallevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", removallevels, sexlevels)
gontsne <- subsetmaketsne("gonads", removallevels, sexlevels)

pittsne <- addgroupings(pittsne)

t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
  facet_wrap(~sex) 
t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank())
t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank())
t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
t456

# replacement
hyptsne <- subsetmaketsne("hypothalamus", timinglevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", timinglevels, sexlevels)
gontsne <- subsetmaketsne("gonads", timinglevels, sexlevels)

pittsne <- addgroupings(pittsne)

t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
  facet_wrap(~sex) 
t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank())
t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank())
t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
t456

```

## variance stabilized gene expression  (vsd)

```{r vsd}
vsd_path <- "../results/DEseq2/treatment/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") 
## before pivoting, check names of df with
## head(names(allvsd)) and tail(names(allvsd))
allvsd <- allvsd %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts") 
allvsd <- as_tibble(allvsd)
head(allvsd[2:4])
tail(allvsd[2:4])

```

```{r candidatevsd}

# for plotting gene expression over time
getcandidatevsdmanip <- function(whichgenes, whichtissue){

  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2),
                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
                treatment =  factor(treatment, levels = alllevels),
                earlylate = fct_collapse(treatment, 
                                  early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
                                         "inc.d9",  "bldg"),
                                  late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
                                         "m.n2", "extend", "n5"),
                                  reference = "control"),
                extint = fct_collapse(treatment, 
                                  nest = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"),
                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong"),
                                  chicks = c("early", "hatch", "extend", "n5"),
                                  reference = "control"))  %>%
    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples)  
  return(candidates)
}

candidatevsd <- getcandidatevsdmanip(candidategenes)

hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus")  
pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary") 
gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")

hypvsdf <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "female")  
pitvsdf <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "female") 
gonvsdf <- candidatevsd %>% filter(tissue %in% "gonad", sex == "female")

hypvsdm <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "male")  
pitvsdm <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "male") 
gonvsdm <- candidatevsd %>% filter(tissue %in% "gonad", sex == "male")

# not all genes appear to be expressed in the gonads... these are
filter(gonvsd, treatment == "control") %>% head()
```

## candidate gene box plots

```{r}
plotcandidatemanp <- function(df, whichgene, whichtissue){

  p <- df %>%
    dplyr::filter(gene  == whichgene)  %>%
    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
    geom_boxplot(outlier.shape = NA, lwd=0.5) +
    facet_wrap(~sex) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    theme_B3() +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(face = "italic")) +
    scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
    labs(title = paste(whichtissue, sep = " "), subtitle = whichgene, y = "gene expression") +
    geom_signif(comparisons = list(c("inc.d9", "early"), 
                                  c( "inc.d17", "prolong"), 
                                  c( "prolong", "hatch")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075) +
  geom_signif(comparisons = list(c( "inc.d3", "m.inc.d3"),
                                   c( "inc.d9", "m.inc.d9"),
                                   c( "inc.d17", "m.inc.d17"),
                                   c( "hatch", "m.n2")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0,  size = 0.5, step_increase = 0.075) +
    geom_signif(comparisons = list(c( "hatch", "extend"), 
                                  c( "n5", "extend")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075)

  return(p)

}

a <- plotcandidatemanp(hypvsdf, "HTR2C", "hypothalamus") 
c <- plotcandidatemanp(pitvsdf, "PRL", "pituitary")
e <- plotcandidatemanp(gonvsdf, "ESR2", "gonads") 
b <- plotcandidatemanp(hypvsdm, "HTR2C", "hypothalamus") 
d <- plotcandidatemanp(pitvsdm, "PRL", "pituitary")
f <- plotcandidatemanp(gonvsdm, "ESR2", "gonads") 
boxplots3 <- plot_grid(a,b,c,d,e,f, nrow = 3, 
                       labels = c("A", "", "B", " ", "C"),
                       hjust = 0, label_size = 8)
boxplots3

```

## DEGs

```{r allDEGs}
DEG_path <- "../results/DEseq2/treatment/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) # %>%
 # mutate(tissue = factor(tissue, levels = tissuelevel),
  #       comparison = factor(comparison , levels = comparisonlevelsremoval),
  #       direction = factor(direction, levels = alllevels))
head(allDEG)

candidateDEGS <- allDEG %>%
  filter(gene %in% parentalcaregenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", 
                         "gonad" = "G", "gonads" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, inc.d3_m.inc.d3, inc.d9_m.inc.d9, inc.d17_m.inc.d17, hatch_m.n2,
         inc.d17_prolong, hatch_prolong,extend_hatch, n5_extend, 
         early_inc.d9 ,hatch_early)
head(candidateDEGS[1:5])
```

## make figure

```{r fig4, fig.width=7.25, fig.height=7.25}

design <- png::readPNG("../figures/images/fig_fig4a.png")
design <- ggdraw() +  draw_image(design, scale = 1)

boxplots <- plot_grid(boxplots1, boxplots2, nrow = 1, rel_widths = c(1.5,1),
                      labels = "AUTO", label_size = 8)
fig4 <- plot_grid(boxplots, design, t456, ncol = 1, rel_heights = c(2.75,1.3,1),
                      labels = c(" ", " " , "C"), label_size = 8)
fig4

```

## suppl fig 1

```{r supplfig-4}

comparisonlevelsmanip <- c("inc.d3_m.inc.d3", "inc.d9_m.inc.d9", "inc.d17_m.inc.d17", "hatch_m.n2",
         "early_inc.d9" ,"hatch_early", "inc.d17_prolong", "hatch_prolong","extend_hatch", "n5_extend")

# for fig 1
DEGmanip <- allDEG %>% 
  filter(comparison %in% comparisonlevelsmanip) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsmanip))

DEGmanip %>% 
  group_by(comparison, sex, tissue) %>%
  summarize(totalDEGs = n())  %>%
  arrange(desc(totalDEGs))

s1a <- makebargraph(DEGmanip, "hypothalamus","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank()) + 
  labs(subtitle = "hypothalamus") + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey")

s1b <- makebargraph(DEGmanip, "pituitary","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), strip.text = element_blank()) + 
  labs(subtitle = "pituitary")  + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey")

s1c <- makebargraph(DEGmanip, "gonad","DEGs w/ + LFC", 0, 3200, DEGmanip$comparison) + 
  theme(strip.text = element_blank()) + 
  labs(subtitle = "gonad", x = "Comparison of parental stage and manipulation groups") + geom_vline(xintercept = 4.5, linetype = "dashed", color = "grey") + 
  scale_x_discrete(breaks= comparisonlevelsmanip,
                      labels= comparisonlevelsmanip,
                   drop=FALSE)

contrasts <- png::readPNG("../figures/images/fig_fig4sup1.png")
contrasts <- ggdraw() +  draw_image(contrasts, scale = 1)

supplfig4 <- plot_grid(contrasts, s1a, s1b, s1c, ncol = 1, rel_heights = c(0.5, 1,1,1.5))
supplfig4

```

## write files

```{r write}

pdf(file="../figures/fig4-1.pdf", width=7.25, height=7.25)
plot(fig4)
dev.off()

pdf(file="../figures/supplfig-4-1.pdf", width=5, height=5)
plot(supplfig4)
dev.off()

write.csv(candidateDEGS, "../results/table2.csv", row.names = F)
```