### 2020-08-21 analysis/figure-2.Rmd
figs 2-4 made with 1 script

---
title: "Fig 2"
output: md_document
---

# Fig 2

```{r setup, include=T}
library(tidyverse)
library(cowplot)
library(ggsignif)

source("../R/themes.R")
source("../R/functions.R")
source("../R/genelists.R")

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

## Treatment specific DEGs

```{r suppltables}
allDEG <- read_csv("../results/03_allDEG.csv") %>%
  mutate(tissue = factor(tissue, levels = tissuelevel),
         direction = factor(direction, levels = alllevels),
         label = gsub("_","\nvs. ", comparison)) %>%
  filter(lfc > 1.1 | lfc < -1.1 )
head(allDEG)

# for suppl figures
DEGcontrol <- allDEG %>% 
  filter(grepl("control", comparison),
         !grepl("m.|early|extend|prolong", comparison))  %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelscontrol)) %>%
  group_by(sex, tissue, comparison, direction, label) %>%
    summarise(n = n()) %>%
    mutate(n = ifelse(direction == "control", n*-1, n*1 ))

DEGbldg <- allDEG %>% 
  filter(grepl("bldg", comparison),
         !grepl("m.|early|extend|prolong|control", comparison))  %>%
  drop_na() %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsbldg))  %>%
  group_by(sex, tissue, comparison, direction, label) %>%
    summarise(n = n()) %>%
    mutate(n = ifelse(direction == "bldg", n*-1, n*1 ))

DEGchar <- allDEG %>% 
  filter(comparison %in% comparisonlevelschar,
         !grepl("control|bldg", comparison)) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelschar)) %>%
  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
  group_by(sex, tissue, comparison, direction, label, updown) %>%
  summarise(n = n()) %>%
  mutate(n = n*updown ) %>%
  select(-updown)

DEGremove <- allDEG %>% 
  filter(comparison %in% comparisonlevelsremoval) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsremoval)) %>%
  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
  group_by(sex, tissue, comparison, direction, label, updown) %>%
  summarise(n = n()) %>%
  mutate(n = n*updown ) %>%
  select(-updown)
DEGremove

DEGreplace <- allDEG %>% 
  filter(comparison %in% comparisonlevelsreplace) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsreplace)) %>%
  mutate(updown = ifelse(lfc > 0, 1, -1)) %>%
  group_by(sex, tissue, comparison, direction, label, updown) %>%
  summarise(n = n()) %>%
  mutate(n = n*updown ) %>%
  select(-updown)
DEGreplace
```

## Variance stabilized data

```{r data}
countData <- read_csv("../results/01_limma.csv") %>%
  column_to_rownames(var = "X1")
countData <- as.data.frame(t(countData))
head(countData[1:3])

colData <- read_csv("../metadata/00_colData.csv") %>%
  mutate(treatment = factor(treatment, levels = alllevels),
         tissue = factor(tissue, levels = tissuelevels)) %>% 
  column_to_rownames(var = "X1") 
head(colData)
# check ready for analysis
# row.names(countData) == row.names(colData)
head(row.names(countData) == row.names(colData))

vsd_path <- "../results/DEseq2/treatment/"   # path to the data
vsd_files <- dir(vsd_path, pattern = "*vsd.csv") # get file names
vsd_pathfiles <- paste0(vsd_path, vsd_files)
vsd_files

allvsd <- vsd_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name")  %>% 
  dplyr::rename("gene" = "X1") 
## before pivoting, check names of df with
## head(names(allvsd)) and tail(names(allvsd))
allvsd <- allvsd %>% 
  pivot_longer(cols = L.G118_female_gonad_control:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "samples", values_to = "counts")  %>% 
    drop_na() 
allvsd <- as_tibble(allvsd)
head(allvsd[2:4])
tail(allvsd[2:4])

# for plotting gene expression over time
getcandidatevsdmanip <- function(whichgenes, whichtissue){

  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/treatment/'), "[", 2),
                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                sex = factor(sex),
                  tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
                treatment =  factor(treatment, levels = alllevels),
                earlylate = fct_collapse(treatment, 
                                  early = c("early", "m.inc.d3", "m.inc.d9",
                                           "lay", "inc.d3", "inc.d9" ),

                                  late = c("inc.d17", "m.inc.d17", "prolong" ,  "hatch" , 
                                         "m.n2", "extend", "n5", "n9"),
                                  reference = c("control", "bldg")),
                extint = fct_collapse(treatment, 
                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17"),
                                  eggs = c("lay","inc.d3", "inc.d9", "inc.d17", "prolong"),
                                  chicks = c("early", "hatch", "extend", "n5", "n9"),
                                  reference = c("control", "bldg")))  %>%
    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples) 
  return(candidates)
}

candidatevsd <- getcandidatevsdmanip(candidategenes)

hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus")  
pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary") 
gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")

hypvsdf <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "female")  
pitvsdf <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "female")
gonvsdf <- candidatevsd %>% filter(tissue %in% "gonad", sex == "female")  
hypvsdm <- candidatevsd %>% filter(tissue %in% "hypothalamus", sex == "male")  
pitvsdm <- candidatevsd %>% filter(tissue %in% "pituitary", sex == "male") 
gonvsdm <- candidatevsd %>% filter(tissue %in% "gonad", sex == "male")

summary(hypvsd)
summary(pitvsd)
```

```{r fig2, fig.width=7.5, fig.height=7.25}
a <- plot.volcano("hypothalamus", sexlevels,  "control_bldg") + 
  facet_wrap(~sex)

b <- makebargraphv3(DEGcontrol, "hypothalamus","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
  labs(x = "Control versus all other reproductive and parental stages",
       title = "Hypothalamus") +
  geom_rect(mapping=aes(xmin=0.5, xmax=1.5, ymin=-1000, ymax = 350, fill = F), color="black", alpha=0.5) +
  annotate("text", x = 1, y = -1075, label = "D", size = 2.5)

c <- makebargraphv3(DEGbldg, "hypothalamus", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
  labs(x = "Nest-building versus all other parental stages")
d <- makebargraphv3(DEGchar, "hypothalamus", NULL,  comparisonlabelscharnobldg) +
  labs(x = "Comparison of sequential parental stages")

e <- makebargraphv3(DEGremove, "hypothalamus", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsremoval) +
  labs(x = "Offspring removal versus temporal control")
f <- makebargraphv3(DEGreplace, "hypothalamus", NULL,  comparisonlevelsreplace) +
  labs(x = "Offspring removal versus temporal or external control")

g <- plotcandidatechar(hypvsdf, "AVP") + labs(subtitle = "Hypothalamus")
h <- plotcandidatechar(hypvsdm, "AVP") + labs(y = NULL, x = "") + labs(subtitle = " ")
i <- plotremoval(hypvsdf, "AVP")+ labs(y = NULL) + labs(subtitle = " ")
j <- plotremoval(hypvsdm, "AVP") + labs(y = NULL, x = "")+ labs(subtitle = " ")
k <- plotreplacement(hypvsdf, "AVP") + labs(y = NULL)+ labs(subtitle = " ")
l <- plotreplacement(hypvsdm, "AVP") + labs(y = NULL, x = "")+ labs(subtitle = " ")

ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
                labels = c("D", "E"), label_size = 8, hjust = 0)
cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
                labels = c("F", "G"), label_size = 8, hjust = 0)
ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
                labels = c("H", "I"), label_size = 8, hjust = 0)

#h <- boxplotextint(hypvsdf, "HTR2C", " ")

ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
                rel_widths = c(9,9,8,8,6,6))

fig2 <- plot_grid(ghi,ab,cd,ef,  ncol = 1)
fig2
```

```{r fig3, fig.width=7.5, fig.height=7.25}
a <- plot.volcano("pituitary", sexlevels,  "control_bldg") + 
  facet_wrap(~sex)

b <- makebargraphv3(DEGcontrol, "pituitary","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
  labs(x = "Control versus all other reproductive and parental stages",
       title = "Pituitary")

c <- makebargraphv3(DEGbldg, "pituitary", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
  labs(x = "Nest-building versus all other parental stages")
d <- makebargraphv3(DEGchar, "pituitary", NULL,  comparisonlabelscharnobldg) +
  labs(x = "Comparison of sequential parental stages")

e <- makebargraphv3(DEGremove, "pituitary", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsremoval) +
  labs(x = "Offspring removal versus temporal control")
f <- makebargraphv3(DEGreplace, "pituitary", NULL,  comparisonlevelsreplace) +
  labs(x = "Offspring removal versus temporal or external control")

g <- plotcandidatechar(pitvsdf, "PRL") + labs(subtitle = "Pituitary")
h <- plotcandidatechar(pitvsdm, "PRL") + labs(y = NULL, x = "") + labs(subtitle = " ")
i <- plotremoval(pitvsdf, "PRL")+ labs(y = NULL) + labs(subtitle = " ")
j <- plotremoval(pitvsdm, "PRL") + labs(y = NULL, x = "")+ labs(subtitle = " ")
k <- plotreplacement(pitvsdf, "PRL") + labs(y = NULL)+ labs(subtitle = " ")
l <- plotreplacement(pitvsdm, "PRL") + labs(y = NULL, x = "")+ labs(subtitle = " ")

ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
                labels = c("D", "E"), label_size = 8, hjust = 0)
cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
                labels = c("F", "G"), label_size = 8, hjust = 0)
ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
                labels = c("H", "I"), label_size = 8, hjust = 0)

ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
                rel_widths = c(9,9,8,8,6,6))

fig3 <- plot_grid(ghi,ab,cd,ef,  ncol = 1)
fig3
```

```{r fig4, fig.width=7.5, fig.height=7.25}
a <- plot.volcano("gonad", sexlevels,  "control_bldg") + 
  facet_wrap(~sex)

b <- makebargraphv3(DEGcontrol, "gonad","No. of DEGs\n(-) decreased  increased (+)", comparisonlabelscontrol) +
  labs(x = "Control versus all other reproductive and parental stages",
       title = "Pituitary")

c <- makebargraphv3(DEGbldg, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlabelsbldg) +
  labs(x = "Nest-building versus all other parental stages")
d <- makebargraphv3(DEGchar, "gonad", NULL,  comparisonlabelscharnobldg) +
  labs(x = "Comparison of sequential parental stages")

e <- makebargraphv3(DEGremove, "gonad", "No. of DEGs\n (-)decreased  increased (+)", comparisonlevelsremoval) +
  labs(x = "Offspring removal versus temporal control")
f <- makebargraphv3(DEGreplace, "gonad", NULL,  comparisonlevelsreplace) +
  labs(x = "Offspring removal versus temporal or external control")

g <- plotcandidatechar(pitvsdf, "ESR1") + labs(subtitle = "Gonads")
h <- plotcandidatechar(pitvsdm, "ESR1") + labs(y = NULL, x = "") + labs(subtitle = " ")
i <- plotremoval(pitvsdf, "ESR1")+ labs(y = NULL) + labs(subtitle = " ")
j <- plotremoval(pitvsdm, "ESR1") + labs(y = NULL, x = "")+ labs(subtitle = " ")
k <- plotreplacement(pitvsdf, "ESR1") + labs(y = NULL)+ labs(subtitle = " ")
l <- plotreplacement(pitvsdm, "ESR1") + labs(y = NULL, x = "")+ labs(subtitle = " ")

ab <- plot_grid(a,b,rel_widths = c(1,2.5), 
                labels = c("D", "E"), label_size = 8, hjust = 0)
cd <- plot_grid(c,d,rel_widths = c(1.1,1), align = "h",
                labels = c("F", "G"), label_size = 8, hjust = 0)
ef <- plot_grid(e,f,rel_widths = c(1.1,1), align = "h",
                labels = c("H", "I"), label_size = 8, hjust = 0)

ghi <- plot_grid(g,h,i, j,k,l,nrow = 1,
                labels = c("A", "", "B", "", "C"), label_size = 8, hjust = 0,
                rel_widths = c(9,9,8,8,6,6))

fig4 <- plot_grid(ghi,ab,cd,ef,  ncol = 1)
fig4
```

## Save files

```{r write}
pdf(file="../figures/fig2-1.pdf", width=7, height=7)
plot(fig2)
dev.off()

png("../figures/fig2-1.png", width = 7, height = 7, 
    units = 'in', res = 300)
plot(fig2) 
dev.off()

pdf(file="../figures/fig3-1.pdf", width=7, height=7)
plot(fig3)
dev.off()

png("../figures/fig3-1.png", width = 7, height = 7, 
    units = 'in', res = 300)
plot(fig3) 
dev.off()

pdf(file="../figures/fig4-1.pdf", width=7, height=7)
plot(fig4)
dev.off()

png("../figures/fig4-1.png", width = 7, height = 7, 
    units = 'in', res = 300)
plot(fig4) 
dev.off()

sessionInfo()
```